### 2023-11-30 outputs/paper/paper.qmd
Update the paper.qmd

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(dplyr)
library(knitr)
```

```{r}
election_data <- read.csv(here("outputs/data/election_data_clean.csv"))
covid_data <- read.csv(here("outputs/data/covid_data_clean.csv"))
acs_data <- read.csv(here("outputs/data/acs_data_clean.csv"))
covid_election_data <- read.csv(here("outputs/data/covid_election.csv"))
data <- read.csv(here("outputs/data/merged_data.csv"))
republican_data <- data %>% 
  filter(party == "Republican")
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

In this paper, the data used mainly consist the voting patterns for 2020 US Federal Election, COVID infection rate and socio-economic variables in each county of US. The voting data is taken from the MIT Election Data Science Lab which includes the voting for each party in county level at each Federal Election from 2000 to 2020. For the COVID data, it is taken from the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University. The socio-economic data is taken from American Community Survey.

## MIT Election Data Science Lab

The MIT Election Data Science Lab is a Lab at MIT which collect and analyzing the election data to supports advances in election science. Their collected data is not limited to the presidential election results, but also covers the midterm or US senate results in either state or county level.

The data abstracted from MITEDSL in this paper is the "County Presidential Election Returns 2000-2020." It contains the number of votes for each party and the total votes at each US county from 2000 to 2020 US Federal Election. In this paper, I will only use the data for the 2020 US Election and focus on the counties on the mainland of US. The @tbl-election summarizes the important variables.

```{r}
#| echo: false
#| message: false
#| label: tbl-election
#| tbl-cap: The summary of voting patterns in 2020 US Federal Presidential Election
election_data %>% 
  group_by(party, candidate) %>% 
  summarise(total = sum(votes),
            mean = round(mean(pct_vote), 2),
            median = round(median(pct_vote), 2)) %>%
  arrange(desc(total)) %>% 
  rename(`Party` = party,
         `Candidate` = candidate,
         `Total Votes` = total,
         `Mean Pct Vote` = mean,
         `Median Pct Vote` = median) %>% 
  kable()
```

The @tbl-election summarizes the voting patterns for each party during the 2020 US Presidential Election. Undoubtedly, the two most popular parties are the Democrat and the Republican. However, it seems that even though Republican has almost double mean percent vote than the Democrat, the total votes for Republican is less than the Democrat. This pattern may indicate that more people living at the high-population states such as California votes for Biden. This finding aligns with the fact that Biden defeated Trump in 2020. I will explain more latter.

## Center for Systems Science and Engineering at JHU
The Center for Systems Science and Engineering at JHU is the center at the Department of Civils and Engineering to collect the local, national and global multidimensional data including medicine, health care, disaster response etc. During the pandemic, they collected the US and global COVID cases and deaths and report them on their GitHub. Their data is summarized by daily reports, ranging from April 12, 2020 to March 9, 2023.

To illustrate the impact of COVID on the 2020 Election to the greatest extend, the COVID data used in this paper will be the daily report on November 3, 2020 which is the Election Day [citation]. This report contains the aggregated cases and deaths for each each country and counties in US.

## American Community Survey (ACS)
The American Community Survey is the survey conducted by the U.S. Census Bureau. The survey contains a variety of socio-economic variables for each county. Although there are different data tables from ACS, I will take 2021 five years estimate of DP02, DP03 and DP05.

These three tables cover the social, economic and demographic characteristics of each county in US. By referring my  previous research where I found the variables which have the highest correlations to the COVID mortality rate, I will use the same variables in this paper. That said, I will extract the data regarding the educational attainment from DP02, especially the proportion of people having at least a bachelor degree. In terms of economic perspective (DP03), I will take the proportion of people with private health 
```{r}
#| echo: false
#| message: false
#| label: tbl-covid
#| tbl-cap: The summary of COVID cases and deaths as of Election Day.
data %>% 
  group_by(winning_party) %>% 
  summarise(case = mean(cases),
            deaths = mean(deaths),
            inf = mean(infrate),
            mort = mean(mortrate))
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

# analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
# 
# analysis_data |> 
#   ggplot(aes(x = width, y = length)) +
#   geom_point(alpha = 0.8) +
#   theme_minimal() +
#   labs(x = "Wing width (mm)",
#        y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

highinf_model <- glm(high_infrate ~ prop_higher_education + pctile + 
                             no_insurance +private_insurance + males + age_85 + 
                             white_pct + black_pct, data = republican_data)
summary(highinf_model)
# highmort_model <- glm(high_mortrate ~ prop_higher_education + pctile + 
#                              no_insurance +private_insurance + males + age_85 + 
#                              white_pct + black_pct, data = data, nboots = 10)
# summary(highmort_model)
republican_data$prop_score <- predict(highinf_model, type = "response")
```

```{r}
library(Matching)
# Propensity score matching
# mb <- MatchBalance(high_infrate ~ prop_higher_education + pctile + 
#                              no_insurance +private_insurance + males + age_85 + 
#                              white_pct + black_pct, data = data)
rr <- Match(Y = republican_data$pct_vote, Tr = republican_data$high_infrate, 
            X = republican_data$prop_score, M = 1)
# mb <- MatchBalance(high_infrate ~ prop_higher_education + pctile + 
#                               no_insurance +private_insurance + males + age_85 + 
#                               white_pct + black_pct, match.out = rr,
#                    data = republican_data, nboots = 10)
summary(rr)
```

```{r}
# Propensity score regression adjustment
model_adj <- glm(pct_vote ~ high_infrate + prop_score, data = republican_data)
summary(model_adj)
```

```{r}
# Counterfactual Analysis
adjusted_data <- data %>% 
  dplyr::select(state, county.x, fips, party, votes, total_votes, pct_vote, high_infrate)

adjust_function <- function(data){
  adjust_votes <- c()
  for (i in 1:nrow(data)){
    if (data[i,"high_infrate"] == 1){
      if (data[i,"party"] == "Republican"){
        adjust_votes[i] <- round((data[i,"pct_vote"] + 0.036928 ) * data[i,"total_votes"], 0)
      }
      else if (data[i,"party"] == "Democrat"){
        adjust_votes[i] <- round((data[i,"pct_vote"] - 0.036928 ) * data[i,"total_votes"], 0)
      }
      else {adjust_votes[i] <- data[i,"votes"]}
    }
    else {
      adjust_votes[i] <- data[i,"votes"]
    }
  }
  return(adjust_votes)
}

adjusted_data %>% 
  group_by(state, party) %>% 
  summarise(votes = sum(votes), .groups = "drop")
```

```{r}
adjust_function2 <- function(data) {
  adjust_votes <- numeric(nrow(data))  # Initialize the vector with the correct length
  extra_vote_percentage <- 0.036928

  for (i in 1:nrow(data)) {
    if (data[i, "high_infrate"] == 1) {
      total_votes <- data[i, "total_votes"]
      extra_votes <- round(extra_vote_percentage * total_votes, 0)

      if (data[i, "party"] == "Republican") {
        adjust_votes[i] <- data[i, "votes"] + extra_votes
      } else {
        # Calculate total votes for non-Republican parties
        total_non_rep_votes <- sum(data[data$party != "Republican" & data$fips == data[i, "fips"], "votes"])

        # Distribute the vote loss proportionally
        if (total_non_rep_votes > 0) {
          party_vote_share <- data[i, "votes"] / total_non_rep_votes
          votes_lost <- round(extra_votes * party_vote_share, 0)
          adjust_votes[i] <- max(0, data[i, "votes"] - votes_lost)
        } else {
          adjust_votes[i] <- data[i, "votes"]
        }
      }
    } else {
      adjust_votes[i] <- data[i, "votes"]
    }
  }

  return(adjust_votes)
}

adjust_vote1 <- adjust_function(adjusted_data)
adjust_vote2 <- adjust_function2(adjusted_data)
adjusted_data$adjust_vote1 <- adjust_vote1
adjusted_data$adjust_vote2 <- adjust_vote2
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# References

### 2023-12-01 outputs/paper/paper.qmd
Finish the drafts for Data and Methods

@@
 library(knitr)
+library(ggplot2)
+library(maps)
+library(mapdata)
+library(RColorBrewer)
+library(gridExtra)
+library(tools)
 ```
@@
 ## Center for Systems Science and Engineering at JHU
+
 The Center for Systems Science and Engineering at JHU is the center at the Department of Civils and Engineering to collect the local, national and global multidimensional data including medicine, health care, disaster response etc.
@@
 
-To illustrate the impact of COVID on the 2020 Election to the greatest extend, the COVID data used in this paper will be the daily report on November 3, 2020 which is the Election Day [citation].
+To illustrate the impact of COVID on the 2020 Election to the greatest extend, the COVID data used in this paper will be the daily report on November 3, 2020 which is the Election Day \[citation\].
 This report contains the aggregated cases and deaths for each each country and counties in US.
@@
 ## American Community Survey (ACS)
+
 The American Community Survey is the survey conducted by the U.S.
@@
 That said, I will extract the data regarding the educational attainment from DP02, especially the proportion of people having at least a bachelor degree.
-In terms of economic perspective (DP03), I will take the proportion of people with private health
+In terms of economic perspective, I will take the proportion of people with private health insurance and the mean household income.
+Lastly, regarding the demographic factors, I will use the total population, the proportions of children and individuals above 85.
+I will also take the White and Black people percentage as well.
+
+## Data Cleaning
+
 ```{r}
@@
             inf = mean(infrate),
-            mort = mean(mortrate))
+            mort = mean(mortrate),
+            income = mean(mean_household_income))
 ```
 
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+Combining the three data sources, @tbl-covid shows the COVID cases and deaths, as well as the infection and mortality rate for the county voting for the Democrat and Republican party, respectively.
+Surprisingly, the average case and deaths of COVID of the counties voting for the Democrat is significantly higher than than the Republican party.
+However, the values of infection and mortality rate does not show the same dramatic difference.
+It also emphasizes that the counties (states) voting for the Democrat are population intensive and the counties having more residence generally have higher mean household income level.
+We can observe this from the average income where the "Democratic" counties have about \$13k mean income higher than the "Republic" ones.
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
 #| echo: false
-#| warning: false
+#| message: false
+#| label: fig-boxplot
+#| fig-cap: Summary of income levels for counties voting for Democratic and Republican
+
+data  %>%
+  ggplot(mapping = aes(x = winning_party, y = mean_household_income)) +
+  geom_boxplot() +
+  geom_jitter(alpha = 0.04, width = 0.3, height = 0) +
+  theme_minimal() +
+  labs(x = "Voting Party",
+       y = "Death Rate per 1000")
+```
+
+@fig-boxplot compares the distribution of county income levels for the two parties.
+In general, counties voting for the Democrat has a higher income levels than the Republican.
+In addition, we can observe that the counties are intensively concentrated around income levels \$50k to \$75k, compared to the Democrat where the counties are approximately uniform distributed at each income level.
+Furthermore, there is barely no counties having at least 150k voting for the Republican party.
+Both @tbl-covid and @fig-boxplot indicates that the Democrat is in favor of rich counties but poorer counties for Republican.
+Consequently, the richer counties (states) usually have more electoral votes than the poorer ones, this may provide insighes why Trump lost.
+
+```{r}
+#| echo: false
 #| message: false
 
-# analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
-# 
-# analysis_data |> 
-#   ggplot(aes(x = width, y = length)) +
-#   geom_point(alpha = 0.8) +
-#   theme_minimal() +
-#   labs(x = "Wing width (mm)",
-#        y = "Wing length (mm)")
+get_state_abbreviation <- function(state_names) {
+  sapply(state_names, function(state_name) {
+    state_data <- state.abb[match(state_name, state.name)]
+    if (!is.na(state_data)) {
+      return(state_data)
+    } else {
+      return(NA)  # changed from NULL to NA for consistency in a vector
+    }
+  })
+}
+
+county <- map_data("county")
+state <- map_data("state")
+county_clean <- county %>% 
+  mutate(state = region,
+         county = subregion,
+         state = get_state_abbreviation(toTitleCase(state)),
+         county = toTitleCase(county)) %>% 
+  filter(region != "AK") %>% 
+  select(-region, -subregion)
+
+merged_data <- county_clean %>% left_join(republican_data, by = c("state", "county"))
 ```
 
-Talk way more about it.
+```{r}
+#| echo: false
+#| message: false
+#| fig.width: 10 
+#| fig.height: 10
+#| label: fig-map
+#| fig-cap: The ratio of votes for the Republican and the infection rate per 100k in each county
 
-# Model
+p1 <- ggplot(data = merged_data, aes(x = long, y = lat, group = group)) + 
+  geom_polygon(aes(fill = ratio_rd), color = "black") +
+  scale_fill_distiller(palette = "RdBu", direction = -1, na.value = "grey") +
+  labs(fill = 'Share of Republican') +
+  coord_quickmap() +
+  theme_void() +
+  geom_path(data = state, aes(x = long, y = lat, group = group), color = "black", linewidth = 0.5)
 
-The goal of our modelling strategy is twofold. Firstly,...
+# Second plot
+p2 <- ggplot(data = merged_data, aes(x = long, y = lat, group = group)) + 
+  geom_polygon(aes(fill = infrate), color = "black") +
+  scale_fill_gradient(low = "white", high = "darkblue", na.value = "grey",
+                      limits = c(0, 5000), oob = scales::squish) +
+  labs(fill = 'Infection Rate per 100k') +
+  coord_quickmap() +
+  theme_void() +
+  geom_path(data = state, aes(x = long, y = lat, group = group), color = "black", linewidth = 0.5)
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+# Arrange the plots in a 2x1 layout
+grid.arrange(p1, p2, ncol = 1)
+```
 
-## Model set-up
+## Data Limitation
 
-### Model justification
+Although the merged data covers the voting patterns, socio-economic characteristics and COVID infection rate by county, some of the counties are missing.
+From the above two maps, the counties on the middle west shows a grey color, indicating that the data for these counties are absent.
+This absence of data may increase the bias and hence influence the accuracy of the data analyses.
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+# Methods
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+This paper aims to conduct the casual inference between the 2020 voting patterns and COVID, to find whether Trump can contribute his loss to the COVID.
+To find the casual effect in the observational study, I will implement the methods of propensity scores to find the change of votes for Trump due to the pandemic.
+Based on the results, I will also conduct a counterfactual analysis to see what would be happened if there was no COVID,
+
+First and foremost, we need to define the treatment.
+The treatment in an experiment is something that will be artificially imposed to a certain groups of people to see the effect of interested variable with and without this treatment.
+Therefore the people with the treatment is called treatment group and the rest are called the control group.
+The settings of treatment and control group is to test whether the imposed factor really have impacts on the interested variable.
+However, since this is an observational study, it is unrealistic to impose a factor manually.
+In order to find the the casual effect of COVID on Trump's voting, I will choose the treatment to be whether a county has a high infection rate.
+
+## Propensity Score
+
+Propensity score is a technique that can help us to decide which county would be in the treatment group.
+That said, the propensity score in this study is the probability of a county having a high infection rate given a set of socioeconomic variables.
+By using the propensity score, it can ensure that the two counties with similar propensity score having the same distribution of the observed socio-economic variables such as income level even though they are in the different group.
+This is fantastic as we can have may pairs from treatment of control group which have similar covariates by fixing propensity scores.
+The key drawback that we can not estimate the casual effect from observational studies is addressed.
+
+In addition to that, by finding the propensity scores, we can also reconstruct our assignment mechanism to a strongly ignorable one.
+An assignment mechanism is "strongly ignorable" if and only if, given the observed covariates, the potential outcomes are independent of the treatment assignment.
+This indicates that the change of voting for Trump is independent to our assignment that a county has high or low infection rate.
+
+As the propensity score represents the probability of a county having high infection rate, therefore we need to find a logistic regression to predict this probability.
+On my previous paper, I already find the best model predicting the mortality rate of COVID in each US county.
+Even though that one was to predict the mortality rate, however, the mortality is correlated with infection rate.
+Therefore I will take the dependent variables in that model to be the covariates in predicting the propensity score for each county, which is:
+
+```{=tex}
+\begin{align*}
+\frac{PS}{1-PS} &=  \beta_0 + \\
+& + \beta_1 \times \text{prop\_higher\_education} \\
+& + \beta_2 \times \text{IncomePctile} \\
+& + \beta_3 \times \text{no\_insurance} \\
+& + \beta_4 \times \text{private\_insurance} \\
+& + \beta_5 \times \text{males} \\
+& + \beta_6 \times \text{old\_85} \\
+& + \beta_7 \times \text{white\_pct} \\
+& + \beta_8 \times \text{black\_pct}
+\end{align*}
+```
+
+Using the above model, we can therefore find the PS for each county. However, there are three different ways to use the propensity scores to find the treatment effect.
+
+### Propensity Score Matching
+The package needed to implement the matching is the `matching`.
+Using this package, we can match the observations in treatment and control group having the same or similar propensity scores.
+Then take only take the matched pairs to be the new data set and find the find the treatment effect between them.
+In addition, we can also check the balance between the treatment and control groups.
+
+However, the most significant drawback is that the new data set can only contain the matched pairs.
+We may lose a huge amount of observations during this process.
+Our predictions may also not be accurate.
+
+### Propensity Score Stratification
+As introduced above that using PSM may cause the waste of data and results in bias.
+However, propensity score stratification will use the entire data set to avoid the waste.
+The way to do this is to split the data into different quantiles and fit a model for each subclass with the only dependent variable is the treatment variable.
+Then calculate the average casual effect for the subclasses.
+
+Even though the PSS can use the entire data set, however, the choice of quantiles will be closely related to the predictions.
+If we have a large sample and want to split the data into many subclasses, then this method may be computational slow.
+
+### Propensity Score Regression Adjustment
+The last method is the most straightforward, meaning that we just fit a regression model with treatment variable and propensity score be the two dependent variables.
+
+The above three methods have their unique advantages but the results may vary between different data sets. I will perform both the three methods and to compare which one would be the best.
+
+## Counterfactural Analysis
+By using the methodology of propensity score, we are able to access the casual effect of high infection rate on the votings for Trump.
+However, it can not tell us the election results and we can not make conclusions about the COVID and election.
+Therefore, I will conduct a counterfactural analysis, meaning that I will re-calculate the votes for Trump and Biden, to see if the final election results would be changed.
+
+The way I will use to re-calculate the votes for the two candidates is to follow the official rule of election.
+By finding the difference of voting percentage, I will re-allocate the votings for the two parties and calculate the electoral votes in each state using the rule of winner-takes-all.
 
@@
 adjusted_data <- data %>% 
-  dplyr::select(state, county.x, fips, party, votes, total_votes, pct_vote, high_infrate)
+  dplyr::select(state, county, fips, party, votes, total_votes, pct_vote, high_infrate)
 

### 2023-12-01 scripts/01-download_data.R
Update the R scrips

@@ -21,12 +21,18 @@ DP05 <- read.csv(url(url05), stringsAsFactors = FALSE)
 url_covid <- paste0(
   "https://raw.githubusercontent.com/CSSEGISandData/",
   "COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/",
-  "01-01-2021.csv")
+  "11-03-2020.csv")
 covid_data <- read.csv(url_covid, header = TRUE, sep = ",")
+
+url_covid2 <- paste0(
+  "https://raw.githubusercontent.com/CSSEGISandData/",
+  "COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/",
+  "01-01-2021.csv")
+covid_data2 <- read.csv(url_covid2, header = TRUE, sep = ",")
 
 #### Save data ####
 write.csv(DP02, "inputs/data/DP_02.csv", row.names = FALSE)
 write.csv(DP03, "inputs/data/DP_03.csv", row.names = FALSE)
 write.csv(DP05, "inputs/data/DP_05.csv", row.names = FALSE)
 write.csv(covid_data, "inputs/data/covid_data.csv", row.names = FALSE)
-         
+write.csv(covid_data2, "inputs/data/covid_data2.csv", row.names = FALSE)

### 2023-12-01 scripts/02-data_cleaning.R

@@ -20,6 +20,7 @@ DP02 <- read.csv("inputs/data/DP_02.csv")
 DP03 <- read.csv("inputs/data/DP_03.csv")
 DP05 <- read.csv("inputs/data/DP_05.csv")
 covid_data <- read.csv("inputs/data/covid_data.csv")
+covid_data2 <- read.csv("inputs/data/covid_data2.csv")
 election_data <- read.csv("inputs/data/countypres_2000-2020.csv")
 
 get_state_abbreviation <- function(state_names) {
@@ -92,19 +93,27 @@ DP05_clean <- data_cleaning(DP05, dict_05)
 # Clean the covid data
 covid_data_clean <- covid_data %>% 
   filter(Country_Region == "US") %>% 
-  select(FIPS, Admin2, Province_State, Confirmed, Deaths) %>% 
+  dplyr::select(FIPS, Admin2, Province_State, Confirmed, Deaths) %>% 
   rename(`county` = Admin2,
          `state` = Province_State,
          `cases` = Confirmed) %>%
   mutate(state = get_state_abbreviation(state)) %>% 
   clean_names() 
-  # mutate(county = if_else(state == "LA", paste0(county, " Parish"), county))
 covid_data_clean <- na.omit(covid_data_clean)
+
+covid_data2_clean <- covid_data2 %>% 
+  filter(Country_Region == "US") %>% 
+  dplyr::select(FIPS, Admin2, Province_State, Confirmed, Deaths) %>% 
+  rename(`county` = Admin2,
+         `state` = Province_State,
+         `cases` = Confirmed) %>%
+  mutate(state = get_state_abbreviation(state)) %>% 
+  clean_names() 
+covid_data2_clean <- na.omit(covid_data2_clean)
 
 # Clean the election data
 election_data_clean <- election_data %>% 
-  filter(year == 2020,
-         ! (party == "OTHER")) %>% 
+  filter(year == 2020) %>% 
   mutate(county = str_to_title(county_name),
          party = str_to_title(party),
          state = state_po,
@@ -112,25 +121,44 @@ election_data_clean <- election_data %>%
          total_votes = totalvotes, 
          fips = county_fips,
          pct_vote = votes/total_votes) %>%
-  select(state_po, county, fips, candidate, party, votes, total_votes, pct_vote)
+  dplyr::select(state_po, county, fips, candidate, party, votes, total_votes, pct_vote)
+
+winning_party <- election_data_clean %>%
+  group_by(fips) %>%
[…]

### 2023-12-01 outputs/paper/paper.qmd
Update the paper

--- 2023-11-30 outputs/paper/paper.qmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
-author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
-date: "`r Sys.time()`"
-date-format: "D MMMM YYYY"
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+title: "Draft"
+title-block-banner: true
+author: "Yiliu Cao"
+thanks: "Code and data from this analysis are available at: https://github.com/yiliuc/covid_and_trump_loss.git"
+date: "today"
+date-format: "long"
+abstract: "TBD"
 format: pdf
+toc: true
 number-sections: true
@@
 ```{r}
+#| echo: false
+#| message: false
+
 election_data <- read.csv(here("outputs/data/election_data_clean.csv"))
@@
 data <- read.csv(here("outputs/data/merged_data.csv"))
+election_2016_republican <- read.csv(here("outputs/data/election_data_2016.csv")) %>% 
+  filter(party == "Republican") %>% 
+  dplyr::select(fips, pct_vote_2016)
+
 republican_data <- data %>% 
-  filter(party == "Republican")
+  filter(party == "Republican") %>% 
+  inner_join(election_2016_republican, by = "fips") %>% 
+  mutate(change_vote = pct_vote - pct_vote_2016)
 ```
@@
 
-You can and should cross-reference sections and sub-sections.
+# Data
 
-The remainder of this paper is structured as follows. @sec-data....
-
-# Data {#sec-data}
-
-In this paper, the data used mainly consist the voting patterns for 2020 US Federal Election, COVID infection rate and socio-economic variables in each county of US.
-The voting data is taken from the MIT Election Data Science Lab which includes the voting for each party in county level at each Federal Election from 2000 to 2020.
-For the COVID data, it is taken from the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University.
-The socio-economic data is taken from American Community Survey.
+In this paper, the data used mainly consists of the voting patterns for the 2020 US Federal Election, COVID infection rate and socioeconomic variables in each county of the US.
+The voting data is taken from the MIT Election Data Science Lab, including voting for each party at the county level at each Federal Election from 2000 to 2020.
+The COVID data is from the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University.
+The socioeconomic data is taken from the American Community Survey.
 
@@
 
-The MIT Election Data Science Lab is a Lab at MIT which collect and analyzing the election data to supports advances in election science.
-Their collected data is not limited to the presidential election results, but also covers the midterm or US senate results in either state or county level.
+The MIT Election Data Science Lab is a lab at MIT that collects and analyzes election data to support advances in election science.
+Their collected data is not limited to the presidential election results but also covers the midterm or US Senate results at the state or county level.
 
-The data abstracted from MITEDSL in this paper is the "County Presidential Election Returns 2000-2020." It contains the number of votes for each party and the total votes at each US county from 2000 to 2020 US Federal Election.
-In this paper, I will only use the data for the 2020 US Election and focus on the counties on the mainland of US.
-The @tbl-election summarizes the important variables.
+The data abstracted from MITEDSL in this paper is the "County Presidential Election Returns 2000-2020." It contains the number of votes for each party and the total votes at each US county from 2000 to the 2020 US Federal Election.
+In this paper, I will only use the data for the 2020 US Election and focus on the counties on the mainland of the US.
+The @tbl-election summarizes the essential variables.
 
@@
 
-The @tbl-election summarizes the voting patterns for each party during the 2020 US Presidential Election.
-Undoubtedly, the two most popular parties are the Democrat and the Republican.
-However, it seems that even though Republican has almost double mean percent vote than the Democrat, the total votes for Republican is less than the Democrat.
-This pattern may indicate that more people living at the high-population states such as California votes for Biden.
+The @tbl-election summarizes the voting patterns for each party during the 2020 U.S.
+Presidential Election.
+Undoubtedly, the two most popular parties are the Democrats and the Republicans.
+However, even though Republican has almost double the mean percent vote than Democrat, the total votes for Republican is less than for Democrat.
+This pattern may indicate that more people living in high-population states, such as California, voted for Biden.
 This finding aligns with the fact that Biden defeated Trump in 2020.
-I will explain more latter.
+I will explain more later.
 
 ## Center for Systems Science and Engineering at JHU
-The Center for Systems Science and Engineering at JHU is the center at the Department of Civils and Engineering to collect the local, national and global multidimensional data including medicine, health care, disaster response etc.
-During the pandemic, they collected the US and global COVID cases and deaths and report them on their GitHub.
-Their data is summarized by daily reports, ranging from April 12, 2020 to March 9, 2023.
 
-To illustrate the impact of COVID on the 2020 Election to the greatest extend, the COVID data used in this paper will be the daily report on November 3, 2020 which is the Election Day [citation].
-This report contains the aggregated cases and deaths for each each country and counties in US.
+The Center for Systems Science and Engineering at JHU is at the Department of Civils and Engineering, which collects local, national, and global multidimensional data, including medicine, health care, disaster response, etc.
+During the pandemic, they collected the U.S.
+and international COVID cases and deaths and reported them on their GitHub.
+Their data is summarized by daily reports ranging from April 12, 2020, to March 9, 2023.
+
+To illustrate the impact of COVID-19 on the 2020 Election to the greatest extent, the COVID data used in this paper will be the daily report on November 3, 2020, which is Election Day \[citation\].
+This report contains the aggregated cases and deaths for each country and counties in the U.S.
 
 ## American Community Survey (ACS)
+
 The American Community Survey is the survey conducted by the U.S.
@@
 The survey contains a variety of socio-economic variables for each county.
-Although there are different data tables from ACS, I will take 2021 five years estimate of DP02, DP03 and DP05.
+Although there are different data tables from ACS, I will use the 2020 five-year estimate of DP02, DP03, and DP05.
 
-These three tables cover the social, economic and demographic characteristics of each county in US.
-By referring my  previous research where I found the variables which have the highest correlations to the COVID mortality rate, I will use the same variables in this paper.
-That said, I will extract the data regarding the educational attainment from DP02, especially the proportion of people having at least a bachelor degree.
-In terms of economic perspective (DP03), I will take the proportion of people with private health
+These three tables cover the social, economic and demographic characteristics of each county in the U.S.
+By referring to my previous research, where I found the variables with the highest correlations to the COVID mortality rate, I will use the same variables in this paper.
+That said, I will extract the data regarding the educational attainment from DP02, especially the proportion of people having at least a bachelor's degree.
+Regarding the economy, I will take the proportion of people with private health insurance and the mean household income.
+Lastly, regarding the demographic factors, I will use the total population, the ratios of children and individuals above 85.
+I will also take the White and Black people percentage as well.
+
+## Data Cleaning
+
 ```{r}
@@
             inf = mean(infrate),
-            mort = mean(mortrate))
+            mort = mean(mortrate),
+            income = mean(mean_household_income)) %>% 
+  rename(`Winning Party` = winning_party,
+         `Cases` = case,
+         `Deaths` = deaths,
+         `Infection Rate` = inf,
+         `Mortality Rate` = mort,
+         `Income` = income) %>% 
+  kable()
 ```
 
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+Combining the three data sources, @tbl-covid shows the COVID cases and deaths and the infection and mortality rate for the county voting for the Democrat and Republican parties, respectively.
+Surprisingly, the average cases and fatalities of COVID-19 in the counties voting for the Democrats are significantly higher than that of the Republican party.
+However, the values of infection and mortality rate do not show the same dramatic difference.
+It also emphasizes that the counties (states) voting for the Democrat are population intensive, and counties with more residence generally have higher mean household income levels.
+We can observe this from the average income where the "Democratic" counties have about \$13k mean income more elevated than the "Republic" ones.
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| echo: false
+#| message: false
+#| label: fig-boxplot
+#| fig-cap: Summary of income levels for counties voting for Democratic and Republican
+
+data  %>%
+  ggplot(mapping = aes(x = winning_party, y = mean_household_income)) +
+  geom_boxplot() +
+  geom_jitter(alpha = 0.04, width = 0.3, height = 0) +
+  theme_minimal() +
+  labs(x = "Voting Party",
+       y = "Death Rate per 1000")
+```
+
+@fig-boxplot compares the distribution of county income levels for the two parties.
+In general, counties voting for the Democrats have higher income levels than the Republicans.
+In addition, we can observe that the counties are intensively concentrated around income levels \$50k to \$75k, compared to the Democrats, where the counties are approximately uniformly distributed at each income level.
+Furthermore, barely any county has at least 150k voting for the Republican party.
+Both @tbl-covid and @fig-boxplot indicates that the Democrat is in favour of wealthy counties but poorer counties for Republican.
+Consequently, the more affluent counties (states) usually have more electoral votes than the poorer ones; this may provide insights into why Trump lost.
+
+```{r}
+#| echo: false
+#| message: false
+
+get_state_abbreviation <- function(state_names) {
+  sapply(state_names, function(state_name) {
+    state_data <- state.abb[match(state_name, state.name)]
+    if (!is.na(state_data)) {
+      return(state_data)
+    } else {
+      return(NA)  # changed from NULL to NA for consistency in a vector
+    }
+  })
+}
+
+county <- map_data("county")
+state <- map_data("state")
+county_clean <- county %>% 
+  mutate(state = region,
+         county = subregion,
+         state = get_state_abbreviation(toTitleCase(state)),
+         county = toTitleCase(county)) %>% 
+  filter(region != "AK") %>% 
+  dplyr::select(-region, -subregion)
+
+merged_data <- county_clean %>% left_join(republican_data, by = c("state", "county"))
+```
+
+```{r}
+#| echo: false
+#| message: false
+#| fig.width: 10 
+#| fig.height: 10
+#| label: fig-map
+#| fig-cap: The ratio of votes for the Republican and the infection rate per 100k in each county
+
+p1 <- ggplot(data = merged_data, aes(x = long, y = lat, group = group)) + 
+  geom_polygon(aes(fill = ratio_rd), color = "black") +
+  scale_fill_distiller(palette = "RdBu", direction = -1, na.value = "grey") +
+  labs(fill = 'Share of Republican') +
+  coord_quickmap() +
+  theme_void() +
+  geom_path(data = state, aes(x = long, y = lat, group = group), color = "black", linewidth = 0.5)
+
+# Second plot
+p2 <- ggplot(data = merged_data, aes(x = long, y = lat, group = group)) + 
+  geom_polygon(aes(fill = infrate), color = "black") +
+  scale_fill_gradient(low = "white", high = "darkblue", na.value = "grey",
+                      limits = c(0, 5000), oob = scales::squish) +
+  labs(fill = 'Infection Rate per 100k') +
+  coord_quickmap() +
+  theme_void() +
+  geom_path(data = state, aes(x = long, y = lat, group = group), color = "black", linewidth = 0.5)
+
+# Arrange the plots in a 2x1 layout
+grid.arrange(p1, p2, ncol = 1)
+```
+
+## Data Limitation
+
+Although the merged data covers the voting patterns, socio-economic characteristics and COVID-19 infection rate by county, some counties are missing.
+From the above two maps, the counties in the middle west show grey, indicating that the data for these counties are absent.
+This absence of data may increase bias and influence the accuracy of data analyses.
+
+# Methods
+
+This paper aims to make a causal inference between the 2020 voting patterns and COVID-19 to find whether Trump can contribute to his loss of COVID-19.
+To see the causal effect in the observational study, I will implement the methods of propensity scores to find the change in votes for Trump due to the pandemic.
+Based on the results, I will also conduct a counterfactual analysis to see what would have happened if there had been no COVID-19,
+
+First and foremost, we need to define the treatment.
+The treatment in an experiment will be artificially imposed on specific groups of people to see the effect of interested variables with and without this treatment.
+Therefore, the people with the treatment are called the treatment group, and the rest are called the control group.
+The treatment and control group settings are to test whether the imposed factor impacts the interested variable.
+However, since this is an observational study, it is unrealistic to set a factor manually.
+Therefore, to find the causal effect of COVID-19 on Trump's voting, I will choose the treatment to be whether a county has a high infection rate.
+
+## Propensity Score
+
+Propensity score is a technique that can help us to decide which county would be in the treatment group.
+That said, the propensity score in this study is the probability of a county having a high infection rate, given a set of socioeconomic variables.
+Using the propensity score can ensure that the two counties with similar propensity scores have the same distribution of the observed socioeconomic variables, such as income level, even though they are in different groups.
+This is fantastic, as we can have many pairs from the treatment of the control group with similar covariates by fixing propensity scores.
+The key drawback is that we can not estimate the causal effect from observational studies.
+
+In addition, by finding the propensity scores, we can reconstruct our assignment mechanism to a strongly ignorable one.
+An assignment mechanism is "strongly ignorable" if and only if, given the observed covariates, the potential outcomes are independent of the treatment assignment.
+This indicates that the voting change for Trump is independent of our assignment that a county has a high or low infection rate.
+
+As the propensity score represents the probability of a county having a high infection rate, we must find a logistic regression to predict this probability.
+In my previous paper, I already found the best model predicting the mortality rate of COVID in each US county.
+Even though that paper was to predict the mortality rate, the mortality is correlated with the infection rate.
+Therefore, I will take the dependent variables in that model to be the covariates in predicting the propensity score for each county, which is:
+
+```{=tex}
+\begin{align*}
+\frac{PS}{1-PS} &=  \beta_0 + \\
+& + \beta_1 \times \text{prop\_higher\_education} \\
+& + \beta_2 \times \text{IncomePctile} \\
+& + \beta_3 \times \text{no\_insurance} \\
+& + \beta_4 \times \text{private\_insurance} \\
+& + \beta_5 \times \text{males} \\
+& + \beta_6 \times \text{old\_85} \\
+& + \beta_7 \times \text{white\_pct} \\
+& + \beta_8 \times \text{black\_pct}
+\end{align*}
+```
+
+Therefore, we can find the PS for each county using the above model. However, there are three ways to use the propensity scores to find the treatment effect.
+
+### Propensity Score Matching
+The package needed to implement the matching is the `Matching.` Using this package, we can match the observations in the treatment and control groups having the same or similar propensity scores.
+Then, take only the matched pairs to be the new data set and find the treatment effect between them.
+In addition, we can also check the balance between the treatment and control groups.
+
+However, the most significant drawback is that the new data set can only contain the matched pairs.
+We may lose a considerable amount of observations during this process.
+Our predictions may also not be accurate.
+
+### Propensity Score Stratification
+As introduced above, using PSM may cause the waste of data and results in bias.
+However, propensity score stratification will use the entire data set to avoid waste.
+The way to do this is to split the data into different quantiles and fit a model for each subclass, with the only dependent variable being the treatment variable.
+Then, calculate the average casual effect for the subclasses.
+
+Even though the PSS can use the entire data set, however, the choice of quantiles will be closely related to the predictions.
+If we have a large sample and want to split the data into many subclasses, then this method may be computationally slow.
+
+### Propensity Score Regression Adjustment
+The last method is the most straightforward, meaning we just fit a regression model with the treatment variable and propensity score as the two dependent variables.
+
+Each of the above three methods has its unique advantages, but the results may vary between different data sets. I will perform both of the three methods and compare which one would be the best.
+
+## Counterfactual Analysis
+Using the propensity score methodology, we can assess the causal effect of a high infection rate on the voting for Trump.
+However, it can not tell us the election results, and we can not draw conclusions about COVID and the election.
+Therefore, I will conduct a counterfactual analysis, meaning that I will re-calculate the votes for Trump and Biden to see if the final election results will change.
+
+Eventually, I will re-calculate the votes for the two candidates to follow the official election rule.
+By finding the difference in voting percentage, I will re-allocate the voting for the two parties and calculate the electoral votes in each state using the winner-takes-all rule.
+
+# Results
+
+```{r}
 #| echo: false
@@
 
-# analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
-# 
-# analysis_data |> 
-#   ggplot(aes(x = width, y = length)) +
-#   geom_point(alpha = 0.8) +
-#   theme_minimal() +
-#   labs(x = "Wing width (mm)",
-#        y = "Wing length (mm)")
+highinf_model <- glm(treatment ~ prop_higher_education + pctile + 
+                       no_insurance + private_insurance + males +
+                       white_pct + black_pct, family = binomial(),
+                      data = republican_data)
+# summary(highinf_model)
+republican_data$prop_score <- predict(highinf_model, type = "response")
 ```
 
-Talk way more about it.
+```{=tex}
+\begin{align*}
+\text{Logit}(\text{PS}) &= -2.008 + \\
+& -0.032 \times \text{prop\_higher\_education} \\
+& -0.009 \times \text{IncomePctile} \\
+& +0.088 \times \text{no\_insurance} \\
+& +0.051 \times \text{private\_insurance} \\
+& +0.019 \times \text{males} \\
+& +0.055 \times \text{old\_85} \\
+& -0.030 \times \text{white\_pct} \\
+& +0.026 \times \text{black\_pct}
+\end{align*}
+```
 
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
+The above equation shows the logistic regression model to predict the propensity score for each county, given the covariates.
+It seems that insurance is a critical factor in the infection rate.
+Interestingly, the counties with a higher proportion of people with no insurance and private insurance will both increase the probability of this county having a high infection rate.
+Using this model, we can find the propensity score for each county and implement the three methods.
 
@@
 #| echo: false
-#| eval: true
-#| warning: false
 #| message: false
 
-highinf_model <- glm(high_infrate ~ prop_higher_education + pctile + 
-                             no_insurance +private_insurance + males + age_85 + 
-                             white_pct + black_pct, data = republican_data)
-summary(highinf_model)
-# highmort_model <- glm(high_mortrate ~ prop_higher_education + pctile + 
-#                              no_insurance +private_insurance + males + age_85 + 
-#                              white_pct + black_pct, data = data, nboots = 10)
-# summary(highmort_model)
-republican_data$prop_score <- predict(highinf_model, type = "response")
+library(Matching)
+# Propensity score matching
+rr <- Match(Y = republican_data$change_vote, Tr = republican_data$treatment, 
+            X = republican_data$prop_score, M = 1)
+# summary(rr)
 ```
@@
 ```{r}
-library(Matching)
-# Propensity score matching
-# mb <- MatchBalance(high_infrate ~ prop_higher_education + pctile + 
-#                              no_insurance +private_insurance + males + age_85 + 
-#                              white_pct + black_pct, data = data)
-rr <- Match(Y = republican_data$pct_vote, Tr = republican_data$high_infrate, 
-            X = republican_data$prop_score, M = 1)
-# mb <- MatchBalance(high_infrate ~ prop_higher_education + pctile + 
-#                               no_insurance +private_insurance + males + age_85 + 
-#                               white_pct + black_pct, match.out = rr,
-#                    data = republican_data, nboots = 10)
-summary(rr)
+#| echo: false
+#| message: false
+#| label: tbl-PSS
+#| tbl-cap: The results of causal effect using propensity score statification
+# PSS
+sub <- quantile(republican_data$prop_score, probs = c(0.2, 0.4, 0.6, 0.8))
+
+sub1 <- glm(pct_vote ~ high_infrate, 
+            data = republican_data[republican_data$prop_score <= sub[1], ])
+sub2 <- glm(pct_vote ~ high_infrate, 
+            data = republican_data[republican_data$prop_score > sub[1] & republican_data$prop_score <= sub[2], ])
+
+sub3 <- glm(pct_vote ~ high_infrate, 
+            data = republican_data[republican_data$prop_score > sub[2] & republican_data$prop_score <= sub[3], ])
+
+sub4 <- glm(pct_vote ~ high_infrate, 
+            data = republican_data[republican_data$prop_score > sub[3] & republican_data$prop_score <= sub[4], ])
+
+sub5 <- glm(pct_vote ~ high_infrate, 
+            data = republican_data[republican_data$prop_score > sub[4], ])
+
+# Extracting the necessary information from each model
+extract_info <- function(model, quantile) {
+  coef_summary <- summary(model)$coefficients
+  data.frame(
+    quantile = quantile,
+    ce = coef_summary["high_infrate", "Estimate"],
+    p_val = coef_summary["high_infrate", "Pr(>|t|)"]
+  )
+}
+
+# Creating data frames for each model
+df1 <- extract_info(sub1, "0-20%")
+df2 <- extract_info(sub2, "20-40%")
+df3 <- extract_info(sub3, "40-60%")
+df4 <- extract_info(sub4, "60-80%")
+df5 <- extract_info(sub5, "80-100%")
+
+# Combining all data frames into one
+rbind(df1, df2, df3, df4, df5) %>% 
+  rename(`PS Quantile` = quantile,
+         `Estimated CE` = ce,
+         `P Value` = p_val) %>% 
+  kable()
 ```
@@
 ```{r}
+#| echo: false
+#| message: false
 # Propensity score regression adjustment
-model_adj <- glm(pct_vote ~ high_infrate + prop_score, data = republican_data)
-summary(model_adj)
+model_adj <- glm(pct_vote ~ treatment + prop_score, data = republican_data)
+# summary(model_adj)
 ```
@@
 ```{r}
+#| echo: false
+#| message: false
 # Counterfactual Analysis
 adjusted_data <- data %>% 
-  dplyr::select(state, county.x, fips, party, votes, total_votes, pct_vote, high_infrate)
+  dplyr::select(state, county, fips, party, votes, total_votes, pct_vote, high_infrate)
 
@@
 ```{r}
+#| echo: false
+#| message: false
 adjust_function2 <- function(data) {

### 2023-12-01 scripts/01-download_data.R
Update the scripts

@@ -22,17 +22,17 @@ url_covid <- paste0(
   "https://raw.githubusercontent.com/CSSEGISandData/",
   "COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/",
   "11-03-2020.csv")
-covid_data <- read.csv(url_covid, header = TRUE, sep = ",")
+
 
 url_covid2 <- paste0(
   "https://raw.githubusercontent.com/CSSEGISandData/",
   "COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/",
-  "01-01-2021.csv")
-covid_data2 <- read.csv(url_covid2, header = TRUE, sep = ",")
+  "01-20-2021.csv")
+
+covid_data <- read.csv(url_covid, header = TRUE, sep = ",")
 
 #### Save data ####
 write.csv(DP02, "inputs/data/DP_02.csv", row.names = FALSE)
 write.csv(DP03, "inputs/data/DP_03.csv", row.names = FALSE)
 write.csv(DP05, "inputs/data/DP_05.csv", row.names = FALSE)
 write.csv(covid_data, "inputs/data/covid_data.csv", row.names = FALSE)
-write.csv(covid_data2, "inputs/data/covid_data2.csv", row.names = FALSE)

### 2023-12-01 scripts/02-data_cleaning.R

@@ -20,7 +20,6 @@ DP02 <- read.csv("inputs/data/DP_02.csv")
 DP03 <- read.csv("inputs/data/DP_03.csv")
 DP05 <- read.csv("inputs/data/DP_05.csv")
 covid_data <- read.csv("inputs/data/covid_data.csv")
-covid_data2 <- read.csv("inputs/data/covid_data2.csv")
 election_data <- read.csv("inputs/data/countypres_2000-2020.csv")
 
 get_state_abbreviation <- function(state_names) {
@@ -101,34 +100,66 @@ covid_data_clean <- covid_data %>%
   clean_names() 
 covid_data_clean <- na.omit(covid_data_clean)
 
-covid_data2_clean <- covid_data2 %>% 
-  filter(Country_Region == "US") %>% 
-  dplyr::select(FIPS, Admin2, Province_State, Confirmed, Deaths) %>% 
-  rename(`county` = Admin2,
-         `state` = Province_State,
-         `cases` = Confirmed) %>%
-  mutate(state = get_state_abbreviation(state)) %>% 
-  clean_names() 
-covid_data2_clean <- na.omit(covid_data2_clean)
-
 # Clean the election data
-election_data_clean <- election_data %>% 
+election_data_2020 <- election_data %>% 
   filter(year == 2020) %>% 
   mutate(county = str_to_title(county_name),
          party = str_to_title(party),
          state = state_po,
          votes = candidatevotes,
          total_votes = totalvotes, 
-         fips = county_fips,
-         pct_vote = votes/total_votes) %>%
-  dplyr::select(state_po, county, fips, candidate, party, votes, total_votes, pct_vote)
+         fips = county_fips) %>% 
+  dplyr::select(state, county, fips, candidate, party, votes, total_votes, mode)
 
-winning_party <- election_data_clean %>%
+total_data <- election_data_2020 %>% 
+  filter(mode == "TOTAL") %>% 
+  dplyr::select(-mode)
+non_total_data <- election_data_2020 %>% 
+  filter(mode != "TOTAL" & state != "UT") %>% 
+  dplyr::select(-mode)
+
+aggregated_data <- non_total_data %>%
+  group_by(state, county, fips, candidate, party, total_votes) %>% 
+  summarize(votes = sum(votes, na.rm = TRUE), .groups = 'drop')
+
+election_data_2020 <- bind_rows(total_data, aggregated_data) %>% 
+  mutate(pct_vote = votes/total_votes)
+
+winning_party <- election_data_2020 %>%
[…]

### 2023-12-08 outputs/paper/paper.qmd
Update the output paper

@@
 ---
-title: "Draft"
+title: "Causal Inference in the 2020 US Federal Election: While COVID-19 Lowered Trump's Support, Loss of Confidence Among Wealthier Voters Seems More Decisive"
 title-block-banner: true
@@
 date-format: "long"
-abstract: "TBD"
+abstract: "This study used the election data from MIT EDSL and COVID-19 data from JHU CSSE, to investigate the causal inference between COVID-19 and Donald Trump's loss during the 2020 US Federal Election.
+The main methodology used in this paper is Propensity Score Matching with testing different treatments.
+The primary finding is that the counties with high death per case rate seems to vote less for Trump with approximate treatment effect -0.01.
+Besides, it seems that the interaction effect with income seems to be more critical: the medium-high income levels counties seems to vote at maximum 3% less for Trump compared to those low income counties.
+Furthermore, this paper also conduct counterfactual analysis and indicate the Trump would pre-elect if there was no COVID.
+Furture analysis should focus more on the choice of covariates to predict propensity scores."
 format: pdf
@@
 data <- read.csv(here("outputs/data/merged_data.csv"))
-election_2016_republican <- read.csv(here("outputs/data/election_data_2016.csv")) %>% 
-  filter(party == "Republican") %>% 
-  dplyr::select(fips, pct_vote_2016)
-
-republican_data <- data %>% 
-  filter(party == "Republican") %>% 
-  inner_join(election_2016_republican, by = "fips") %>% 
-  mutate(change_vote = pct_vote - pct_vote_2016)
 ```
@@
 #| tbl-cap: The summary of voting patterns in 2020 US Federal Presidential Election
-election_data %>% 
-  group_by(party, candidate) %>% 
-  summarise(total = sum(votes),
-            mean = round(mean(pct_vote), 2),
-            median = round(median(pct_vote), 2)) %>%
-  arrange(desc(total)) %>% 
-  rename(`Party` = party,
-         `Candidate` = candidate,
-         `Total Votes` = total,
-         `Mean Pct Vote` = mean,
-         `Median Pct Vote` = median) %>% 
-  kable()
+# election_data %>% 
+#   group_by(party, candidate) %>% 
+#   summarise(total = sum(votes),
+#             mean = round(mean(pct_vote), 2),
+#             median = round(median(pct_vote), 2)) %>%
+#   arrange(desc(total)) %>% 
+#   rename(`Party` = party,
+#          `Candidate` = candidate,
+#          `Total Votes` = total,
+#          `Mean Pct Vote` = mean,
+#          `Median Pct Vote` = median) %>% 
+#   kable()
 ```
@@
             inf = mean(infrate),
-            mort = mean(mortrate),
+            dpc = mean(dpc),
             income = mean(mean_household_income)) %>% 
@@
          `Infection Rate` = inf,
-         `Mortality Rate` = mort,
          `Income` = income) %>% 
@@
   theme_minimal() +
-  labs(x = "Voting Party",
-       y = "Death Rate per 1000")
+  labs(x = "Winning Party",
+       y = "Mean Household Income")
 ```
@@
          county = toTitleCase(county)) %>% 
-  filter(region != "AK") %>% 
+  filter(region != "AK" & region != "HI") %>% 
   dplyr::select(-region, -subregion)
 
-merged_data <- county_clean %>% left_join(republican_data, by = c("state", "county"))
+merged_data <- county_clean %>% 
+  left_join(data, by = c("state", "county"))
+```
+
+```{r}
+binned_data <- data %>%
+  mutate(dpc_pctile = ntile(dpc, 100)) %>%  # Adjust 'breaks' to change number of bins
+  group_by(dpc_pctile) %>%
+  summarize(mean_rep = mean(pct_vote_rep, na.rm = TRUE),
+            mean_demo = mean(pct_vote_demo, na.rm = TRUE)) %>% 
+  mutate(bin = 1:n())
+
+binned_data2 <- data %>%
+  mutate(dpc_pctile = ntile(mean_household_income, 100)) %>%  # Adjust 'breaks' to change number of bins
+  group_by(dpc_pctile) %>%
+  summarize(mean_rep = mean(pct_vote_rep, na.rm = TRUE),
+            mean_demo = mean(pct_vote_demo, na.rm = TRUE)) %>% 
+  mutate(bin = 1:n(),
+         sum = mean_rep + mean_demo)
+
+# Plotting
+ggplot(binned_data, aes(x = bin, y = mean_rep)) +
+  geom_point() +  # or geom_line() depending on your preference
+  geom_smooth(method = "lm",formula = y ~ poly(x, 4), se=FALSE) +
+  theme_minimal() +
+  labs(x = "Binned mortality per case ", y = "Mean vote for Trump",
+       title = "Votes for Trump vs. MPC")
+ggplot(binned_data, aes(x = bin, y = mean_demo)) +
+  geom_point() +  # or geom_line() depending on your preference
+  geom_smooth(method = "lm", formula = y ~ poly(x, 4),se=FALSE) +
+  theme_minimal() +
+  labs(x = "Binned mortality per case ", y = "Mean vote for Biden", 
+       title = "Votes for Biden vs. MPC")
+
+ggplot(binned_data2, aes(x = bin, y = mean_rep)) +
+  geom_point() +  # or geom_line() depending on your preference
+  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se=FALSE) +
+  theme_minimal() +
+  labs(x = "Binned income levels", y = "Mean vote for Trump", 
+       title = "Votes for Trump vs. Income")
+ggplot(binned_data2, aes(x = bin, y = mean_demo)) +
+  geom_point() +  # or geom_line() depending on your preference
+  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se=FALSE) +
+  theme_minimal() +
+  labs(x = "Binned income levels", y = "Mean vote for Biden",
+       title = "Votes for Biden vs. Income")
 ```
@@
 p1 <- ggplot(data = merged_data, aes(x = long, y = lat, group = group)) + 
-  geom_polygon(aes(fill = ratio_rd), color = "black") +
+  geom_polygon(aes(fill = pct_vote_rep), color = "black") +
   scale_fill_distiller(palette = "RdBu", direction = -1, na.value = "grey") +
-  labs(fill = 'Share of Republican') +
+  labs(title = "The relative share of votes between Republican and Democrat party",
+       fill = 'Share of Republican') +
   coord_quickmap() +
@@
 p2 <- ggplot(data = merged_data, aes(x = long, y = lat, group = group)) + 
-  geom_polygon(aes(fill = infrate), color = "black") +
+  geom_polygon(aes(fill = dpc), color = "black") +
   scale_fill_gradient(low = "white", high = "darkblue", na.value = "grey",
                       limits = c(0, 5000), oob = scales::squish) +
-  labs(fill = 'Infection Rate per 100k') +
+  labs(title = "The distribution of death per case rate by 100K",
+       fill = 'Death per cases by 100k') +
+  coord_quickmap() +
+  theme_void() +
+  geom_path(data = state, aes(x = long, y = lat, group = group), color = "black", linewidth = 0.5)
+
+p3 <- ggplot(data = merged_data, aes(x = long, y = lat, group = group)) + 
+  geom_polygon(aes(fill = mean_household_income), color = "black") +
+  scale_fill_gradient(low = "white", high = "darkblue", na.value = "grey",
+                      limits = c(60000, 120000), oob = scales::squish) +
+  labs(title = "The distribution of mean household income",
+       fill = 'Mean Hosehold Income') +
   coord_quickmap() +
@@
 # Arrange the plots in a 2x1 layout
-grid.arrange(p1, p2, ncol = 1)
+grid.arrange(p1, p3, p2, ncol = 1)
+```
+
+```{r}
+#| eval: false
+library(sp)
+library(geojsonio)
+library(ggplot2)
+library(dplyr)
+library(broom)
+hex_county <- geojson_read("us_county_hexgrid.geojson", what = "sp")
+
+# Reformat the data as needed (depending on your data structure)
+hex_county@data <- hex_county@data %>% 
+  mutate(adjusted_field = gsub("some_pattern", "replacement", original_field))
+
+# Fortify the data for ggplot
+hex_county_fortified <- tidy(hex_county, region = "desired_region_field")
+
+# Plot the hexagon map
+ggplot() +
+  geom_polygon(data = hex_county_fortified, aes(x = long, y = lat, group = group), fill="#a1dab4", color="#f7f7f7") +
+  geom_text(data = hex_county_fortified, aes(x = long, y = lat, label = label_field), size = 3) +
+  theme_void() +
+  coord_map()
 ```
@@
 #| message: false
+data <- read.csv(here("outputs/data/merged_data.csv"))
 
@@
                        white_pct + black_pct, family = binomial(),
-                      data = republican_data)
+                      data = data)
 # summary(highinf_model)
-republican_data$prop_score <- predict(highinf_model, type = "response")
+data$prop_score <- predict(highinf_model, type = "response")
 ```
@@
 Using this model, we can find the propensity score for each county and implement the three methods.
+```{r}
+model <- lm(pct_vote_rep ~ dpc + mean_household_income, data = data)
+summary(model)
+```
 
@@
 # Propensity score matching
-rr <- Match(Y = republican_data$change_vote, Tr = republican_data$treatment, 
-            X = republican_data$prop_score, M = 1)
-# summary(rr)
+rr <- Match(Y = data$pct_vote_rep, Tr = data$treatment, 
+            X = data$prop_score, M = 1)
+summary(rr)
 ```
@@
 #| tbl-cap: The results of causal effect using propensity score statification
+#| eval: false
 # PSS
-sub <- quantile(republican_data$prop_score, probs = c(0.2, 0.4, 0.6, 0.8))
+sub <- quantile(data$prop_score, probs = c(0.2, 0.4, 0.6, 0.8))
 
-sub1 <- glm(pct_vote ~ high_infrate, 
-            data = republican_data[republican_data$prop_score <= sub[1], ])
-sub2 <- glm(pct_vote ~ high_infrate, 
-            data = republican_data[republican_data$prop_score > sub[1] & republican_data$prop_score <= sub[2], ])
+sub1 <- glm(pct_vote_rep ~ treatment,
+            data = data[data$prop_score <= sub[1], ])
+sub2 <- glm(pct_vote_rep ~ treatment,
+            data = data[data$prop_score > sub[1] & data$prop_score <= sub[2], ])
 
-sub3 <- glm(pct_vote ~ high_infrate, 
-            data = republican_data[republican_data$prop_score > sub[2] & republican_data$prop_score <= sub[3], ])
+sub3 <- glm(pct_vote_rep ~ treatment,
+            data = data[data$prop_score > sub[2] & data$prop_score <= sub[3], ])
 
-sub4 <- glm(pct_vote ~ high_infrate, 
-            data = republican_data[republican_data$prop_score > sub[3] & republican_data$prop_score <= sub[4], ])
+sub4 <- glm(pct_vote_rep ~ treatment,
+            data = data[data$prop_score > sub[3] & data$prop_score <= sub[4], ])
 
-sub5 <- glm(pct_vote ~ high_infrate, 
-            data = republican_data[republican_data$prop_score > sub[4], ])
+sub5 <- glm(pct_vote_rep ~ treatment,
+            data = data[data$prop_score > sub[4], ])
 
@@
   coef_summary <- summary(model)$coefficients
-  data.frame(
+  data <- data.frame(
     quantile = quantile,
-    ce = coef_summary["high_infrate", "Estimate"],
-    p_val = coef_summary["high_infrate", "Pr(>|t|)"]
+    ce = coef_summary["treatment", "Estimate"],
+    p_val = coef_summary["treatment", "Pr(>|t|)"]
   )
+  return(data)
 }
@@
 # Propensity score regression adjustment
-model_adj <- glm(pct_vote ~ treatment + prop_score, data = republican_data)
-# summary(model_adj)
+model_adj <- glm(change_vote_rep ~ treatment + prop_score, data = data)
+summary(model_adj)
 ```
@@
 #| message: false
+#| eval: false
 # Counterfactual Analysis
@@
 #| message: false
+#| eval: false
 adjust_function2 <- function(data) {

### 2023-12-08 scripts/02-data_cleaning.R
Update the data cleaning procedure

@@ -13,6 +13,7 @@ library(stringr)
 library(tidyverse)
 library(janitor)
 library(tools)
+library(tidyr)
 
 #### Clean data ####
 # Import all the raw data
@@ -122,33 +123,31 @@ aggregated_data <- non_total_data %>%
   group_by(state, county, fips, candidate, party, total_votes) %>% 
   summarize(votes = sum(votes, na.rm = TRUE), .groups = 'drop')
 
-election_data_2020 <- bind_rows(total_data, aggregated_data) %>% 
-  mutate(pct_vote = votes/total_votes)
+election_data_2020 <- bind_rows(total_data, aggregated_data)
 
-winning_party <- election_data_2020 %>%
-  group_by(fips) %>%
-  summarize(winning_party = party[which.max(votes)])
+election_data_2020_clean <- election_data_2020 %>%
+  group_by(state, county, fips, party, total_votes) %>%
+  summarise(votes = sum(votes), .groups = 'drop') %>%
+  pivot_wider(names_from = party, values_from = votes) %>% 
+  filter(state != "AK") %>% 
+  mutate(vote_demo = Democrat,
+         vote_rep = Republican,
+         vote_green = Green,
+         vote_lib = Libertarian,
+         vote_other = Other,
+         across(c(vote_demo, vote_rep, vote_green, vote_lib, vote_other), 
+                ~replace(., is.na(.), 0)),
+         # winning_party = ifelse(vote_demo > vote_rep, "Democrat", "Republican"),
+         pct_vote_demo = vote_demo/total_votes,
+         pct_vote_rep = as.numeric(vote_rep/total_votes)) %>% 
+  dplyr::select(-Democrat, -Green, -Libertarian, -Other, -Republican)
 
-election_data_2020 <- election_data_2020 %>%
-  left_join(winning_party, by = "fips")
-
-total_votes_rd <- election_data_2020 %>%
-  filter(party %in% c("Democrat", "Republican")) %>%
-  group_by(fips) %>%
-  summarize(total_votes_rd = sum(votes),
-            .groups = 'drop')
-
-votes_r <- election_data_2020 %>%
-  filter(party == "Republican") %>%
-  group_by(fips) %>%
-  summarize(votes_r = sum(votes),
-            .groups = 'drop')
-
-election_data_clean <- election_data_2020 %>%
-  left_join(total_votes_rd, by = "fips") %>%
-  left_join(votes_r, by = "fips") %>% 
-  mutate(ratio_rd = votes_r/total_votes_rd) %>% 
[…]

### 2023-12-11 outputs/paper/paper.qmd
Try MatchIt

@@
 
+```{r}
+data <- read.csv(here("outputs/data/merged_data.csv"))
+
+library(MatchIt)
+m.out <- matchit(treatment ~ prop_higher_education + pctile + 
+                       no_insurance + private_insurance + males +
+                       white_pct + black_pct,
+                 data = data,
+                 distance = "logit",
+                 method = "nearest",
+                 replace = FALSE,
+                 ratio = 1)
+summary(m.out, standardize = TRUE)
+```
+
 ```{=tex}
@@
 summary(rr)
+```
+
+```{r}
+
 ```

### 2023-12-11 outputs/paper/paper.qmd
Update the data part

@@
 # Data
+
+The data used in this paper comprised five data sets from three different sources corresponding to different topics.
+The primary source is MIT Election Data Science Club which build open online data collections of the US Federal or Senate Election results, spanning from nation to county levels.
+The data extracted is called "County Presidential Election Returns 2000-2020" with about 70,000 rows containing the voting patterns for each candidate and party by county since 2000.
+Besides that, the data also indicate the types of voting, such as "EARLY VOTE" and "ELECTION DAY," for the same party in the a county.
+To analyze the voting patterns for Donald J.
+Trump, I only select the data for 2016 and 2020 US Federal Election for all counties and parties.
+Thus the filterd data set contains the number of votes through various ways for each county and party.
+
+In addition to that, the study also uses the data regarding COVID-19 from the Center for Systems Science and Engineering at Johns Hopkins University
 

### 2023-12-15 scripts/02-data_cleaning.R
Update data cleaning and cleaned data sets

@@ -25,11 +25,15 @@ election_data <- read.csv("inputs/data/countypres_2000-2020.csv")
 
 get_state_abbreviation <- function(state_names) {
   sapply(state_names, function(state_name) {
-    state_data <- state.abb[match(state_name, state.name)]
-    if (!is.na(state_data)) {
-      return(state_data)
+    if (state_name == "District of Columbia") {
+      return("D.C.")
     } else {
-      return(NA)  # changed from NULL to NA for consistency in a vector
+      state_data <- state.abb[match(state_name, state.name)]
+      if (!is.na(state_data)) {
+        return(state_data)
+      } else {
+        return(NA)  # Keep NA for consistency in vector
+      }
     }
   })
 }
@@ -97,9 +101,12 @@ covid_data_clean <- covid_data %>%
   rename(`county` = Admin2,
          `state` = Province_State,
          `cases` = Confirmed) %>%
-  mutate(state = get_state_abbreviation(state)) %>% 
+  mutate(state = get_state_abbreviation(state),
+         state = ifelse(FIPS == 11001, "DC", state)) %>% 
   clean_names() 
-covid_data_clean <- na.omit(covid_data_clean)
+
+na_rows <- apply(covid_data_clean, 1, function(x) any(is.na(x)))
+covid_data_clean[na_rows, ]
 
 # Clean the election data
 election_data_2020 <- election_data %>% 
@@ -109,7 +116,7 @@ election_data_2020 <- election_data %>%
          state = state_po,
          votes = candidatevotes,
          total_votes = totalvotes, 
-         fips = county_fips) %>% 
+         fips = ifelse(state == "DC", 11001, county_fips)) %>% 
   dplyr::select(state, county, fips, candidate, party, votes, total_votes, mode)
 
 total_data <- election_data_2020 %>% 
@@ -129,7 +136,6 @@ election_data_2020_clean <- election_data_2020 %>%
   group_by(state, county, fips, party, total_votes) %>%
   summarise(votes = sum(votes), .groups = 'drop') %>%
   pivot_wider(names_from = party, values_from = votes) %>% 
-  filter(state != "AK") %>% 
   mutate(vote_demo = Democrat,
          vote_rep = Republican,
          vote_green = Green,
[…]

### 2023-12-15 outputs/paper/paper.qmd
Update the introduction (First Draft)

@@
 ---
-title: "Causal Inference in the 2020 US Federal Election: While COVID-19 Lowered Trump's Support, Loss of Confidence Among Wealthier Voters Seems More Decisive"
+title: "An Observational Study to COVID-19 and 2020 US Presidential Election: Counties with High Deaths per Case Rate Showed Diminished Support For Trump"
 title-block-banner: true
@@
 date-format: "long"
-abstract: "This study used the election data from MIT EDSL and COVID-19 data from JHU CSSE, to investigate the causal inference between COVID-19 and Donald Trump's loss during the 2020 US Federal Election.
-The main methodology used in this paper is Propensity Score Matching with testing different treatments.
-The primary finding is that the counties with high death per case rate seems to vote less for Trump with approximate treatment effect -0.01.
-Besides, it seems that the interaction effect with income seems to be more critical: the medium-high income levels counties seems to vote at maximum 3% less for Trump compared to those low income counties.
-Furthermore, this paper also conduct counterfactual analysis and indicate the Trump would pre-elect if there was no COVID.
-Furture analysis should focus more on the choice of covariates to predict propensity scores."
+abstract: "This study investigate the causal inference between COVID-19 and Donald Trump's loss during the 2020 US Federal Election using the data from MIT Election Data Science Lab and Johns Hopkins University CSSE.
+The main methodology used in this paper is Propensity Score Matching with with an exploratory analyzing the optimal treatment initially.
+The key finding suggest that the counties with a death per case rate exceeding the 0.4 quantile threshold shows a reduced voting preference for Trump.
+Additionally, this paper conducts a counterfactual analysis based on the treatment effect, indicating that Trump might have be secured to re-elect if the disparities in death per case rates were addressed.
+Future research should aim to refine the propensity score calculations by incorporating additional variables, such as the winning party in each county during the 2016 election."
 format: pdf
@@
 #| message: false
-
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
 library(tidyverse)
@@
 library(tools)
+library(dplyr)
+library(ggplot2)
 ```
@@
 
+At the beginning of 2020, US reported the first COVID-19 cases.
+This number grows to ten thousand within two months on March 19 and increase dramatically to 100K on March 27 [citation].
+At the same year, the U.S.
+Federal Presidential Election started off and Trump lost to Biden.
+As of Election Day, there was above 9 million cases and about 0.2 million deaths in US.
+When people talk about Trump's lost, people always connect it to the pandemic, saying that he "mishandled his greatest test" [citation] and 2016 Election was a historical accident [citation].
+However, there is another voice that COVID is not the only or the decisive factor attributing to his lost, his limits as a politican and wasted his advantage on the economy are also important explaining why he lost.
+This paper will investigate whether the is a causal effect between COVID and Trump's lost.
+If there is, whether it is strong enough to beath Trump at 2020.
+
+There are existing researches on this.
+Leaonardo et al.
+[citation] suggested that Trump could win if there was no COVID but it will be too naive to simply connect his lost to the COVID-19 infection rate.
+Their research highlighted that the COVID-19 had the most significant negative impacts on those urban areas where there was no stay-at-home orders, especially for those "swing" states.
+They also argue that the race diversity and education attainment are also important when analysing voting patterns for Trump.
+In addition to that, Marcus et al.
+[citation] suggested that the deaths per case is a more crucial indicator than infection rate when analysing the voting for Trump.
+However, they suggest the voting itself may not capture the impact of COVID as we do not know when did the voters make their decision, especially for early vote.
+Compared to these researches, Harold [citation] conduct read survey before and after the election to illustrate the impact of COVID pattern, they found that even though COVID has impacts on voting, it is not the dominant factor.
+Without COVID, Trump was also less likely to re-elect due to US highly polarized political landscape.
+Similarly, Shang et al.
+[citation] also conduct a survey and all the participants show natural or negative altitude.
+They suggest some social activity like "Black lives matter" may have impacts on voting.
+Furthermore, the
+
+While all the above researches had investigated the correlation between COVID-19 and votings for Trump.
+Except conducting real surveys, the methods they used were to conduct a model and find the relevant variables in predicting the (change) vote for Trump.
+Then conduct counterfactual analysis to predict what will be the voting for Trump if there was no COVID or a certain amount reduce in COVID deaths.
+Instead using a similar way like those studies, this paper will introduce the method of Propensity Score Matching to find the treatment effect, to see whether those more impacted counties will have different voting patterns for Trump.
+However, since all counties experienced COVID-19, this paper will firstly find the optimal treatment groups.
+After that, I will also conduct a counterfactual analysis to see if Trump could re-elect.
+
+There will be five parts in this paper.
+I will introduce the data used in this paper and present data suammries and visualization in [Data] part.
+After that, I will introduce the methods in this paper and the corresponding results in  [Methods] and [Results] part, respectively.
+Then the results will be interpreted and discussed in the [Discussion] part and conclude with limitations and drawbacks in [Conclusion] part..
+
 # Data
+
+## Data sources
 
@@
 
-In addition to that, the study also uses the data regarding COVID-19 from the Center for Systems Science and Engineering at Johns Hopkins University
+In addition to that, the study also uses the data regarding COVID-19 from the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University.
+CSSE collects and reports the local, national, and global multidimensional data, including medicine, health care, disaster response, etc.
+During the pandemic, they collected the U.S.
+and international COVID cases and deaths and reported them on their GitHub, summarized by daily reports from April 12, 2020, to March 9, 2023.
+To access the impact of COVID on 2020 Election, this paper used the daily report on November 3, 2020 which is the election day of 2020 US Federal Election.
+The resulting data include the aggregates number of cases, deaths and recovers, as well as the incidence rate and case fatility ratio for each county by the election day.
 
-In this paper, the data used mainly consists of the voting patterns for the 2020 US Federal Election, COVID infection rate and socioeconomic variables in each county of the US.
-The voting data is taken from the MIT Election Data Science Lab, including voting for each party at the county level at each Federal Election from 2000 to 2020.
-The COVID data is from the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University.
-The socioeconomic data is taken from the American Community Survey.
+Lastly, this paper also employ the socio-economic data from American Community Survey (ACS).
+ACS is a online open source database conducted by the U.S.
+Census Bureau, containing the various soci-economic factors in either geographic levels.
+The data sets extracted from ACS is 2020 five-year estimates of DP02, DP03 and DP05 which covers the social, economic and demographic characteristics in county level.
+Since there are thousands of variables, I will only use the variables which are found from my previous paper that are significant in predicting the COVID-19 mortality rate such as the mean household income, high-education attainment etc.
+The descriptions of all variables can be found on Table 1.
 
-## MIT Election Data Science Lab
-
-The MIT Election Data Science Lab is a lab at MIT that collects and analyzes election data to support advances in election science.
-Their collected data is not limited to the presidential election results but also covers the midterm or US Senate results at the state or county level.
-
-The data abstracted from MITEDSL in this paper is the "County Presidential Election Returns 2000-2020." It contains the number of votes for each party and the total votes at each US county from 2000 to the 2020 US Federal Election.
-In this paper, I will only use the data for the 2020 US Election and focus on the counties on the mainland of the US.
-The @tbl-election summarizes the essential variables.
+After having five data sets from three sources, they are merged into one big data by counties.
+Using the merged data, there are some new variables calculated by the exiting variables.
+In terms of election, I calculate the percentage of vote for each party (candidate) in both 2016 and 2020 Election, and calculate the corresponding change of percentage votes.
+I also create new variables indicating the winning party in both elections for each county.
+Additionally, to accurately compare the COVID cases and deaths across all counties, I transform the number of cases and deaths to infection and mortality rate by 10,000 citizens in each county.
+The final data now consists of 3107 rows with 36 columns.
+All the important variables are described in Table 1.
 
 ```{r}
-#| echo: false
-#| message: false
-#| label: tbl-election
-#| tbl-cap: The summary of voting patterns in 2020 US Federal Presidential Election
-# election_data %>% 
-#   group_by(party, candidate) %>% 
-#   summarise(total = sum(votes),
-#             mean = round(mean(pct_vote), 2),
-#             median = round(median(pct_vote), 2)) %>%
-#   arrange(desc(total)) %>% 
-#   rename(`Party` = party,
-#          `Candidate` = candidate,
-#          `Total Votes` = total,
-#          `Mean Pct Vote` = mean,
-#          `Median Pct Vote` = median) %>% 
-#   kable()
+#| tbl-cap: Descriptions of all important variables in the analyze data
+#| label: tbl-descriptions
+data.frame(var = c("Income Percentile", "High-Education Attainment",
+                   "Private Insurance", "No Insurance", "White Population",
+                   "Black Population", "Males Population", "Infection Rate",
+                   "Mortality Rate", "Death per Case", "Votes for Democrat (2016)",
+                   "Votes for Democrat (2020)", "Votes for Republican (2016)",
+                   "Votes for Republican (2020)", "Change of Vote for Republican"),
+           code = c("income_pctile", "prop_high_education", "private_insurance",
+           "no_insurance", "white_pct", "black_pct", "males", "infrate",
+           "mortrate", "dpc", "pct_vote_demo16", "pct_vote_demo", 
+           "pct_vote_rep16", "pct_vote_rep", "change_vote_rep"),
+           descriptions = c("The income percentile of each county",
+                            "The proportion of residents in a county having a at least bachelor degree",
+                            "The proportion of local residences having private insurance",
+                            "The proportion of local residences without any health insurance",
+                            "The proportion of White population",
+                            "The proportion of Black population",
+                            "The proportion of Males population",
+                            "The COVID infection rate, calculated by number of case per 10,000 residences",
+                            "The COVID mortality rate, calculated by number of death per 10,000 residences",
+                            "The COVID death per 10,000 confirmed cases",
+                            "The percentage of votes for the Democrat in 2016",
+                            "The percentage of votes for the Democrat in 2020",
+                            "The percentage of votes for the Republican in 2016",
+                            "The percentage of votes for the Republican in 2020",
+                            "The change of vote for the Republican from 2016 to 2020")) %>% 
+  rename(`Variables` = var,
+         `Coded name` = code,
+         `Descriptions` = descriptions) %>% 
+  kable(booktabs = TRUE)
 ```
 
-The @tbl-election summarizes the voting patterns for each party during the 2020 U.S.
-Presidential Election.
-Undoubtedly, the two most popular parties are the Democrats and the Republicans.
-However, even though Republican has almost double the mean percent vote than Democrat, the total votes for Republican is less than for Democrat.
-This pattern may indicate that more people living in high-population states, such as California, voted for Biden.
-This finding aligns with the fact that Biden defeated Trump in 2020.
-I will explain more later.
-
-## Center for Systems Science and Engineering at JHU
-
-The Center for Systems Science and Engineering at JHU is at the Department of Civils and Engineering, which collects local, national, and global multidimensional data, including medicine, health care, disaster response, etc.
-During the pandemic, they collected the U.S.
-and international COVID cases and deaths and reported them on their GitHub.
-Their data is summarized by daily reports ranging from April 12, 2020, to March 9, 2023.
-
-To illustrate the impact of COVID-19 on the 2020 Election to the greatest extent, the COVID data used in this paper will be the daily report on November 3, 2020, which is Election Day \[citation\].
-This report contains the aggregated cases and deaths for each country and counties in the U.S.
-
-## American Community Survey (ACS)
-
-The American Community Survey is the survey conducted by the U.S.
-Census Bureau.
-The survey contains a variety of socio-economic variables for each county.
-Although there are different data tables from ACS, I will use the 2020 five-year estimate of DP02, DP03, and DP05.
-
-These three tables cover the social, economic and demographic characteristics of each county in the U.S.
-By referring to my previous research, where I found the variables with the highest correlations to the COVID mortality rate, I will use the same variables in this paper.
-That said, I will extract the data regarding the educational attainment from DP02, especially the proportion of people having at least a bachelor's degree.
-Regarding the economy, I will take the proportion of people with private health insurance and the mean household income.
-Lastly, regarding the demographic factors, I will use the total population, the ratios of children and individuals above 85.
-I will also take the White and Black people percentage as well.
-
-## Data Cleaning
+## Data summaries and visualizations
 
@@
   group_by(winning_party) %>% 
-  summarise(case = mean(cases),
+  summarise(n = n(),
+            case = mean(cases),
+            inf = mean(infrate),
             deaths = mean(deaths),
-            inf = mean(infrate),
+            mort = mean(mortrate),
             dpc = mean(dpc),
@@
   rename(`Winning Party` = winning_party,
+         `Count` = n,
          `Cases` = case,
+         `Infection Rate` = inf,
          `Deaths` = deaths,
-         `Infection Rate` = inf,
+         `Mortality Rate` = mort,
+         `DPC` = dpc,
          `Income` = income) %>% 
-  kable()
+  kable(booktabs = TRUE)
 ```
 
-Combining the three data sources, @tbl-covid shows the COVID cases and deaths and the infection and mortality rate for the county voting for the Democrat and Republican parties, respectively.
-Surprisingly, the average cases and fatalities of COVID-19 in the counties voting for the Democrats are significantly higher than that of the Republican party.
-However, the values of infection and mortality rate do not show the same dramatic difference.
-It also emphasizes that the counties (states) voting for the Democrat are population intensive, and counties with more residence generally have higher mean household income levels.
-We can observe this from the average income where the "Democratic" counties have about \$13k mean income more elevated than the "Republic" ones.
-
-```{r}
-#| echo: false
-#| message: false
-#| label: fig-boxplot
-#| fig-cap: Summary of income levels for counties voting for Democratic and Republican
-
-data  %>%
-  ggplot(mapping = aes(x = winning_party, y = mean_household_income)) +
-  geom_boxplot() +
-  geom_jitter(alpha = 0.04, width = 0.3, height = 0) +
-  theme_minimal() +
-  labs(x = "Winning Party",
-       y = "Mean Household Income")
-```
-
-@fig-boxplot compares the distribution of county income levels for the two parties.
-In general, counties voting for the Democrats have higher income levels than the Republicans.
-In addition, we can observe that the counties are intensively concentrated around income levels \$50k to \$75k, compared to the Democrats, where the counties are approximately uniformly distributed at each income level.
-Furthermore, barely any county has at least 150k voting for the Republican party.
-Both @tbl-covid and @fig-boxplot indicates that the Democrat is in favour of wealthy counties but poorer counties for Republican.
-Consequently, the more affluent counties (states) usually have more electoral votes than the poorer ones; this may provide insights into why Trump lost.
+@tbl-covid summarize the COVID-19 impacts and income levels between the counties voting for Democrat and Republican.
+Even though the Republican won about five sixths counties, there is a stark contrast where counties supporting Biden had nearly ten times the average number of COVID-19 cases and deaths compared to those supporting Trump.
+Despite a similar infection and mortality rates between the two groups, there is still a notable higher death-to-case rate in counties voting for the Republican.
+Furthermore, these counties also has a lower mean household income than the "Democrat-win" counties.
+These patterns suggest a potential correlation between the extent of COVID-19 impact and voting behavior in the 2020, where areas more severely affected by COVID-19 were less likely to vote for the Trump, vice-versa.
+This aligns with mainstream media and intuitive expectations.
 
@@
 ```{r}
-binned_data <- data %>%
-  mutate(dpc_pctile = ntile(dpc, 100)) %>%  # Adjust 'breaks' to change number of bins
-  group_by(dpc_pctile) %>%
-  summarize(mean_rep = mean(pct_vote_rep, na.rm = TRUE),
-            mean_demo = mean(pct_vote_demo, na.rm = TRUE)) %>% 
-  mutate(bin = 1:n())
-
-binned_data2 <- data %>%
-  mutate(dpc_pctile = ntile(mean_household_income, 100)) %>%  # Adjust 'breaks' to change number of bins
-  group_by(dpc_pctile) %>%
-  summarize(mean_rep = mean(pct_vote_rep, na.rm = TRUE),
-            mean_demo = mean(pct_vote_demo, na.rm = TRUE)) %>% 
-  mutate(bin = 1:n(),
-         sum = mean_rep + mean_demo)
-
-# Plotting
-ggplot(binned_data, aes(x = bin, y = mean_rep)) +
-  geom_point() +  # or geom_line() depending on your preference
-  geom_smooth(method = "lm",formula = y ~ poly(x, 4), se=FALSE) +
-  theme_minimal() +
-  labs(x = "Binned mortality per case ", y = "Mean vote for Trump",
-       title = "Votes for Trump vs. MPC")
-ggplot(binned_data, aes(x = bin, y = mean_demo)) +
-  geom_point() +  # or geom_line() depending on your preference
-  geom_smooth(method = "lm", formula = y ~ poly(x, 4),se=FALSE) +
-  theme_minimal() +
-  labs(x = "Binned mortality per case ", y = "Mean vote for Biden", 
-       title = "Votes for Biden vs. MPC")
-
-ggplot(binned_data2, aes(x = bin, y = mean_rep)) +
-  geom_point() +  # or geom_line() depending on your preference
-  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se=FALSE) +
-  theme_minimal() +
-  labs(x = "Binned income levels", y = "Mean vote for Trump", 
-       title = "Votes for Trump vs. Income")
-ggplot(binned_data2, aes(x = bin, y = mean_demo)) +
-  geom_point() +  # or geom_line() depending on your preference
-  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se=FALSE) +
-  theme_minimal() +
-  labs(x = "Binned income levels", y = "Mean vote for Biden",
-       title = "Votes for Biden vs. Income")
-```
-
-```{r}
 #| echo: false
 #| message: false
-#| fig.width: 10 
-#| fig.height: 10
+#| fig.width: 8 
+#| fig.height: 8
 #| label: fig-map
@@
   theme_void() +
-  geom_path(data = state, aes(x = long, y = lat, group = group), color = "black", linewidth = 0.5)
+  geom_path(data = state, aes(x = long, y = lat, group = group), color = "black", linewidth = 0.9)
 
@@
   theme_void() +
-  geom_path(data = state, aes(x = long, y = lat, group = group), color = "black", linewidth = 0.5)
-
-p3 <- ggplot(data = merged_data, aes(x = long, y = lat, group = group)) + 
-  geom_polygon(aes(fill = mean_household_income), color = "black") +
-  scale_fill_gradient(low = "white", high = "darkblue", na.value = "grey",
-                      limits = c(60000, 120000), oob = scales::squish) +
-  labs(title = "The distribution of mean household income",
-       fill = 'Mean Hosehold Income') +
-  coord_quickmap() +
-  theme_void() +
-  geom_path(data = state, aes(x = long, y = lat, group = group), color = "black", linewidth = 0.5)
+  geom_path(data = state, aes(x = long, y = lat, group = group), color = "black", linewidth = 0.9)
 
 # Arrange the plots in a 2x1 layout
-grid.arrange(p1, p3, p2, ncol = 1)
+grid.arrange(p1, p2, ncol = 1)
+```
+
+We can validate our guess from @fig-map which compares the relative ratio of votes for the two parties and the death to cases rate in maps.
+From the maps, it is clearly more counties had preference for the Republican but the counties with higher death per case rate have a smaller ratio of vote for the Republican and these counties are usually richer ones.
+The most significant example can be California and New York.
+In contrast, the states with less impacted by COVID, such as Utah, vote more for Republican.
+
+```{r}
+#| fig-cap: The correlation between voting behaviors to income and DPC rate for the Republican and Democrat
+#| label: fig-dpc-income
+
+binned_data <- data %>%
+  mutate(dpc_pctile = ntile(dpc, 100)) %>%  
+  group_by(dpc_pctile) %>%
+  summarize(mean_rep = mean(pct_vote_rep, na.rm = TRUE),
+            mean_demo = mean(pct_vote_demo, na.rm = TRUE),
+            mean_vote_rep = mean(vote_rep, na.rm = TRUE),
+            mean_vote_demo = mean(vote_demo, na.rm = TRUE),
+            mean_income = mean(mean_household_income))
+
+binned_data2 <- data %>%
+  mutate(income_pctile = ntile(mean_household_income, 100)) %>%  
+  group_by(income_pctile) %>%
+  summarize(mean_rep = mean(pct_vote_rep, na.rm = TRUE),
+            mean_demo = mean(pct_vote_demo, na.rm = TRUE),
+            mean_vote_rep = mean(vote_rep, na.rm = TRUE),
+            mean_vote_demo = mean(vote_demo, na.rm = TRUE)) %>% 
+  mutate(sum = mean_rep + mean_demo)
+
+plot1 <- ggplot(binned_data, aes(x = dpc_pctile)) +
+  geom_point(aes(y = mean_rep, colour = "Republican"), alpha = 0.25) +  
+  geom_smooth(aes(y = mean_rep, colour = "Republican"), 
+              method = "lm", formula = y ~ poly(x, 3), se = FALSE) +
+  geom_point(aes(y = mean_demo, colour = "Democrat"), alpha = 0.25) + 
+  geom_smooth(aes(y = mean_demo, colour = "Democrat"), 
+              method = "lm", formula = y ~ poly(x, 3), se = FALSE) +
+  geom_vline(xintercept = 25, 
+             color = "black", linetype = "dashed", size = 0.6) +
+  geom_vline(xintercept = 75, 
+             color = "black", linetype = "dashed", size = 0.6) +
+  scale_colour_manual(values = c("Republican" = "#FF9999", "Democrat" = "#9999FF"),
+                      name = "Party",  
+                      labels = c("Democrat", "Republican")) +
+  theme_minimal() +
+  theme(legend.position = "none") + 
+  labs(x = "DPC Rate Percentiles", 
+       y = "Votes Share (%)")
+
+plot2 <- ggplot(binned_data, aes(x = dpc_pctile)) +
+  geom_point(aes(y = mean_vote_rep, colour = "Republican"), alpha = 0.25) +  
+  geom_smooth(aes(y = mean_vote_rep, colour = "Republican"), 
+              method = "lm", formula = y ~ poly(x, 3), se = FALSE) +
+  geom_point(aes(y = mean_vote_demo, colour = "Democrat"), alpha = 0.25) + 
+  geom_smooth(aes(y = mean_vote_demo, colour = "Democrat"), 
+              method = "lm", formula = y ~ poly(x, 3), se = FALSE) +
+  geom_vline(xintercept = 40, 
+             color = "black", linetype = "dashed", size = 0.6) +
+  scale_colour_manual(values = c("Republican" = "#FF9999", "Democrat" = "#9999FF"),
+                      name = "Party",  
+                      labels = c("Democrat", "Republican")) +
+  theme_minimal() +
+  theme(legend.position = "none") + 
+  labs(x = "DPC Rate Percentiles", 
+       y = "Number of Votes")
+
+pred <- predict(lm(mean_rep ~ poly(income_pctile, 3), data = binned_data2), 
+                           newdata = binned_data2)
+
+plot3 <- ggplot(binned_data2, aes(x = income_pctile)) +
+  geom_point(aes(y = mean_rep, colour = "Republican"), alpha = 0.25) +  
+  geom_smooth(aes(y = mean_rep, colour = "Republican"), 
+              method = "lm", formula = y ~ poly(x, 3), se = FALSE) +
+  geom_point(aes(y = mean_demo, colour = "Democrat"), alpha = 0.25) + 
+  geom_smooth(aes(y = mean_demo, colour = "Democrat"), 
+              method = "lm", formula = y ~ poly(x, 3), se = FALSE) +
+  geom_vline(xintercept = which.max(pred), 
+             color = "black", linetype = "dashed", size = 0.6) +
+  scale_colour_manual(values = c("Republican" = "#FF9999", "Democrat" = "#9999FF"),
+                      name = "Party",  
+                      labels = c("Democrat", "Republican")) +
+  theme_minimal() +
+  theme(legend.position = "none") + 
+  labs(x = "Income Percentiles", 
+       y = "Votes Share (%)")
+
+plot4 <- ggplot(binned_data2, aes(x = income_pctile)) +
+  geom_point(aes(y = mean_vote_rep, colour = "Republican"), alpha = 0.25) +  
+  geom_smooth(aes(y = mean_vote_rep, colour = "Republican"), 
+              method = "lm", formula = y ~ poly(x, 3), se = FALSE) +
+  geom_point(aes(y = mean_vote_demo, colour = "Democrat"), alpha = 0.25) + 
+  geom_smooth(aes(y = mean_vote_demo, colour = "Democrat"), 
+              method = "lm", formula = y ~ poly(x, 3), se = FALSE) +
+  geom_vline(xintercept = 25, 
+             color = "black", linetype = "dashed", size = 0.6) +
+  geom_vline(xintercept = 78, 
+             color = "black", linetype = "dashed", size = 0.6) +
+  scale_colour_manual(values = c("Republican" = "#FF9999", "Democrat" = "#9999FF"),
+                      name = "Party",  
+                      labels = c("Democrat", "Republican")) +
+  theme_minimal() +
+  theme(legend.position = "none") + 
+  labs(x = "Income Percentiles", 
+       y = "Number of Votes")
+
+library(cowplot)
+legend <- get_legend(plot1 + theme(legend.position = "bottom"))
+combined_plot <- plot_grid(plot1, plot2, plot3, plot4, ncol = 2)
+combined_plot_with_legend <- plot_grid(combined_plot, legend, ncol = 1, rel_heights = c(1, 0.1))
+
+print(combined_plot_with_legend)
+```
+
+Given the insights from @tbl-covid and @fig-map, @fig-dpc-income further examine the impact of COVID and income levels on the voting behaviors for the two parties.
+@fig-dpc-income shows the correlation between the votes of the two parties to the death per case rate and income levels.
+The two graphs on the left indicates the share of votes whereas the right two graphs indicate the number of votes.
+The two parties are represented by different colors.
+In terms of share of votes, in general, there is a positive relationship between the DPC and votes for the Democrat but a negative one for the Republican.
+However, the counties with medium value of DPC is less sensitive to the voting for the two parties, compared to the "tail" counties which their support change significantly as DPC increases or decreases.
+Meanwhile, there is a quadratic patterns between the vote for the two parties and income, it seems that the Republican is in favor of low income counties, but as income increase, more counties vote for Bide not Trump.
+
+In addition, even though the average share of votes for Republican is always higher than 0.5, it does not necessarily mean that the Democrat lost in all counties.
+The reason behind this is, as mentioned in @tbl-covid, the counties which the Republican won is about five times as higher as the Democrat.
+Since Republican has more counties, it definitely has a higher average share of votes than Democrat.
+This pattern also indicate that Democrat usually won the densely populated, urban areas but Republican won in more sparsely populated, rural areas \[citation\].
+The high population density areas has more votes and explains why Biden beat Trump.
+We can verify this from the number of votes for the two parties, where the number of votes for Biden exceeds Trump as DPC and income level increases.
+This pattern is especially significant for the top 50% richest counties and the counties with DPC larger than 75 quantile.
+
+```{r}
+data_population <- data %>% 
+  mutate(pop_pctile = ntile(total_population, 100)) %>% 
+  group_by(pop_pctile) %>% 
+  summarise(county_demo = sum(winning_party == "Democrat")/n(),
+            county_rep = sum(winning_party == "Republican")/n())
+
+data_population %>% 
+  ggplot(aes(x = pop_pctile)) +
+  geom_point(aes(y = county_rep, colour = "Republican"), alpha = 0.25) +  
+  geom_smooth(aes(y = county_rep, colour = "Republican"), 
+              method = "lm", formula = y ~ poly(x, 3), se = FALSE) +
+  geom_point(aes(y = county_demo, colour = "Democrat"), alpha = 0.25) + 
+  geom_smooth(aes(y = county_demo, colour = "Democrat"), 
+              method = "lm", formula = y ~ poly(x, 3), se = FALSE) +
+  scale_colour_manual(values = c("Republican" = "#FF9999", "Democrat" = "#9999FF"),
+                      name = "Party",  
+                      labels = c("Democrat", "Republican")) +
+  theme_minimal() +
+  theme(legend.position = "bottom") + 
+  labs(x = "Population Percentiles", 
+       y = "Share of winning counties")
 ```
@@
 ```{r}
-#| eval: false
-library(sp)
-library(geojsonio)
-library(ggplot2)
-library(dplyr)
-library(broom)
-hex_county <- geojson_read("us_county_hexgrid.geojson", what = "sp")
+boundaries <- c(0, 5000, 9000, 14000, 19000, 26000, 37000, 54000, 94000, 215000, Inf)
 
-# Reformat the data as needed (depending on your data structure)
-hex_county@data <- hex_county@data %>% 
-  mutate(adjusted_field = gsub("some_pattern", "replacement", original_field))
+# Classify the counties into percentiles
+data_population <- data %>% 
+  mutate(pop_pctile = cut(total_population, breaks = boundaries, include.lowest = TRUE, labels = FALSE))
 
-# Fortify the data for ggplot
-hex_county_fortified <- tidy(hex_county, region = "desired_region_field")
+# data_population %>% 
+#   filter(winning_party == "Republican") %>% 
+#   ggplot(aes(x = pop_pctile)) +
+#   geom_bar(aes(y = n()))
+# 
+# data_population %>% 
+#   group_by(pop_pctile, winning_party) %>% 
+#   summarise(n = n())
+# 
+# county_sum <- data_population %>%
+#   filter(winning_party == "Republican") %>%
+#   group_by(pop_pctile) %>%
+#   summarize(count = n())
 
-# Plot the hexagon map
-ggplot() +
-  geom_polygon(data = hex_county_fortified, aes(x = long, y = lat, group = group), fill="#a1dab4", color="#f7f7f7") +
-  geom_text(data = hex_county_fortified, aes(x = long, y = lat, label = label_field), size = 3) +
-  theme_void() +
-  coord_map()
+# Plot the data
+# ggplot(county_sum, aes(x = pop_pctile, y = count)) +
+#   geom_bar(stat = "identity")
 ```
 
-## Data Limitation
+From four graphs in @fig-dpc-income, we can conclude that, in general, the higher the DPC rate, the lower the support for Trump but higher for Biden.
+Besides, the Democrat is in favor of richer voters, where as Republican is in favor of low income counties.
+It is direct to ask whether these two variables have interaction effect on the voting patterns?
+However, it will be difficult to visualize the interaction effect between two continuous factors.
+Instead, I will manually set a new categorical variable `high_income` to denote whether the county has a high income with certain cutoff.
+Then the difference of correlations of the two groups as death per case rate changes.
+I will set different cutoffs to fully test the interaction effect.
 
-Although the merged data covers the voting patterns, socio-economic characteristics and COVID-19 infection rate by county, some counties are missing.
-From the above two maps, the counties in the middle west show grey, indicating that the data for these counties are absent.
-This absence of data may increase bias and influence the accuracy of data analyses.
+```{r}
+#| echo: false
+#| message: false
+#| label: fig-boxplot
+#| fig-cap: Summary of income levels for counties voting for Democratic and Republican
+
+data  %>%
+  ggplot(mapping = aes(x = winning_party, y = mean_household_income)) +
+  geom_boxplot() +
+  geom_jitter(alpha = 0.04, width = 0.3, height = 0) +
+  theme_minimal() +
+  labs(x = "Winning Party",
+       y = "Mean Household Income")
+```
+
+```{r, fig.width=10, fig.height=4.5}
+#| echo: false
+#| message: false
+#| label: fig-income_state
+#| fig-width: 10
+#| fig-height: 4.5
+#| fig-cap: Variation in COVID-19 Mortality Rates Across States, Categorized by Income Levels and Dominant Political Preferences
+
+# Find the wining data for each state
+data_win_party <- data %>%
+  group_by(state) %>%
+  summarise(dem_votes = sum(vote_demo),
+            rep_votes = sum(vote_rep), 
+            .groups = 'drop') %>% 
+  mutate(winning_party_state = ifelse(dem_votes > rep_votes, 'Democratic', 'Republican')) %>% 
+  dplyr::select(state, winning_party_state)
+# Calculate the average income
+average_income_by_state <- data %>%
+  group_by(state) %>%
+  summarise(average_income = mean(mean_household_income), .groups = 'drop')
+# Join the winning party and average income back to the original data
+data_win_party_income <- data %>%
+  left_join(data_win_party, by = "state") %>%
+  left_join(average_income_by_state, by = "state")
+# Arrange states by average income
+ordered_states <- data_win_party_income %>%
+  arrange(average_income) %>%
+  .$state %>%
+  unique()
+# Create a factor with levels based on income-ordered states for plotting
+data_win_party_income$state <- factor(data_win_party_income$state, levels = ordered_states)
+
+average_dpc <- mean(data_win_party_income$mortrate)
+
+ggplot(data_win_party_income, aes(x = state, y = mortrate, fill = winning_party_state)) +
+  geom_boxplot() +
+  geom_hline(yintercept = average_dpc, linetype = "dashed", color = "black") +
+  annotate("text", x = Inf, y = average_dpc, label = " National Average", hjust = 1, vjust = -0.5, color = "black", size = 3) +
+  scale_fill_manual(values = c('Democratic' = 'blue', 'Republican' = 'red')) +
+  labs(y = 'Death Rate', x = 'State (Ascending Income)', 
+       fill = "Winning Party") +
+  theme_minimal() +
+  theme(legend.position = "bottom") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+```
+
+@fig-boxplot compares the distribution of county income levels for the two parties.
+In general, counties voting for the Democrats have higher income levels than the Republicans.
+In addition, we can observe that the counties are intensively concentrated around income levels \$50k to \$75k, compared to the Democrats, where the counties are approximately uniformly distributed at each income level.
+Furthermore, barely any county has at least 150k voting for the Republican party.
+Both @tbl-covid and @fig-boxplot indicates that the Democrat is in favour of wealthy counties but poorer counties for Republican.
+Consequently, the more affluent counties (states) usually have more electoral votes than the poorer ones; this may provide insights into why Trump lost.
 
@@
 
-This paper aims to make a causal inference between the 2020 voting patterns and COVID-19 to find whether Trump can contribute to his loss of COVID-19.
-To see the causal effect in the observational study, I will implement the methods of propensity scores to find the change in votes for Trump due to the pandemic.
-Based on the results, I will also conduct a counterfactual analysis to see what would have happened if there had been no COVID-19,
+The objective in this paper is to conduct the causal inference between COVID-19 and Trump's loss during 2020, the main methods implemented in this study will be the Propensity Score Matching (PSM).
+After find the causal inference, this paper will also conduct a counterfactual analysis to see whether Trump can re-elect if there was no COVID.
 
-First and foremost, we need to define the treatment.
-The treatment in an experiment will be artificially imposed on specific groups of people to see the effect of interested variables with and without this treatment.
-Therefore, the people with the treatment are called the treatment group, and the rest are called the control group.
-The treatment and control group settings are to test whether the imposed factor impacts the interested variable.
-However, since this is an observational study, it is unrealistic to set a factor manually.
-Therefore, to find the causal effect of COVID-19 on Trump's voting, I will choose the treatment to be whether a county has a high infection rate.
+## Treatment
 
-## Propensity Score
+The treatment refers to the intervention or exposure that is being studied to understand its potential impact on the outcome.
+In other words, we implement certain interventions to a group of people but not for the rest.
+The group of people receiving the treatment is called the treatment group and control group for the rest.
+For instance, Jochen \[citation\] examined whether a new medical treatment called clampless off-pump coronary artery bypass (clampless OPCAB) will reduce the in-hospital mortality rate, compared to the traditional way.
+In their studies, the treatment is the new medical surgery and patients who took this surgery is the treatment group.
+They compared the mortality rates between the two groups and concluded that the new treatment can significantly lower the stroke and mortality rate.
 
-Propensity score is a technique that can help us to decide which county would be in the treatment group.
-That said, the propensity score in this study is the probability of a county having a high infection rate, given a set of socioeconomic variables.
-Using the propensity score can ensure that the two counties with similar propensity scores have the same distribution of the observed socioeconomic variables, such as income level, even though they are in different groups.
-This is fantastic, as we can have many pairs from the treatment of the control group with similar covariates by fixing propensity scores.
-The key drawback is that we can not estimate the causal effect from observational studies.
+Normally, the treatment should be established before conduct the analysis, this aims to reduce the risk of p-hacking.
+However, the objective of this paper is to find whether Trump lost the last Election due to COVID-19, it may not be possible to state the treatment in advance as all the counties in US had experienced COVI-19.
+Therefore this paper will firstly perform exploratory analysis to firstly find the treatment group which has the most significant difference in voting for Trump.
+This paper will set a cutoff defining whether a county has a high or low death per case rate.
+In addition, from @fig-dpc-income, low and high income counties also seem to have different voting behaviors.
+Therefore the treatment in this paper will incorporate whether a county has a high DPC and a high income levels with certain cutoffs.
 
-In addition, by finding the propensity scores, we can reconstruct our assignment mechanism to a strongly ignorable one.
-An assignment mechanism is "strongly ignorable" if and only if, given the observed covariates, the potential outcomes are independent of the treatment assignment.
-This indicates that the voting change for Trump is independent of our assignment that a county has a high or low infection rate.
+However, we need to be very careful about the cutoffs for defining "high" in DPC and income is crucial as the risk of p-hacking or data dreging may raise.
+P-hacking means the manipulation of data analysis until we can find the statistically significant results [citation].
+If we simply choose one cutoff and find the significant results that match the common sense, it may raise the p-hacking risk and the analyses will be less convincing.
+Therefore, to make the analysis more comprehensive, instead of choosing a single cutoff, I will set a grid of cutoffs for each variable to see how the separate or combine treatment effects varies.
+For example, one treatment groups can be counties with death per case and income levels higher than 0.3 quantiles.
+By this way, we should be able to find the optimal treatment group and avoid the risk of p-hacking.
 
-As the propensity score represents the probability of a county having a high infection rate, we must find a logistic regression to predict this probability.
-In my previous paper, I already found the best model predicting the mortality rate of COVID in each US county.
-Even though that paper was to predict the mortality rate, the mortality is correlated with the infection rate.
-Therefore, I will take the dependent variables in that model to be the covariates in predicting the propensity score for each county, which is:
+## Propensity Score Matching (PSM)
+
+To estimate the treatment effects, randomized controlled trials (RCT) is the most ideal and the golden way to do this \[citation\].
+The reason is that RCTs is the only way that guarantees the equal distribution of known and unknown parameters.
+Therefore the outcomes between the treatment and control groups will not be confounded.
+However, conducting a RCT is nearly impractical and impossible.
+Back to the coronary artery bypass example \[citation\], you can not force a patient to take new or traditional treatment, the researchers can not control it and hence RCT is impossible.
+Instead, we can take observational studies to find the treatment effects.
+
+One method that can be applied to observational studies to find the treatment effect is Propensity Score Matching (PSM).
+Propensity score is the probability of an observation that receive the treatment based on the existing covariates using logistic regression.
+Using propensity scores have one main advantages which is dimension reduction [citation].
+Dimension reduction means we can describe an observation by only its propensity score instead of a list of covariates.
+Therefore, we can match two observations having similar propensity scores but from treatment and control groups separately.
+With this algorithm, we can have many matched pairs and the matching process is called propensity score matching.
+Here, we have another advantage of PSM which is design separation [citation], meaning that it can separate the covariates balancing and effects estimation.
+I will show the balance of covariates later.
+Finally, we can directly estimate their outcomes and solve the drawback that we can not have causal inference from observational studies.
+That is why it can also separate designs and this process is called propensity score matching.
+
+In this paper, the propensity scores are used to estimate how likely a county having a high DPC or high income levels (or both) using multiple linear logistic regression. The model will be look like:
 
@@
 \begin{align*}
-\frac{PS}{1-PS} &=  \beta_0 + \\
-& + \beta_1 \times \text{prop\_higher\_education} \\
-& + \beta_2 \times \text{IncomePctile} \\
-& + \beta_3 \times \text{no\_insurance} \\
-& + \beta_4 \times \text{private\_insurance} \\
-& + \beta_5 \times \text{males} \\
-& + \beta_6 \times \text{old\_85} \\
-& + \beta_7 \times \text{white\_pct} \\
-& + \beta_8 \times \text{black\_pct}
+\log\left(\frac{P(T=1|X)}{1 - P(T=1|X)}\right) = \beta_0 + \beta_1 X_1 +\cdots +\beta_p X_p
+\end{align*}
+```
+where:\
+* $P(T=1|X)$ is the probability of a county receiving treatment\
+* $X_j\:(1\leq j\leq p)$ are the covariates\
+* $\beta_0$ is the intercept
+* $\beta_j\:(1\leq j\leq p)$ is the corresponding coefficients of each covariate\
+
+Using the model, I will calculate the propensity scores for each county and match the data with similar propensity scores.
+Notice that the values of propensity scores changes as the treatment changes.
+That said, I will calculate the propensity scores and match the data for number of treatment times.
+
+## Counterfactual Analysis
+
+Using the PSM described previously, we can find the treatment effects under different settings of treatment groups.
+However, to conclude whether Trump was lost due to COVID, we need to perform a counterfactual analysis with re-calculating the votes for Trump of counties in the treatment group.
+
+The couterfactual analysis is particularly necessary for those "swing" states in 2020.
+In 2016 Federal Election, one key factory why Trump defeated Hillary is that he won seven out of eleven swing states.
+However, he only won three of them in 2020 [citaion (wikipedia)].
+Besides, if Donald Trump were able to hold onto three swing states which are Georgia, Arizona and Wisconsin where the average margin of Democrat is only about 0.3%, the result would have been a 269-269 electoral tie decided in the House of Representatives [citation (wikipedia)].
+Then the presidential election is left up to members of the House of Representatives and it is possible that Trump can win.
+
+To find the voting patterns for Trump, especially for the "swing" states, I will follow the official election procedure and use the winners-takes-all rule.
+That said, I will re-calculate the votes for Trump in county level and summarize by county level.
+Then the party with higher total votes will take all the electoral votes in this state.
+Eventually, I will calculate the total electoral votes for each party to see whether Trump could re-elect if the treatment effects diminish.
+Besides, I will also predict the votes if the death number can reduce 5%, 10%..., to see if a tiny decrease in deaths will make Trump to re-elect.
+
+# Results
+
+## Choise of treatment groups
+
+```{r}
+#| fig-width: 8
+#| fig-height: 5
+#| fig-cap: 'The Correlation Between Voting for Trump and Income Levels, Categorized by DPC Rate at Varied Quantile Cutoffs (e.g., Top 40% as High-DPC and bottom 60% as Low-DPC for "Cutoff: 0.6")'
+#| label: fig-dpc
+#| message: false
+#| warning: false
+
+generate_plot2 <- function(cutoff, y_var, colours, y_limits) {
+  data_processed <- data %>%
+    mutate(income_pctile = ntile(mean_household_income, 20),
+           high_dpc = as.factor(ifelse(dpc > quantile(dpc, cutoff), 1, 0))) %>%
+    group_by(income_pctile, high_dpc) %>%
+    summarise(y_value = mean({{ y_var }}), .groups = "drop")  # Use dynamic variable
+
+  ggplot(data_processed, aes(x = income_pctile, y = y_value, color = high_dpc)) +
+    geom_point(alpha = 0.25) +
+    geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE) +
+    theme_minimal() +
+    scale_color_manual(values = c("1" = colours[1], "0" = colours[2]),
+                       name = "DPC Level",
+                       labels = c("0" = "Low DPC", "1" = "High DPC")) +
+    ylim(y_limits[1], y_limits[2]) +
+    theme(legend.position = "none",
+          axis.title.x = element_blank(),
+          axis.title.y = element_blank(),
+          plot.title = element_text(size = 10)) + 
+    labs(x = "Income Percentile", y = deparse(substitute(y_var)),
+         title = paste("Cutoff:", cutoff))
+}
+
+cutoffs <- seq(0.1, 0.9, by = 0.1)
+
+# Assuming 'pct_vote_rep' is a column in your 'data' dataframe
+plots <- lapply(cutoffs, function(cutoff) generate_plot2(cutoff, 
+                                                        y_var = pct_vote_rep,
+                                                        c("#FF6347", "#FFB6C1"),
+                                                        c(0.4, 0.8)))
+
+library(cowplot)
+legend <- get_legend(plots[[1]] + theme(legend.position = "bottom",  # Place legend at the bottom
+        legend.justification = "center",  # Justify legend in the center
+        legend.box.just = "bottom"))
+combined_plot <- plot_grid(plotlist = plots[1:9], ncol = 3)
+combined_plot_with_legend <- plot_grid(combined_plot, legend, ncol = 1, rel_heights = c(1, 0.1))
+ggdraw() +
+  draw_plot(combined_plot_with_legend) +
+  draw_label("Income Rate Percentiles", x = 0.5, y = 0, hjust = 0.5, size = 12) +
+  draw_label("Share of Vote for Republican", x = -0.03, y = 0.5, angle = 90, vjust = 0.5, size = 12) +
+  theme(plot.margin = margin(t = 10, r = 10, 
+                             b = 10, l = 40, unit = "pt"))
+```
+
+@fig-dpc shows the correlation between votes for Trump and income levels, segmented into "High-DPC" and "Low-DPC" county groups at varying DPC cutoffs.
+For instance, a cutoff of 0.3 classifies counties with a DPC rate below this threshold as "Low DPC," and those above as "High DPC.".
+By comparing these groups across different cutoffs, we aim to identify the cutoff where the difference in voting for Trump across difference income levels is most pronounced.
+With that said, we can find that except cutoff 0.4, the two line converge as income increases (right tail).
+Only for the 0.4 cutoff, the lines of two groups converge at the middle-income but do not overlap and diverge at the tails.
+Additionally, regardless of the cutoff applied, there is a common pattern that the support for Trump increases in poorer counties up to around the eighth bin, which is approximately the 0.4 income quantile, and then decreases in wealthier counties.
+This indicates a significant voting disparity between counties below and above the 0.4 income quantile.
+
+```{r}
+#| fig-width: 8
+#| fig-height: 5
+#| fig-cap: 'The Correlation Between Voting for Trump and DPC Rate, Categorized by Income Levels at Varied Quantile Cutoffs (e.g., Top 40% as High Income and bottom 60% as Low Income for "Cutoff: 0.6")'
+#| label: fig-income
+#| message: false
+#| warning: false
+
+generate_plot <- function(cutoff, y_var, colours, y_limits) {
+  data_processed <- data %>%
+    mutate(dpc_pctile = ntile(dpc, 20),
+           high_income = as.factor(ifelse(mean_household_income > quantile(mean_household_income, cutoff), 1, 0))) %>%
+    group_by(dpc_pctile, high_income) %>%
+    summarise(y_value = mean({{ y_var }}), .groups = "drop")  # Use dynamic variable
+
+  ggplot(data_processed, aes(x = dpc_pctile, y = y_value, color = high_income)) +
+    geom_point(alpha = 0.25) +
+    geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE) +
+    theme_minimal() +
+    scale_color_manual(values = c("1" = colours[1], "0" = colours[2]),
+                       name = "Income Level",
+                       labels = c("0" = "Low Income", "1" = "High Income")) +
+    ylim(y_limits[1], y_limits[2]) +
+    theme(legend.position = "none",
+          axis.title.x = element_blank(),
+          axis.title.y = element_blank(),
+          plot.title = element_text(size = 11)) + 
+    labs(x = "DPC Percentile", y = deparse(substitute(y_var)),
+         title = paste("Cutoff:", cutoff))
+}
+
+cutoffs <- seq(0.1, 0.9, by = 0.1)
+
+# Assuming 'pct_vote_rep' is a column in your 'data' dataframe
+plots <- lapply(cutoffs, function(cutoff) generate_plot(cutoff, 
+                                                        y_var = pct_vote_rep,
+                                                        c("#FF6347", "#FFB6C1"),
+                                                        c(0.4, 0.8)))
+
+library(cowplot)
+legend <- get_legend(plots[[1]] + theme(legend.position = "bottom",  # Place legend at the bottom
+        legend.justification = "center",  # Justify legend in the center
+        legend.box.just = "bottom"))
+combined_plot <- plot_grid(plotlist = plots[1:9], ncol = 3)
+combined_plot_with_legend <- plot_grid(combined_plot, legend, ncol = 1, rel_heights = c(1, 0.1))
+ggdraw() +
+  draw_plot(combined_plot_with_legend) +
+  draw_label("DPC Rate Percentiles", x = 0.5, y = 0, hjust = 0.5, size = 12) +
+  draw_label("Share of Vote for Republican", x = -0.03, y = 0.5, angle = 90, vjust = 0.5, size = 12) +
+  theme(plot.margin = margin(t = 10, r = 10, 
+                             b = 10, l = 40, unit = "pt"))
+```
+
+Furthering on our analysis, @fig-income shows the correlation between Trump's vote share and the DPC rate but this time dividing counties based on income levels into "High Income" and "Low Income" groups.
+Each subplot in this analysis corresponds to a different income cutoff.
+Compared to @fig-dpc, all curves, regardless of the cutoff, display a linear or near-linear relationship with voting patterns.
+As the income cutoff increases, the difference in voting behavior between the high and low-income counties becomes more distinct, especially at the 0.9 cutoff.
+This suggests that in counties with higher incomes, voting patterns vary significantly from those in lower-income counties as the DPC rate changes.
+Therefore, the most significant pattern is observed at the 0.9 income cutoff, contradicted to the patterns identified in @fig-dpc.
+
+So, should we include all the income levels and DPC in the settings of treatment?
+The answer is NO and we should only include the DPC.
+From @fig-dpc, a consistent quadratic relationship between income and voting patterns for Trump is evident, regardless of how the cutoff for defining high DPC is adjusted.
+This pattern demonstrates that poorer counties (in the first half of the income spectrum) may show a positive or a positive correlation with voting for Trump depending on the cutoff we choose, whereas richer counties will only exhibit a negative correlation.
+This quadratic pattern can be verifies from @fig-dpc-income, which illustrates a similar relationship between votes and income.
+Conversely, @fig-income presents a different narrative.
+Here, all regression lines, irrespective of the DPC cutoff, display a linear or nearly linear relationship.
+Besides, unlike the @fig-dpc, all the lines in @fig-income shows a negative slope, meaning that the choice of cutoff for "High DPC" group will not change the general voting patterns for Trump.
+The optimal cutoff we found previously, 0.4, is the one that the "High DPC" and "Low DPC" have the maximum difference.
+This finding can be also verified from @fig-dpc-income, which shows a monotonically decreasing cubic correlation between votes and DPC.
+The @fig-treatment shows a simple visualization of why is the case.
+
+```{r}
+#| fig-width: 8
+#| fig-height: 3
+#| label: fig-treatment
+# Splitting the data into two subsets
+left_data <- subset(binned_data, dpc_pctile <= 65)
+right_data <- subset(binned_data, dpc_pctile > 65)
+
+# Creating new subsets for the x-intercept at 40
+left_data_40 <- subset(binned_data, dpc_pctile <= 40)
+right_data_40 <- subset(binned_data, dpc_pctile > 40)
+
+# Base plot with cubic regression and vertical lines
+p1 <- ggplot(binned_data, aes(x = dpc_pctile)) +
+  geom_smooth(aes(y = mean_rep), colour = "black", method = "lm", 
+              formula = y ~ poly(x, 3), se = FALSE, size = 1) +
+  geom_vline(xintercept = 65, color = "lightgrey", linetype = "dashed", size = 0.8) +
+  geom_vline(xintercept = 40, color = "#E69F00", linetype = "dashed", size = 0.8) +
+  geom_smooth(data = left_data, aes(y = mean_rep, colour = "65"), 
+                     method = "lm", formula = y ~ x, se = FALSE, linetype = "dashed", size = 0.8) +
+  geom_smooth(data = right_data, aes(y = mean_rep, colour = "65"), 
+                     method = "lm", formula = y ~ x, se = FALSE, linetype = "dashed", size = 0.8) + 
+  geom_smooth(data = left_data_40, aes(y = mean_rep, colour = "40"), 
+                method = "lm", formula = y ~ x, se = FALSE, linetype = "longdash", size = 0.8) +
+  geom_smooth(data = right_data_40, aes(y = mean_rep, colour = "40"), 
+                method = "lm", formula = y ~ x, se = FALSE, linetype = "longdash", size = 0.8) +
+  theme_minimal() +
+  scale_color_manual(values = c("65" = "lightgrey", "40" = "#E69F00"),
+                       name = "Cutoffs",
+                       labels = c("65" = "0.65 Quantile", "40" = "0.4 Quantile")) +
+  theme(legend.position = "bottom",
+          axis.title.y = element_blank()) +
+  labs(x = "DPC Rate Percentiles", y = "Votes Share (%)")
+
+# Splitting the data into two subsets
+left_data <- subset(binned_data2, income_pctile <= 90)
+right_data <- subset(binned_data2, income_pctile > 90)
+
+# Creating new subsets for the x-intercept at 40
+left_data_40 <- subset(binned_data2, income_pctile <= 40)
+right_data_40 <- subset(binned_data2, income_pctile > 40)
+
+p2 <- ggplot(binned_data2, aes(x = income_pctile)) +
+  geom_smooth(aes(y = mean_rep), colour = "black", 
+              method = "lm", formula = y ~ poly(x, 3), se = FALSE, size = 1) +
+  geom_vline(xintercept = 90, color = "lightgrey", linetype = "dashed", size = 0.8) +
+  geom_vline(xintercept = 40, color = "#E69F00", linetype = "dashed", size = 0.8) +
+  geom_smooth(data = left_data, aes(y = mean_rep, colour = "90"), 
+                     method = "lm", formula = y ~ x, se = FALSE, linetype = "dashed", size = 0.8) +
+  geom_smooth(data = right_data, aes(y = mean_rep, colour = "90"), 
+                     method = "lm", formula = y ~ x, se = FALSE, linetype = "dashed", size = 0.8) + 
+  geom_smooth(data = left_data_40, aes(y = mean_rep, colour = "40"), 
+                method = "lm", formula = y ~ x, se = FALSE, linetype = "longdash", size = 0.8) +
+  geom_smooth(data = right_data_40, aes(y = mean_rep, colour = "40"), 
+                method = "lm", formula = y ~ x, se = FALSE, linetype = "longdash", size = 0.8) +
+  theme_minimal() +
+  scale_color_manual(values = c("90" = "lightgrey", "40" = "#E69F00"),
+                       name = "Cutoffs",
+                       labels = c("90" = "0.90 Quantile", "40" = "0.4 Quantile")) +
+  theme(legend.position = "bottom",
+          axis.title.y = element_blank()) +
+  labs(x = "Income Percentiles", y = "Votes Share (%)")
+
+library(cowplot)
+
+combined_plot <- plot_grid(p1, p2, nrow = 1)
+ggdraw() +
+  draw_plot(combined_plot) +
+  draw_label("Share of Vote for Republican", x = -0.03, y = 0.5, angle = 90, vjust = 0.5, size = 12) +
+  theme(plot.margin = margin(t = 10, r = 10, 
+                             b = 10, l = 40, unit = "pt"))
+```
+
+@fig-treatment shows a simple example when we test different cutoffs for DPC rate and income.
+From the graph, we can see that no matter which cutoff we set, the low DPC and high DPC counties always have the a negative value of slope.
+The only difference is the absolute values of the slope.
+However, for the income, we can see that for the cutoff 0.4 quantile, the low income and high income groups show a opposite patterns for voting for Trump.
+But when we increase the cutoff to 0.9 quantile.
+The two groups are both negatively related to the change.
+This indicates that it is not appropriate to incorporate income in treatment as the choice of cutoff will significant change the
+
+no matter which cutoff we set, the two groups will always have opposite altitude for voting for Trump.
+Suppose we have enough samples, when we include income into the treatment, then treatment effect will always be more negative when we increase the cutoff.
+The treatment effect will be more and more significant when we increase the cutoff.
+However, the number of counties in treatment group will decrease and hence it will be unnecessary to have income into treatment settings.
+
+From @fig-dpc, we already verify that cutoff 0.4 is the optimal threshold for defining high and low DPC group, therefore the treatment group will be the counties having DPC rate larger than 0.4 quantile.
+Furthermore, to balance the effect of income, I will incorporate income into the ways to calculate the propensity scores for each county.
+
+## Propensity Score Matching
+
+In my previous paper, I already show that proportion of residence with at least a bachelor degree, income, proportion of people with private insurance and withou health insurance, ratio of males, ratio of black and black residence are statistically in predicting the probability of a county has a high mortality rate or not.
+In this paper, I will directly use the results and only consider these variables when predicting the propensity scores.
+
+```{=tex}
+\begin{align*}
+\log\left(\frac{P(T=1|X)}{1 - P(T=1|X)}\right) &= 8.06 \\
+& -0.02 \times \text{prop\_higher\_education} \\
+& +0.01 \times \text{pctile} \\
+& +0.01 \times \text{no\_insurance} \\
+& -0.03 \times \text{private\_insurance} \\
+& -0.13 \times \text{males} \\
+& +0.00 \times \text{white\_pct} \\
+& +0.06 \times \text{black\_pct}
 \end{align*}
@@
 
-Therefore, we can find the PS for each county using the above model. However, there are three ways to use the propensity scores to find the treatment effect.
-
-### Propensity Score Matching
-The package needed to implement the matching is the `Matching.` Using this package, we can match the observations in the treatment and control groups having the same or similar propensity scores.
-Then, take only the matched pairs to be the new data set and find the treatment effect between them.
-In addition, we can also check the balance between the treatment and control groups.
-
-However, the most significant drawback is that the new data set can only contain the matched pairs.
-We may lose a considerable amount of observations during this process.
-Our predictions may also not be accurate.
-
-### Propensity Score Stratification
-As introduced above, using PSM may cause the waste of data and results in bias.
-However, propensity score stratification will use the entire data set to avoid waste.
-The way to do this is to split the data into different quantiles and fit a model for each subclass, with the only dependent variable being the treatment variable.
-Then, calculate the average casual effect for the subclasses.
-
-Even though the PSS can use the entire data set, however, the choice of quantiles will be closely related to the predictions.
-If we have a large sample and want to split the data into many subclasses, then this method may be computationally slow.
-
-### Propensity Score Regression Adjustment
-The last method is the most straightforward, meaning we just fit a regression model with the treatment variable and propensity score as the two dependent variables.
-
-Each of the above three methods has its unique advantages, but the results may vary between different data sets. I will perform both of the three methods and compare which one would be the best.
-
-## Counterfactual Analysis
-Using the propensity score methodology, we can assess the causal effect of a high infection rate on the voting for Trump.
-However, it can not tell us the election results, and we can not draw conclusions about COVID and the election.
-Therefore, I will conduct a counterfactual analysis, meaning that I will re-calculate the votes for Trump and Biden to see if the final election results will change.
-
-Eventually, I will re-calculate the votes for the two candidates to follow the official election rule.
-By finding the difference in voting percentage, I will re-allocate the voting for the two parties and calculate the electoral votes in each state using the winner-takes-all rule.
-
-# Results
+The above equation shows the logistic regression model to predict the propensity score for each county.
+It seems that males is the most critical factor when predict the propensity score.
+Perhaps the majority of COVID deaths are males.
+Interestingly, the counties with a higher income levels will have a higher death per case rate.
+This may explain why richer counties are less likely to vote for Trump.
 
@@
 #| message: false
+#| tbl-cap: The summary of Propensity Score Matching
+#| label: tbl-psm
+
+library(dplyr)
 data <- read.csv(here("outputs/data/merged_data.csv"))
 
+data_te <- data %>% 
+  mutate(high_dpc = ifelse(dpc > quantile(dpc, 0.4), 1, 0),
+         high_income = ifelse(mean_household_income > quantile(mean_household_income, 0.9), 1, 0),
+         treatment = high_dpc)
 highinf_model <- glm(treatment ~ prop_higher_education  + pctile +
@@
                        white_pct + black_pct, family = binomial(),
-                      data = data)
-# summary(highinf_model)
-data$prop_score <- predict(highinf_model, type = "response")
+                      data = data_te)
+data_te$prop_score <- predict(highinf_model, type = "response")
+
+library(Matching)
+library(kableExtra)
+# Propensity score matching
+rr <- Match(Y = data_te$pct_vote_rep, Tr = data_te$treatment,
+            X = data_te$prop_score, M = 1)
+
+data.frame(estimate = -0.018342,
+           p = 0.021404,
+           on = 3115,
+           ot = 1869,
+           mo=1869,
+           mou = 5415) %>% 
+  rename(`Treatment Effect` = estimate,
+         `P value` = p,
+         `Obs.` = on,
+         `Treat. Obs.` = ot,
+         `Treat. Obs` = mo,
+         `Treat. Obs. Unw.` = mou) %>% 
+  kable(booktabs = TRUE) %>% 
+  add_header_above(c(" " = 2,
+                     "Original" = 2, "Matched" = 2))
 ```
 
+@tbl-psm summarize the results of Propensity Score Matching.
+The treatment effect is about -0.18 means that the counties with DPC rate higher than 0.4 will in average vote 0.018 less for Trump, compared to the counties with lower DPC rate.
+This value is statistically significant as its p value is about 0.02 which is lower than 0.05.
+In addition, there are 1869 treatment observations in both original and matched data, meaning that there some counties in control group that are matched more than once.
+
 ```{r}
-data <- read.csv(here("outputs/data/merged_data.csv"))
+#| label: tbl-balance
+#| tbl-cap: The balance of each covariates before and after the mathcing
+# Load necessary libraries
+library(knitr)
+library(kableExtra)
 
-library(MatchIt)
-m.out <- matchit(treatment ~ prop_higher_education + pctile + 
-                       no_insurance + private_insurance + males +
-                       white_pct + black_pct,
-                 data = data,
-                 distance = "logit",
-                 method = "nearest",
-                 replace = FALSE,
-                 ratio = 1)
-summary(m.out, standardize = TRUE)
+# Prepare your data
+data.frame(
+  var = c("prop_higher_education", "pctile", "no_insurance", "private_insurance", "males", "white_pct", "black_pct"),
+  before_mean_control = c(23.246, 51.774, 8.7107, 67.656, 50.512, 87.137, 4.1685),
+  before_std_mean_diff = c(-10.772, -10.389, 24.449, -35.913, -32.492, -47.207, 49.083),
+  after_mean_control = c(21.633, 47.804, 9.7494, 63.409, 49.552, 79.636, 11.456),
+  after_std_mean_diff = c(5.6159, 2.9221, 4.6766, 5.955, 9.6593, -5.5169, 5.4646)
+) %>% 
+  rename(`Variables` = var,
+         `Mean Contr` = before_mean_control,
+         `Std Mean Diff` = before_std_mean_diff,
+         `Mean Contr.` = after_mean_control,
+         `Std Mean Diff.` = after_std_mean_diff) %>% 
+  kable(booktabs = TRUE) %>%
+  kable_styling(full_width = FALSE) %>%
+  add_header_above(c(" " = 1, "Before Matching" = 2, "After Matching" = 2))
 ```
 
-```{=tex}
-\begin{align*}
-\text{Logit}(\text{PS}) &= -2.008 + \\
-& -0.032 \times \text{prop\_higher\_education} \\
-& -0.009 \times \text{IncomePctile} \\
-& +0.088 \times \text{no\_insurance} \\
-& +0.051 \times \text{private\_insurance} \\
-& +0.019 \times \text{males} \\
-& +0.055 \times \text{old\_85} \\
-& -0.030 \times \text{white\_pct} \\
-& +0.026 \times \text{black\_pct}
-\end{align*}
+@tbl-balance summarize each variable's mean values in the control group and standard mean difference before and after the matching.
+We can see that the mean values in the control group approximately remain same after the matching.
+Besides, there is a significant decrease in the standard mean difference for each variable.
+This indicates that the Propensity Score Matching reduced the imbalance between the treatment and control group considerablely.
+
+## Counterfactual Analysis
+
+```{r}
+#| tbl-cap: The summary of counterfactual votes for the eleven swing staes in 2020
+#| label: tbl-counter
+electoral_votes_state <- read.csv(here("outputs/data/electoral_votes_state.csv"))
+swing_states_2020 <- c("FL", "PA", "MI", "WI", "AZ", "NC", "GA", "IA", "NV", "MN", "NH")
+
+data_new <- data_te %>%
+  mutate(vote_rep = ifelse(treatment == 1, total_votes*(pct_vote_rep + 0.018342), vote_rep),
+         vote_demo = ifelse(treatment == 1, total_votes*(pct_vote_demo - 0.018342), vote_demo))
+
+data_c <- data_new %>%
+  filter(state %in% swing_states_2020) %>% 
+  group_by(state) %>% 
+  summarise(total_votes_rep_counter = sum(vote_rep),
+            total_votes_demo_counter = sum(vote_demo))
+
+data_counter <- data %>% 
+  filter(state %in% swing_states_2020) %>% 
+  group_by(state) %>% 
+  summarise(total_votes_rep_real = sum(vote_rep),
+            total_votes_demo_real = sum(vote_demo)) %>% 
+  left_join(data_c, by = "state") %>% 
+  mutate(margin_real = ifelse(total_votes_rep_real > total_votes_demo_real, 
+                         paste0(round(100 * (total_votes_rep_real - total_votes_demo_real) / (total_votes_rep_real + total_votes_demo_real), 2), "% R"),
+                         ifelse(total_votes_rep_real < total_votes_demo_real, 
+                                paste0(round(100 * (total_votes_demo_real - total_votes_rep_real) / (total_votes_rep_real + total_votes_demo_real), 2), "% D"),
+                                "Equal")),
+         margin_counter = ifelse(total_votes_rep_counter > total_votes_demo_counter, 
+                         paste0(round(100 * (total_votes_rep_counter - total_votes_demo_counter) / (total_votes_rep_counter + total_votes_demo_counter), 2), "% R"),
+                         ifelse(total_votes_rep_counter < total_votes_demo_counter, 
+                                paste0(round(100 * (total_votes_demo_counter - total_votes_rep_counter) / (total_votes_rep_counter + total_votes_demo_counter), 2), "% D"),
+                                "Equal"))) %>% 
+  arrange(desc(total_votes_demo_real/(total_votes_rep_real + total_votes_demo_real))) %>% 
+  left_join(electoral_votes_state, by = "state") %>% 
+  dplyr::select(state, electoral_votes, total_votes_rep_real, 
+                total_votes_demo_real, margin_real, total_votes_rep_counter, 
+                total_votes_demo_counter, margin_counter)
+
+data_counter %>% 
+  rename(`States` = state,
+         `Electoral Votes` = electoral_votes,
+         `Trump` = total_votes_rep_real,
+         `Biden` = total_votes_demo_real,
+         `Margin` = margin_real,
+         `Trump*` = total_votes_rep_counter,
+         `Biden*` = total_votes_demo_counter,
+         `Margin*` = margin_counter) %>% 
+  kable(booktabs = TRUE) %>% 
+  add_header_above(c(" " = 2, "Actual Results" = 3, "Simulated Results*" = 3))
 ```
 
-The above equation shows the logistic regression model to predict the propensity score for each county, given the covariates.
-It seems that insurance is a critical factor in the infection rate.
-Interestingly, the counties with a higher proportion of people with no insurance and private insurance will both increase the probability of this county having a high infection rate.
-Using this model, we can find the propensity score for each county and implement the three methods.
-```{r}
-model <- lm(pct_vote_rep ~ dpc + mean_household_income, data = data)
-summary(model)
-```
-
-```{r}
-#| echo: false
-#| message: false
-
-library(Matching)
-# Propensity score matching
-rr <- Match(Y = data$pct_vote_rep, Tr = data$treatment, 
-            X = data$prop_score, M = 1)
-summary(rr)
-```
-
-```{r}
-
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| label: tbl-PSS
-#| tbl-cap: The results of causal effect using propensity score statification
-#| eval: false
-# PSS
-sub <- quantile(data$prop_score, probs = c(0.2, 0.4, 0.6, 0.8))
-
-sub1 <- glm(pct_vote_rep ~ treatment,
-            data = data[data$prop_score <= sub[1], ])
-sub2 <- glm(pct_vote_rep ~ treatment,
-            data = data[data$prop_score > sub[1] & data$prop_score <= sub[2], ])
-
-sub3 <- glm(pct_vote_rep ~ treatment,
-            data = data[data$prop_score > sub[2] & data$prop_score <= sub[3], ])
-
-sub4 <- glm(pct_vote_rep ~ treatment,
-            data = data[data$prop_score > sub[3] & data$prop_score <= sub[4], ])
-
-sub5 <- glm(pct_vote_rep ~ treatment,
-            data = data[data$prop_score > sub[4], ])
-
-# Extracting the necessary information from each model
-extract_info <- function(model, quantile) {
-  coef_summary <- summary(model)$coefficients
-  data <- data.frame(
-    quantile = quantile,
-    ce = coef_summary["treatment", "Estimate"],
-    p_val = coef_summary["treatment", "Pr(>|t|)"]
-  )
-  return(data)
-}
-
-# Creating data frames for each model
-df1 <- extract_info(sub1, "0-20%")
-df2 <- extract_info(sub2, "20-40%")
-df3 <- extract_info(sub3, "40-60%")
-df4 <- extract_info(sub4, "60-80%")
-df5 <- extract_info(sub5, "80-100%")
-
-# Combining all data frames into one
-rbind(df1, df2, df3, df4, df5) %>%
-  rename(`PS Quantile` = quantile,
-         `Estimated CE` = ce,
-         `P Value` = p_val) %>%
-  kable()
-```
-
-```{r}
-#| echo: false
-#| message: false
-# Propensity score regression adjustment
-model_adj <- glm(change_vote_rep ~ treatment + prop_score, data = data)
-summary(model_adj)
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| eval: false
-# Counterfactual Analysis
-adjusted_data <- data %>% 
-  dplyr::select(state, county, fips, party, votes, total_votes, pct_vote, high_infrate)
-
-adjust_function <- function(data){
-  adjust_votes <- c()
-  for (i in 1:nrow(data)){
-    if (data[i,"high_infrate"] == 1){
-      if (data[i,"party"] == "Republican"){
-        adjust_votes[i] <- round((data[i,"pct_vote"] + 0.036928 ) * data[i,"total_votes"], 0)
-      }
-      else if (data[i,"party"] == "Democrat"){
-        adjust_votes[i] <- round((data[i,"pct_vote"] - 0.036928 ) * data[i,"total_votes"], 0)
-      }
-      else {adjust_votes[i] <- data[i,"votes"]}
-    }
-    else {
-      adjust_votes[i] <- data[i,"votes"]
-    }
-  }
-  return(adjust_votes)
-}
-
-adjusted_data %>% 
-  group_by(state, party) %>% 
-  summarise(votes = sum(votes), .groups = "drop")
-```
-
-```{r}
-#| echo: false
-#| message: false
-#| eval: false
-adjust_function2 <- function(data) {
-  adjust_votes <- numeric(nrow(data))  # Initialize the vector with the correct length
-  extra_vote_percentage <- 0.036928
-
-  for (i in 1:nrow(data)) {
-    if (data[i, "high_infrate"] == 1) {
-      total_votes <- data[i, "total_votes"]
-      extra_votes <- round(extra_vote_percentage * total_votes, 0)
-
-      if (data[i, "party"] == "Republican") {
-        adjust_votes[i] <- data[i, "votes"] + extra_votes
-      } else {
-        # Calculate total votes for non-Republican parties
-        total_non_rep_votes <- sum(data[data$party != "Republican" & data$fips == data[i, "fips"], "votes"])
-
-        # Distribute the vote loss proportionally
-        if (total_non_rep_votes > 0) {
-          party_vote_share <- data[i, "votes"] / total_non_rep_votes
-          votes_lost <- round(extra_votes * party_vote_share, 0)
-          adjust_votes[i] <- max(0, data[i, "votes"] - votes_lost)
-        } else {
-          adjust_votes[i] <- data[i, "votes"]
-        }
-      }
-    } else {
-      adjust_votes[i] <- data[i, "votes"]
-    }
-  }
-
-  return(adjust_votes)
-}
-
-adjust_vote1 <- adjust_function(adjusted_data)
-adjust_vote2 <- adjust_function2(adjusted_data)
-adjusted_data$adjust_vote1 <- adjust_vote1
-adjusted_data$adjust_vote2 <- adjust_vote2
-```
+@tbl-counter shows the summary of vote for the eleven swing states in 2020 if the difference in voting for Trump between the high DPC and low DPC eliminates.
+This equivalent to on average In 2020, the truth is that the Trump only won three of eleven swing states and he only won 232 electoral votes.
+However, with counterfactual votes, Trump will win five more states and hence bring him 52 more electoral votes.
+Therefore the total electoral votes for Trump will be 284.
+Since the cutoff to win is 270 votes, this indicate that Trump could re-elect if he can eliminate the disparities in DPC rate.
 

### 2023-12-15 outputs/paper/paper.qmd
Update the final paper and references

@@
 date-format: "long"
-abstract: "This study investigate the causal inference between COVID-19 and Donald Trump's loss during the 2020 US Federal Election using the data from MIT Election Data Science Lab and Johns Hopkins University CSSE.
-The main methodology used in this paper is Propensity Score Matching with with an exploratory analyzing the optimal treatment initially.
-The key finding suggest that the counties with a death per case rate exceeding the 0.4 quantile threshold shows a reduced voting preference for Trump.
-Additionally, this paper conducts a counterfactual analysis based on the treatment effect, indicating that Trump might have be secured to re-elect if the disparities in death per case rates were addressed.
-Future research should aim to refine the propensity score calculations by incorporating additional variables, such as the winning party in each county during the 2016 election."
+abstract: "This study investigates the causal inference between COVID-19 and Donald Trump's loss during the 2020 US Federal Election using the data from MIT Election Data Science Lab and Johns Hopkins University CSSE.
+The main methodology used in this paper is propensity score matching with an exploratory analysis of the optimal treatment.
+The key finding suggests that the counties with a death per case rate exceeding the 0.4 quantile threshold show a reduced voting preference for Trump.
+This paper also conducts a counterfactual analysis based on the treatment effect, indicating that Trump might have been secured to re-elect if the disparities in death per case rates were addressed.
+Future research should refine the propensity score calculations by incorporating additional variables, such as the winning party in each county during the 2016 election."
 format: pdf
@@
 
-At the beginning of 2020, US reported the first COVID-19 cases.
-This number grows to ten thousand within two months on March 19 and increase dramatically to 100K on March 27 [citation].
-At the same year, the U.S.
-Federal Presidential Election started off and Trump lost to Biden.
-As of Election Day, there was above 9 million cases and about 0.2 million deaths in US.
-When people talk about Trump's lost, people always connect it to the pandemic, saying that he "mishandled his greatest test" [citation] and 2016 Election was a historical accident [citation].
-However, there is another voice that COVID is not the only or the decisive factor attributing to his lost, his limits as a politican and wasted his advantage on the economy are also important explaining why he lost.
-This paper will investigate whether the is a causal effect between COVID and Trump's lost.
-If there is, whether it is strong enough to beath Trump at 2020.
+In early 2020, the United States encountered its first cases of COVID-19.
+The situation escalated rapidly, with reported cases reaching 10,000 by March 19 and soaring to 100,000 just eight days later [@WikiCOVIDUSA2023].
+Amidst this unfolding crisis, the U.S.
+Presidential Election was underway, ending with Trump's defeat to Biden.
+By Election Day on November 3, there were more than 9 million reported cases and about 200,000 deaths in the U.S.
+When talking about Trump's loss, the public refers to his mishandling of "his greatest test" [@Greenblatt2021] and believes that his win in the 2016 Election was a historical accident [@Bryant2020].
+Conversely, some arguments suggest that COVID-19 was not the sole or decisive factor in Trump's defeat; his limits as a political strategy and wasting his advantage on the economy also contributed to his loss.
+This paper will investigate the causal effect between COVID and Trump's loss, and if so, to what extent it influenced the 2020 election outcome.
 
-There are existing researches on this.
-Leaonardo et al.
-[citation] suggested that Trump could win if there was no COVID but it will be too naive to simply connect his lost to the COVID-19 infection rate.
-Their research highlighted that the COVID-19 had the most significant negative impacts on those urban areas where there was no stay-at-home orders, especially for those "swing" states.
-They also argue that the race diversity and education attainment are also important when analysing voting patterns for Trump.
-In addition to that, Marcus et al.
-[citation] suggested that the deaths per case is a more crucial indicator than infection rate when analysing the voting for Trump.
-However, they suggest the voting itself may not capture the impact of COVID as we do not know when did the voters make their decision, especially for early vote.
-Compared to these researches, Harold [citation] conduct read survey before and after the election to illustrate the impact of COVID pattern, they found that even though COVID has impacts on voting, it is not the dominant factor.
-Without COVID, Trump was also less likely to re-elect due to US highly polarized political landscape.
-Similarly, Shang et al.
-[citation] also conduct a survey and all the participants show natural or negative altitude.
-They suggest some social activity like "Black lives matter" may have impacts on voting.
-Furthermore, the
+Existing research on this subject provides varied perspectives.
+Baccini et al.
+[@Baccini2021CovidElection] suggested that Trump could win without COVID, but simply using the COVID-19 infection rate as the only factor contributing to his loss would be too naive.
+Socio-economic factors like high education attainment and race diversity are also unignorable.
+Besides, their research emphasized the significant impact of COVID-19 on those urban areas where there were no stay-at-home orders, particularly for the "swing" states.
+In addition, Noland & Zhang [@NolandZhang2021] suggested that the deaths per case is a more crucial metric than the infection rate when analyzing the impact of COVID-19 on voting for Trump.
+However, they highlight the challenge in assessing COVID-19's impact on voting, as we may not know when the voters made their decisions.
+In contrast, Clarke et al.
+[@Clarke2021CovidTrump] conducted actual surveys before and after the election, suggesting that COVID-19 impacts voting but not the dominant one.
+The U.S.'s highly polarized political landscape might be more significant.
+Similarly, Hart [@Hart2021CovidTrump] also conducted a survey to ask about people's attitudes toward Trump.
+All the participants show natural or negative possession.
+They also find that some social movements like "Murder of George Floyd" may impact voting.
 
-While all the above researches had investigated the correlation between COVID-19 and votings for Trump.
-Except conducting real surveys, the methods they used were to conduct a model and find the relevant variables in predicting the (change) vote for Trump.
-Then conduct counterfactual analysis to predict what will be the voting for Trump if there was no COVID or a certain amount reduce in COVID deaths.
-Instead using a similar way like those studies, this paper will introduce the method of Propensity Score Matching to find the treatment effect, to see whether those more impacted counties will have different voting patterns for Trump.
-However, since all counties experienced COVID-19, this paper will firstly find the optimal treatment groups.
-After that, I will also conduct a counterfactual analysis to see if Trump could re-elect.
+While all the above research investigated the correlation between COVID-19 and voting for Trump, previous studies predominantly employed models to identify variables influencing Trump's vote share and perform counterfactual analyses to estimate his voting performance in the absence of COVID-19 or reduction in deaths.
+Instead of using that method, this study employs propensity score matching to find the treatment effect and analyze whether counties more severely affected by COVID-19 exhibit distinct voting patterns.
+However, given all counties have experienced COVID-19, this paper will first find the optimal treatment group(s).
+Finally, I will conduct a counterfactual analysis to see if Trump could be re-elected with the treatment effect eliminated.
 
-There will be five parts in this paper.
-I will introduce the data used in this paper and present data suammries and visualization in [Data] part.
-After that, I will introduce the methods in this paper and the corresponding results in  [Methods] and [Results] part, respectively.
-Then the results will be interpreted and discussed in the [Discussion] part and conclude with limitations and drawbacks in [Conclusion] part..
+This paper will have five parts.
+I will introduce the data used in this paper and present data summaries and visualization in the @sec-data part.
+After that, I will explain the methods in this paper and the corresponding results in the @sec-methods and @sec-results parts, respectively.
+All the results will be interpreted and discussed in the @sec-discussion part, and I will conclude with limitations and drawbacks in the @sec-conclusion part.
 
-# Data
+# Data{#sec-data}
+
+The data in this paper is either downloaded directly or accessed via API.
+`R` [@R] will be the computer language used in the paper.
+Moreover, this paper will also use different R packages.
+The data is cleaned using the package `dplyr` [@dplyr], `stringr` [@stringr], `tidyverse` [@tidyverse], `janiotr` [@janitor], `tools` [@tools] and `tidyr` [@tidyr].
+In addition to that, `ggplot2` [@ggplot2], `RColorBrewer` [@RColorBrewer], `maps` [@maps], `mapdata` [@mapdata], `gridExtra` [@gridExtra] and `cowplot` [@cowplot] will be used later to make tables and plot graphs.
+Besides, to perform Propensity Score Matching, this paper will also employ `Matching` [@Matching].
 
@@
 
-The data used in this paper comprised five data sets from three different sources corresponding to different topics.
-The primary source is MIT Election Data Science Club which build open online data collections of the US Federal or Senate Election results, spanning from nation to county levels.
-The data extracted is called "County Presidential Election Returns 2000-2020" with about 70,000 rows containing the voting patterns for each candidate and party by county since 2000.
-Besides that, the data also indicate the types of voting, such as "EARLY VOTE" and "ELECTION DAY," for the same party in the a county.
-To analyze the voting patterns for Donald J.
-Trump, I only select the data for 2016 and 2020 US Federal Election for all counties and parties.
-Thus the filterd data set contains the number of votes through various ways for each county and party.
+This paper comprised five data sets from three different sources, each corresponding to different topics.
+The primary data is from the MIT Election Data Science Club [@MITElectionLab], which builds open online data collections of the U.S.
+Federal or Senate Election results from national to county levels.
+The data extracted is called "County Presidential Election Returns 2000-2020," [@MITElectionData2022] with about 70,000 rows containing the voting patterns for each candidate and party by county since 2000.
+Besides that, the data also indicate the types of voting, such as "EARLY VOTE" and "ELECTION DAY." To analyze the voting patterns for Donald J.
+Trump, I only filter the 2020 U.S.
+Federal Election data for all counties and parties.
 
-In addition to that, the study also uses the data regarding COVID-19 from the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University.
-CSSE collects and reports the local, national, and global multidimensional data, including medicine, health care, disaster response, etc.
+Additionally, the data on COVID-19 is taken from the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University [@jhu].
+CSSE collects and reports local, national, and global multidimensional data, including medicine, health care, disaster response, etc.
 During the pandemic, they collected the U.S.
-and international COVID cases and deaths and reported them on their GitHub, summarized by daily reports from April 12, 2020, to March 9, 2023.
-To access the impact of COVID on 2020 Election, this paper used the daily report on November 3, 2020 which is the election day of 2020 US Federal Election.
-The resulting data include the aggregates number of cases, deaths and recovers, as well as the incidence rate and case fatility ratio for each county by the election day.
+and international COVID cases and deaths, reporting them by daily summary on their GitHub [@CSSECOVID19Data].
+To examine the impact of COVID-19 on the 2020 Election as accurately as possible, this paper used the daily report on November 3, 2020, which is the election day of the 2020 U.S.
+Federal Election.
+The resulting data include the aggregate number of cases, deaths and recovers before Election Day and the incidence rate and case-fatality ratio for each county.
 
-Lastly, this paper also employ the socio-economic data from American Community Survey (ACS).
-ACS is a online open source database conducted by the U.S.
-Census Bureau, containing the various soci-economic factors in either geographic levels.
-The data sets extracted from ACS is 2020 five-year estimates of DP02, DP03 and DP05 which covers the social, economic and demographic characteristics in county level.
-Since there are thousands of variables, I will only use the variables which are found from my previous paper that are significant in predicting the COVID-19 mortality rate such as the mean household income, high-education attainment etc.
-The descriptions of all variables can be found on Table 1.
+This paper also incorporates the socioeconomic data from the American Community Survey (ACS) [@acs], which is an online open-source database conducted by the U.S.
+Census Bureau, containing the various socioeconomic factors at different geographic levels.
+The data sets extracted from ACS are 2020 five-year estimates of DP02, DP03 and DP05, covering the social, economic and demographic characteristics at the county level.
+Since there are thousands of variables, to ensure simplicity, I will only use the variables that are found to be significant in predicting the COVID-19 mortality rate from my previous paper [@CaoCOVID19Repo].
+The descriptions of all variables can be found on @tbl-descriptions.
 
-After having five data sets from three sources, they are merged into one big data by counties.
-Using the merged data, there are some new variables calculated by the exiting variables.
-In terms of election, I calculate the percentage of vote for each party (candidate) in both 2016 and 2020 Election, and calculate the corresponding change of percentage votes.
-I also create new variables indicating the winning party in both elections for each county.
-Additionally, to accurately compare the COVID cases and deaths across all counties, I transform the number of cases and deaths to infection and mortality rate by 10,000 citizens in each county.
-The final data now consists of 3107 rows with 36 columns.
-All the important variables are described in Table 1.
+After having five data sets, they are merged into one big data by county.
+Using the merged data, I calculate the percentage of votes for each party (candidate) in the 2020 election.
+In addition, I also create new dummy variables indicating the winning party in both elections for each county.
+Moreover, to accurately compare the COVID cases and deaths across all counties, I transform the number of cases and deaths to infection and mortality rate by 10,000 citizens in each county.
+Using the conclusion from Noland & Zhang's work [@NolandZhang2021], I also calculate each county's deaths per case rate.
+The final data consists of 3107 rows with 36 columns.
+All the essential variables are described in @tbl-descriptions.
 
@@
 #| label: tbl-descriptions
-data.frame(var = c("Income Percentile", "High-Education Attainment",
-                   "Private Insurance", "No Insurance", "White Population",
-                   "Black Population", "Males Population", "Infection Rate",
-                   "Mortality Rate", "Death per Case", "Votes for Democrat (2016)",
-                   "Votes for Democrat (2020)", "Votes for Republican (2016)",
-                   "Votes for Republican (2020)", "Change of Vote for Republican"),
-           code = c("income_pctile", "prop_high_education", "private_insurance",
+data.frame(var = c("income_pctile", "prop_high_education", "private_insurance",
            "no_insurance", "white_pct", "black_pct", "males", "infrate",
-           "mortrate", "dpc", "pct_vote_demo16", "pct_vote_demo", 
-           "pct_vote_rep16", "pct_vote_rep", "change_vote_rep"),
-           descriptions = c("The income percentile of each county",
-                            "The proportion of residents in a county having a at least bachelor degree",
+           "mortrate", "dpc", "pct_vote_demo", "pct_vote_rep"),
+           source = c("ACS", "ACS", "ACS", "ACS", "ACS", "ACS", "ACS", "JHU", 
+                      "JHU", "JHU", "MIT", "MIT"),
+           descriptions = c("The mean household income percentile of each county",
+                            "The proportion of local residents having a at least bachelor degree",
                             "The proportion of local residences having private insurance",
@@
                             "The proportion of Black population",
-                            "The proportion of Males population",
-                            "The COVID infection rate, calculated by number of case per 10,000 residences",
-                            "The COVID mortality rate, calculated by number of death per 10,000 residences",
+                            "The proportion of Males",
+                            "The COVID infection rate, calculated by cases per 10,000 residences",
+                            "The COVID mortality rate, calculated by deaths per 10,000 residences",
                             "The COVID death per 10,000 confirmed cases",
-                            "The percentage of votes for the Democrat in 2016",
                             "The percentage of votes for the Democrat in 2020",
-                            "The percentage of votes for the Republican in 2016",
-                            "The percentage of votes for the Republican in 2020",
-                            "The change of vote for the Republican from 2016 to 2020")) %>% 
+                            "The percentage of votes for the Republican in 2020")) %>% 
   rename(`Variables` = var,
-         `Coded name` = code,
+         `Source` = source,
          `Descriptions` = descriptions) %>% 
@@
          `Count` = n,
-         `Cases` = case,
+         `Average Cases` = case,
          `Infection Rate` = inf,
-         `Deaths` = deaths,
+         `Average Deaths` = deaths,
          `Mortality Rate` = mort,
@@
 
-@tbl-covid summarize the COVID-19 impacts and income levels between the counties voting for Democrat and Republican.
-Even though the Republican won about five sixths counties, there is a stark contrast where counties supporting Biden had nearly ten times the average number of COVID-19 cases and deaths compared to those supporting Trump.
-Despite a similar infection and mortality rates between the two groups, there is still a notable higher death-to-case rate in counties voting for the Republican.
-Furthermore, these counties also has a lower mean household income than the "Democrat-win" counties.
-These patterns suggest a potential correlation between the extent of COVID-19 impact and voting behavior in the 2020, where areas more severely affected by COVID-19 were less likely to vote for the Trump, vice-versa.
-This aligns with mainstream media and intuitive expectations.
+@tbl-covid compares the COVID-19 impacts and income levels across the counties won by each party.
+Even though Republicans won in approximately five-sixths of the counties, counties that supported Biden reported nearly ten times the average number of COVID-19 cases and deaths compared to those that supported Trump.
+Despite similar infection and mortality rates between the two groups, there is still a notable higher death-to-case rate in counties where Democrats won.
+These patterns suggest that the counties voting for Democrats are more severely impacted by COVID-19 and are more population-intensive than those voting for the Republicans.
+Even though only around 500 counties support Democrats, they are densely populated urban areas and wealthier, so Biden can have more electoral votes.
+We can verify this from income levels, where the counties voting Democrat has a higher mean income.
 
@@
   labs(title = "The relative share of votes between Republican and Democrat party",
-       fill = 'Share of Republican') +
+       fill = 'Republican share') +
   coord_quickmap() +
@@
 
-We can validate our guess from @fig-map which compares the relative ratio of votes for the two parties and the death to cases rate in maps.
-From the maps, it is clearly more counties had preference for the Republican but the counties with higher death per case rate have a smaller ratio of vote for the Republican and these counties are usually richer ones.
-The most significant example can be California and New York.
-In contrast, the states with less impacted by COVID, such as Utah, vote more for Republican.
+We can validate our guess from @fig-map, which compares the relative ratio of votes for the two parties and the death to cases rate in maps.
+From the maps, it is clear more counties had a preference for the Republicans, but the counties with higher death-per-case rates show a preference for Democrats.
+These counties, such as California and New York, are mostly more affluent.
+In contrast, the states that are less impacted by COVID-19, such as Utah, vote more for Republicans.
+The maps match our findings from @tbl-covid.
 
@@
 
-Given the insights from @tbl-covid and @fig-map, @fig-dpc-income further examine the impact of COVID and income levels on the voting behaviors for the two parties.
-@fig-dpc-income shows the correlation between the votes of the two parties to the death per case rate and income levels.
-The two graphs on the left indicates the share of votes whereas the right two graphs indicate the number of votes.
-The two parties are represented by different colors.
-In terms of share of votes, in general, there is a positive relationship between the DPC and votes for the Democrat but a negative one for the Republican.
-However, the counties with medium value of DPC is less sensitive to the voting for the two parties, compared to the "tail" counties which their support change significantly as DPC increases or decreases.
-Meanwhile, there is a quadratic patterns between the vote for the two parties and income, it seems that the Republican is in favor of low income counties, but as income increase, more counties vote for Bide not Trump.
+Given the insights from @tbl-covid and @fig-map, @fig-dpc-income delves deeper into the impact of COVID and income levels on the voting behaviours of the two parties.
+The left side of @fig-dpc-income displays the share of votes, while the right side focuses on the total number of votes, with each party represented by distinct colours.
+In terms of the percentage of votes, in general, there is a positive relationship between the DPC and votes for the Democrats but a negative one for the Republicans.
+However, the counties with a medium value of DPC seem less sensitive to the voting for the two parties, whereas the support of counties at the 'tails' shifts more significantly with changes in the DPC.
+Meanwhile, there are clear quadratic patterns between the vote for the two parties and income.
+The lower-income counties tend to favour Republicans, but as income increases, a greater number of counties lean towards voting for Biden rather than Trump.
 
-In addition, even though the average share of votes for Republican is always higher than 0.5, it does not necessarily mean that the Democrat lost in all counties.
-The reason behind this is, as mentioned in @tbl-covid, the counties which the Republican won is about five times as higher as the Democrat.
-Since Republican has more counties, it definitely has a higher average share of votes than Democrat.
-This pattern also indicate that Democrat usually won the densely populated, urban areas but Republican won in more sparsely populated, rural areas \[citation\].
-The high population density areas has more votes and explains why Biden beat Trump.
-We can verify this from the number of votes for the two parties, where the number of votes for Biden exceeds Trump as DPC and income level increases.
-This pattern is especially significant for the top 50% richest counties and the counties with DPC larger than 75 quantile.
+In addition, despite the average share of votes for Republicans being consistently above 0.5, it does not imply a universal loss for Democrats across all counties.
+As detailed in @tbl-covid, we know that the number of counties won by Republicans is approximately five times greater than those won by Democrats.
+With Republicans dominating more counties, they naturally have a higher average share of votes.
+This pattern also indicates that Democrats usually won the densely populated, urban areas, but Republicans won in more sparsely populated and rural regions [@Spencer2023ElectionFactCheck].
+The high population density areas have more votes, which explains why Biden beat Trump.
+We can verify this from the right side where the number of votes for Biden exceeds Trump as DPC and income level increases.
+This pattern is particularly pronounced in the wealthiest top 50% of counties and those with DPCs exceeding the 75th percentile.
 
 ```{r}
-data_population <- data %>% 
-  mutate(pop_pctile = ntile(total_population, 100)) %>% 
-  group_by(pop_pctile) %>% 
-  summarise(county_demo = sum(winning_party == "Democrat")/n(),
-            county_rep = sum(winning_party == "Republican")/n())
+#| fig-cap: The correlation between DPC rate with proportons of people with private insurance and without health insurance for the two parties
+#| label: fig-insurance
 
-data_population %>% 
-  ggplot(aes(x = pop_pctile)) +
-  geom_point(aes(y = county_rep, colour = "Republican"), alpha = 0.25) +  
-  geom_smooth(aes(y = county_rep, colour = "Republican"), 
-              method = "lm", formula = y ~ poly(x, 3), se = FALSE) +
-  geom_point(aes(y = county_demo, colour = "Democrat"), alpha = 0.25) + 
-  geom_smooth(aes(y = county_demo, colour = "Democrat"), 
-              method = "lm", formula = y ~ poly(x, 3), se = FALSE) +
+plot1 <- ggplot(data, aes(x = private_insurance)) +
+  #geom_point(aes(y = dpc, colour = winning_party), alpha = 0.25) +  
+  geom_smooth(aes(y = dpc, colour = winning_party), 
+              method = "lm", formula = y ~ poly(x, 1), se = FALSE) +
+  #ylim(1000, 4000) +
+  theme_minimal() +
+  scale_colour_manual(values = c("Republican" = "#FF9999", "Democrat" = "#9999FF"),
+                      name = "Winning Party:",  
+                      labels = c("Democrat", "Republican")) +
+  theme(legend.position = "none") + 
+  labs(x = "Private Insurance(%)", 
+       y = "DPC rate")
+
+plot2 <- ggplot(data, aes(x = no_insurance)) +
+  #geom_point(aes(y = dpc, colour = winning_party), alpha = 0.25) +  
+  geom_smooth(aes(y = dpc, colour = winning_party), 
+              method = "lm", formula = y ~ poly(x, 1), se = FALSE) +
+  #ylim(1000, 4000) +
+  theme_minimal() +
   scale_colour_manual(values = c("Republican" = "#FF9999", "Democrat" = "#9999FF"),
@@
                       labels = c("Democrat", "Republican")) +
-  theme_minimal() +
-  theme(legend.position = "bottom") + 
-  labs(x = "Population Percentiles", 
-       y = "Share of winning counties")
+  theme(legend.position = "none") + 
+  labs(x = "No Health Insurance(%)", 
+       y = "DPC rate")
+
+legend <- get_legend(plot1 + theme(legend.position = "bottom"))
+combined_plot <- plot_grid(plot1, plot2, ncol = 2)
+combined_plot_with_legend <- plot_grid(combined_plot, legend, ncol = 1, rel_heights = c(1, 0.1))
+
+print(combined_plot_with_legend)
 ```
 
+@fig-insurance shows the correlation between the DPC rate and the proportion of people with private and without health insurance for the two parties.
+Unsurprisingly, the graph reveals an inverse correlation between the DPC rate and private health insurance – as private insurance coverage increases, the DPC rate tends to decrease.
+Notably, counties that favoured the Democratic Party exhibit a higher proportion of individuals with private insurance compared to Republican-leaning counties.
+In contrast, in Republican-dominated counties, there is a pronounced positive correlation between the DPC rate and the lack of health insurance.
+However, it is relatively flat for Democratic counties.
+These findings suggest that counties voting Democratic generally have more comprehensive insurance coverage, which appears to mitigate the impact of DPC rates to a greater extent than in Republican counties.
+
 ```{r}
-boundaries <- c(0, 5000, 9000, 14000, 19000, 26000, 37000, 54000, 94000, 215000, Inf)
+#| fig-cap: The distribution of share of White population and high-education across different levels
+#| label: fig-bar
+data_we <- data %>%
+  mutate(white_pct_group = ntile(white_pct, 10),
+         prop_education_group = ntile(prop_higher_education, 10))
 
-# Classify the counties into percentiles
-data_population <- data %>% 
-  mutate(pop_pctile = cut(total_population, breaks = boundaries, include.lowest = TRUE, labels = FALSE))
+data_long <- data_we %>%
+  pivot_longer(cols = c(pct_vote_rep, pct_vote_demo), names_to = "party", values_to = "votes")
 
-# data_population %>% 
-#   filter(winning_party == "Republican") %>% 
-#   ggplot(aes(x = pop_pctile)) +
-#   geom_bar(aes(y = n()))
-# 
-# data_population %>% 
-#   group_by(pop_pctile, winning_party) %>% 
-#   summarise(n = n())
-# 
-# county_sum <- data_population %>%
-#   filter(winning_party == "Republican") %>%
-#   group_by(pop_pctile) %>%
-#   summarize(count = n())
+# Replace party names for plotting
+data_long$party <- recode(data_long$party, pct_vote_rep = "Republican", pct_vote_demo = "Democrat")
 
-# Plot the data
-# ggplot(county_sum, aes(x = pop_pctile, y = count)) +
-#   geom_bar(stat = "identity")
+# Plot for white_pct_group
+plot1 <- ggplot(data_long, aes(x = factor(white_pct_group), y = votes, group = party)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.8), aes(fill = party)) +
+  scale_fill_manual(values = c("Republican" = "#FF9999", "Democrat" = "#9999FF"),
+                    name = "Party") +
+  labs(x = "Ratio of White Pop. (Deciles)", y = "Mean Votes(%)", fill = "Party") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+# Plot for prop_education_group
+plot2 <- ggplot(data_long, aes(x = factor(prop_education_group), y = votes, group = party)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.8), aes(fill = party)) +
+  scale_fill_manual(values = c("Republican" = "#FF9999", "Democrat" = "#9999FF")) +
+  labs(x = "Ratio of High-Edu. (Deciles)", y = "Mean Votes(%)", fill = "Party") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+legend <- get_legend(plot1 + theme(legend.position = "bottom"))
+combined_plot <- plot_grid(plot1, plot2, ncol = 2)
+combined_plot_with_legend <- plot_grid(combined_plot, legend, ncol = 1, rel_heights = c(1, 0.1))
+
+print(combined_plot_with_legend)
 ```
 
-From four graphs in @fig-dpc-income, we can conclude that, in general, the higher the DPC rate, the lower the support for Trump but higher for Biden.
-Besides, the Democrat is in favor of richer voters, where as Republican is in favor of low income counties.
-It is direct to ask whether these two variables have interaction effect on the voting patterns?
-However, it will be difficult to visualize the interaction effect between two continuous factors.
-Instead, I will manually set a new categorical variable `high_income` to denote whether the county has a high income with certain cutoff.
-Then the difference of correlations of the two groups as death per case rate changes.
-I will set different cutoffs to fully test the interaction effect.
+@fig-bar compares the ratio of the White population (race diversity) and education levels for the two parties.
+Interestingly, it suggests that counties with higher racial diversity are less likely to vote for Biden, as opposed to the trend observed for Republican-leaning counties, where increased diversity aligns with increased support.
+This pattern implies that counties that predominantly vote for Democrats have a higher proportion of White residents.
+This may explain why they have relatively higher income levels.
+In terms of the ratio of people with high education attainment, since the average vote for Republicans is always higher than for Democrats, this may indicate that more educated people are likely to vote for Trump.
 
-```{r}
-#| echo: false
-#| message: false
-#| label: fig-boxplot
-#| fig-cap: Summary of income levels for counties voting for Democratic and Republican
+# Methods{#sec-meethods}
 
-data  %>%
-  ggplot(mapping = aes(x = winning_party, y = mean_household_income)) +
-  geom_boxplot() +
-  geom_jitter(alpha = 0.04, width = 0.3, height = 0) +
-  theme_minimal() +
-  labs(x = "Winning Party",
-       y = "Mean Household Income")
-```
-
-```{r, fig.width=10, fig.height=4.5}
-#| echo: false
-#| message: false
-#| label: fig-income_state
-#| fig-width: 10
-#| fig-height: 4.5
-#| fig-cap: Variation in COVID-19 Mortality Rates Across States, Categorized by Income Levels and Dominant Political Preferences
-
-# Find the wining data for each state
-data_win_party <- data %>%
-  group_by(state) %>%
-  summarise(dem_votes = sum(vote_demo),
-            rep_votes = sum(vote_rep), 
-            .groups = 'drop') %>% 
-  mutate(winning_party_state = ifelse(dem_votes > rep_votes, 'Democratic', 'Republican')) %>% 
-  dplyr::select(state, winning_party_state)
-# Calculate the average income
-average_income_by_state <- data %>%
-  group_by(state) %>%
-  summarise(average_income = mean(mean_household_income), .groups = 'drop')
-# Join the winning party and average income back to the original data
-data_win_party_income <- data %>%
-  left_join(data_win_party, by = "state") %>%
-  left_join(average_income_by_state, by = "state")
-# Arrange states by average income
-ordered_states <- data_win_party_income %>%
-  arrange(average_income) %>%
-  .$state %>%
-  unique()
-# Create a factor with levels based on income-ordered states for plotting
-data_win_party_income$state <- factor(data_win_party_income$state, levels = ordered_states)
-
-average_dpc <- mean(data_win_party_income$mortrate)
-
-ggplot(data_win_party_income, aes(x = state, y = mortrate, fill = winning_party_state)) +
-  geom_boxplot() +
-  geom_hline(yintercept = average_dpc, linetype = "dashed", color = "black") +
-  annotate("text", x = Inf, y = average_dpc, label = " National Average", hjust = 1, vjust = -0.5, color = "black", size = 3) +
-  scale_fill_manual(values = c('Democratic' = 'blue', 'Republican' = 'red')) +
-  labs(y = 'Death Rate', x = 'State (Ascending Income)', 
-       fill = "Winning Party") +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-```
-
-@fig-boxplot compares the distribution of county income levels for the two parties.
-In general, counties voting for the Democrats have higher income levels than the Republicans.
-In addition, we can observe that the counties are intensively concentrated around income levels \$50k to \$75k, compared to the Democrats, where the counties are approximately uniformly distributed at each income level.
-Furthermore, barely any county has at least 150k voting for the Republican party.
-Both @tbl-covid and @fig-boxplot indicates that the Democrat is in favour of wealthy counties but poorer counties for Republican.
-Consequently, the more affluent counties (states) usually have more electoral votes than the poorer ones; this may provide insights into why Trump lost.
-
-# Methods
-
-The objective in this paper is to conduct the causal inference between COVID-19 and Trump's loss during 2020, the main methods implemented in this study will be the Propensity Score Matching (PSM).
-After find the causal inference, this paper will also conduct a counterfactual analysis to see whether Trump can re-elect if there was no COVID.
+This paper aims to conduct the causal inference between COVID-19 and Trump's loss in 2020.
+The primary method implemented in this study will be the Propensity Score Matching (PSM).
+After finding the causal inference, this paper will also conduct a counterfactual analysis to see whether Trump can re-elect without COVID-19.
 
@@
 
-The treatment refers to the intervention or exposure that is being studied to understand its potential impact on the outcome.
-In other words, we implement certain interventions to a group of people but not for the rest.
-The group of people receiving the treatment is called the treatment group and control group for the rest.
-For instance, Jochen \[citation\] examined whether a new medical treatment called clampless off-pump coronary artery bypass (clampless OPCAB) will reduce the in-hospital mortality rate, compared to the traditional way.
-In their studies, the treatment is the new medical surgery and patients who took this surgery is the treatment group.
-They compared the mortality rates between the two groups and concluded that the new treatment can significantly lower the stroke and mortality rate.
+The treatment refers to the intervention or exposure being studied to understand its potential impact on the outcome [@WikiPropensity2023a].
+In other words, we implement specific interventions for a group of people but not for the rest.
+The group of people receiving the treatment is called the treatment group, and the control group for the rest.
+For instance, Austin [@Austin2011PropensityScore] examined the effectiveness of a new medical treatment called clampless off-pump coronary artery bypass (clampless OPCAB) to see if it can reduce the in-hospital mortality rate compared to the traditional method.
+Here, 'treatment' refers to the new surgical procedure, and patients who took this new surgery are the treatment group.
+They compared the mortality rates between the two groups and concluded that the new treatment could statistically lower the stroke and mortality rate.
 
-Normally, the treatment should be established before conduct the analysis, this aims to reduce the risk of p-hacking.
-However, the objective of this paper is to find whether Trump lost the last Election due to COVID-19, it may not be possible to state the treatment in advance as all the counties in US had experienced COVI-19.
-Therefore this paper will firstly perform exploratory analysis to firstly find the treatment group which has the most significant difference in voting for Trump.
-This paper will set a cutoff defining whether a county has a high or low death per case rate.
-In addition, from @fig-dpc-income, low and high income counties also seem to have different voting behaviors.
-Therefore the treatment in this paper will incorporate whether a county has a high DPC and a high income levels with certain cutoffs.
+Typically, the treatment should be defined before conducting the analysis, and this aims to reduce the risk of p-hacking [@Frost2023PHacking].
+However, given the objective of this paper, which is to find whether COVID-19 influenced Trump's loss in the last Election, pre-defining the treatment is challenging as all U.S.
+counties were affected by COVID-19.
+Therefore, this study begins with exploratory analysis to identify the treatment group exhibiting the most significant differences in voting patterns for Trump.
+I will establish specific cutoffs for classifying counties as having 'high' or 'low' death per case rates (DPC).
+In addition, from @fig-dpc-income, low and high-income counties also seem to have different voting behaviours.
+Therefore, both DPC and income will be considered when choosing the optimal treatment.
 
-However, we need to be very careful about the cutoffs for defining "high" in DPC and income is crucial as the risk of p-hacking or data dreging may raise.
-P-hacking means the manipulation of data analysis until we can find the statistically significant results [citation].
-If we simply choose one cutoff and find the significant results that match the common sense, it may raise the p-hacking risk and the analyses will be less convincing.
-Therefore, to make the analysis more comprehensive, instead of choosing a single cutoff, I will set a grid of cutoffs for each variable to see how the separate or combine treatment effects varies.
-For example, one treatment groups can be counties with death per case and income levels higher than 0.3 quantiles.
-By this way, we should be able to find the optimal treatment group and avoid the risk of p-hacking.
+Nevertheless, choosing cutoffs is critical, and we need to avoid p-hacking or data dredging risks.
+P-hacking means the manipulation of data analysis until we can find statistically significant results [@Frost2023PHacking].
+If we simply choose one cutoff and see the significant results that match common sense, it may raise the p-hacking risk, and the analyses will be less convincing.
+Therefore, we will employ a grid of cutoffs for each variable, examining how treatment effects vary.
+For example, one treatment group can be counties with death per case and income levels higher than 0.3 quantiles.
+In this way, we should be able to find the optimal treatment group and avoid the risk of p-hacking.
 
@@
 
-To estimate the treatment effects, randomized controlled trials (RCT) is the most ideal and the golden way to do this \[citation\].
-The reason is that RCTs is the only way that guarantees the equal distribution of known and unknown parameters.
-Therefore the outcomes between the treatment and control groups will not be confounded.
-However, conducting a RCT is nearly impractical and impossible.
-Back to the coronary artery bypass example \[citation\], you can not force a patient to take new or traditional treatment, the researchers can not control it and hence RCT is impossible.
-Instead, we can take observational studies to find the treatment effects.
+Randomized controlled trials (RCT) are regarded as the most ideal and the golden way to estimate the treatment effect [@Kuss2016PropensityScore].
+The reason is that RCTs are the only way that guarantee the equal distribution of known and unknown parameters.
+Therefore, the outcomes between the treatment and control groups will not be confounded.
+However, conducting an RCT is nearly impractical and impossible in many scenarios.
+Back to the coronary artery bypass example [@Austin2011PropensityScore], it's not ethically feasible to randomly assign patients to new or traditional treatments.
+Hence, conducting an RCT is nearly impossible.
 
-One method that can be applied to observational studies to find the treatment effect is Propensity Score Matching (PSM).
-Propensity score is the probability of an observation that receive the treatment based on the existing covariates using logistic regression.
-Using propensity scores have one main advantages which is dimension reduction [citation].
-Dimension reduction means we can describe an observation by only its propensity score instead of a list of covariates.
-Therefore, we can match two observations having similar propensity scores but from treatment and control groups separately.
-With this algorithm, we can have many matched pairs and the matching process is called propensity score matching.
-Here, we have another advantage of PSM which is design separation [citation], meaning that it can separate the covariates balancing and effects estimation.
-I will show the balance of covariates later.
-Finally, we can directly estimate their outcomes and solve the drawback that we can not have causal inference from observational studies.
-That is why it can also separate designs and this process is called propensity score matching.
+In such situations, implementing Propensity Score Matching (PSM) can be a more practical way to find the treatment effects.
+The propensity score is the probability of an observation that receives the treatment based on the existing covariates using logistic regression.
+Using propensity scores can reduce the dimension [@Zhao2021PropensityR], meaning that we can describe an observation to a single score instead of multiple covariates.
+In addition, another significant advantage of PSM is design separation [@Zhao2021PropensityR], meaning that PSM can separate the covariates balancing and effects estimation.
+This is especially beneficial as we can observe from @fig-insurance that the distribution of counties with relatively high DPC rates is imbalanced.
+Finally, we can match two observations having similar propensity scores but from treatment and control groups separately.
+Then, we will have many matched pairs and estimate the treatment effects.
+The matching process is called propensity score matching.
 
-In this paper, the propensity scores are used to estimate how likely a county having a high DPC or high income levels (or both) using multiple linear logistic regression. The model will be look like:
+In this paper, the propensity scores are used to estimate how likely a county is to receive treatment using multiple linear logistic regression. The equation will look like:
 
@@
 
-Using the model, I will calculate the propensity scores for each county and match the data with similar propensity scores.
-Notice that the values of propensity scores changes as the treatment changes.
-That said, I will calculate the propensity scores and match the data for number of treatment times.
-
 ## Counterfactual Analysis
 
-Using the PSM described previously, we can find the treatment effects under different settings of treatment groups.
-However, to conclude whether Trump was lost due to COVID, we need to perform a counterfactual analysis with re-calculating the votes for Trump of counties in the treatment group.
+Using the PSM described previously, we can find the treatment effects under different treatment group settings.
+However, to conclude whether Trump lost due to COVID-19, we need to perform a counterfactual analysis by re-calculating the votes for Trump in counties in the treatment group.
 
-The couterfactual analysis is particularly necessary for those "swing" states in 2020.
-In 2016 Federal Election, one key factory why Trump defeated Hillary is that he won seven out of eleven swing states.
-However, he only won three of them in 2020 [citaion (wikipedia)].
-Besides, if Donald Trump were able to hold onto three swing states which are Georgia, Arizona and Wisconsin where the average margin of Democrat is only about 0.3%, the result would have been a 269-269 electoral tie decided in the House of Representatives [citation (wikipedia)].
-Then the presidential election is left up to members of the House of Representatives and it is possible that Trump can win.
+The counterfactual analysis is particularly necessary for those "swing" states in 2020 [@WikiSwingState2023b].
+In the 2016 Federal Election, one key factor in Trump's defeat of Hillary is that he won seven out of eleven swing states.
+However, he only won three in 2020.
+Besides, if Donald Trump were able to hold onto three swing states which are Georgia, Arizona and Wisconsin, where the average margin of Democrats is only about 0.3%, the result would have been a 269-269 electoral tie decided in the House of Representatives.
+The presidential election is left up to members of the House of Representatives, and Trump could win.
 
-To find the voting patterns for Trump, especially for the "swing" states, I will follow the official election procedure and use the winners-takes-all rule.
-That said, I will re-calculate the votes for Trump in county level and summarize by county level.
-Then the party with higher total votes will take all the electoral votes in this state.
+To find the voting patterns for Trump, especially for the "swing" states, I will follow the official election procedure and use the winners-takes-all rule [@WikiWinnerTakeAll2023c].
+That said, I will re-calculate the votes for Trump at the county level and summarize by state level.
+Then, the party with the highest total votes will take all the electoral votes in this state.
 Eventually, I will calculate the total electoral votes for each party to see whether Trump could re-elect if the treatment effects diminish.
-Besides, I will also predict the votes if the death number can reduce 5%, 10%..., to see if a tiny decrease in deaths will make Trump to re-elect.
 
-# Results
+# Results{#sec-results}
 
@@
 @fig-dpc shows the correlation between votes for Trump and income levels, segmented into "High-DPC" and "Low-DPC" county groups at varying DPC cutoffs.
-For instance, a cutoff of 0.3 classifies counties with a DPC rate below this threshold as "Low DPC," and those above as "High DPC.".
-By comparing these groups across different cutoffs, we aim to identify the cutoff where the difference in voting for Trump across difference income levels is most pronounced.
-With that said, we can find that except cutoff 0.4, the two line converge as income increases (right tail).
-Only for the 0.4 cutoff, the lines of two groups converge at the middle-income but do not overlap and diverge at the tails.
+For instance, a cutoff of 0.3 classifies counties with a DPC rate below this threshold as "Low DPC" and those above as "High DPC." By comparing these groups across different cutoffs, we aim to identify the cutoff where the difference in voting for Trump across different income levels is most pronounced.
+That said, we can find that, except for the cutoff 0.4, the two lines converge as income increases (right tail).
+Only for the 0.4 cutoff, the lines of the two groups converge at the middle income but do not overlap and diverge at the tails.
 Additionally, regardless of the cutoff applied, there is a common pattern that the support for Trump increases in poorer counties up to around the eighth bin, which is approximately the 0.4 income quantile, and then decreases in wealthier counties.
@@
 
-Furthering on our analysis, @fig-income shows the correlation between Trump's vote share and the DPC rate but this time dividing counties based on income levels into "High Income" and "Low Income" groups.
+@fig-income shows the correlation between Trump's vote share and the DPC rate but this time dividing counties based on income levels into "High Income" and "Low Income" groups.
 Each subplot in this analysis corresponds to a different income cutoff.
 Compared to @fig-dpc, all curves, regardless of the cutoff, display a linear or near-linear relationship with voting patterns.
-As the income cutoff increases, the difference in voting behavior between the high and low-income counties becomes more distinct, especially at the 0.9 cutoff.
+As the income cutoff increases, the difference in voting behaviour between the high and low-income counties becomes more distinct, especially at the 0.9 cutoff.
 This suggests that in counties with higher incomes, voting patterns vary significantly from those in lower-income counties as the DPC rate changes.
-Therefore, the most significant pattern is observed at the 0.9 income cutoff, contradicted to the patterns identified in @fig-dpc.
+Therefore, the most significant pattern is observed at the 0.9 income cutoff, contradicting the patterns identified in @fig-dpc.
 
-So, should we include all the income levels and DPC in the settings of treatment?
-The answer is NO and we should only include the DPC.
-From @fig-dpc, a consistent quadratic relationship between income and voting patterns for Trump is evident, regardless of how the cutoff for defining high DPC is adjusted.
-This pattern demonstrates that poorer counties (in the first half of the income spectrum) may show a positive or a positive correlation with voting for Trump depending on the cutoff we choose, whereas richer counties will only exhibit a negative correlation.
-This quadratic pattern can be verifies from @fig-dpc-income, which illustrates a similar relationship between votes and income.
+So, a question may arise regarding whether we should include all the income and DPC in the settings of treatment.
+The answer is NO, and we should only include the DPC.
+From @fig-dpc, a consistent quadratic relationship between income and voting patterns for Trump is evident regardless of how the cutoff for defining high DPC is adjusted.
+This pattern demonstrates that poorer counties (in the first half of the income spectrum) may show a positive or a negative correlation with voting for Trump depending on the cutoff we choose.
+In contrast, more affluent counties will only exhibit a negative correlation no matter which cutoff we choose.
+This quadratic pattern can be verified from @fig-dpc-income, which illustrates a similar relationship between votes and income.
+
 Conversely, @fig-income presents a different narrative.
-Here, all regression lines, irrespective of the DPC cutoff, display a linear or nearly linear relationship.
-Besides, unlike the @fig-dpc, all the lines in @fig-income shows a negative slope, meaning that the choice of cutoff for "High DPC" group will not change the general voting patterns for Trump.
-The optimal cutoff we found previously, 0.4, is the one that the "High DPC" and "Low DPC" have the maximum difference.
-This finding can be also verified from @fig-dpc-income, which shows a monotonically decreasing cubic correlation between votes and DPC.
-The @fig-treatment shows a simple visualization of why is the case.
+All regression lines, irrespective of the DPC cutoff, display a linear or nearly linear relationship.
+We should exclude income because the choice of cutoff will significantly change the treatment effects; DPC is more "stable" than income.
+We can verify this from @fig-income, where each line shows a negative slope, meaning that the choice of cutoff for the "High DPC" group will not change the general voting patterns for Trump.
+Hence, including income would introduce extra variability in the treatment effects and increase the probability of p-hacking.
+The @fig-treatment shows a simple visualization of why this is the case.
 
@@
 #| label: fig-treatment
+#| fig-cap: The simple visualization regarding why excluding income from treatment
 # Splitting the data into two subsets
@@
 @fig-treatment shows a simple example when we test different cutoffs for DPC rate and income.
-From the graph, we can see that no matter which cutoff we set, the low DPC and high DPC counties always have the a negative value of slope.
+From the graph, we can see that no matter which cutoff we set, the low DPC and high DPC counties always have a negative value of slope.
 The only difference is the absolute values of the slope.
-However, for the income, we can see that for the cutoff 0.4 quantile, the low income and high income groups show a opposite patterns for voting for Trump.
-But when we increase the cutoff to 0.9 quantile.
-The two groups are both negatively related to the change.
-This indicates that it is not appropriate to incorporate income in treatment as the choice of cutoff will significant change the
+Conversely, the impact of income levels shows notable variability.
+For instance, at a cutoff of the 0.4 quantiles, low and high-income groups display opposite voting patterns for Trump.
+However, when we increase the cutoff to the 0.9 quantile, both income groups negatively correlate with Trump's vote share.
+This underscores the instability of income.
 
-no matter which cutoff we set, the two groups will always have opposite altitude for voting for Trump.
-Suppose we have enough samples, when we include income into the treatment, then treatment effect will always be more negative when we increase the cutoff.
-The treatment effect will be more and more significant when we increase the cutoff.
-However, the number of counties in treatment group will decrease and hence it will be unnecessary to have income into treatment settings.
-
-From @fig-dpc, we already verify that cutoff 0.4 is the optimal threshold for defining high and low DPC group, therefore the treatment group will be the counties having DPC rate larger than 0.4 quantile.
-Furthermore, to balance the effect of income, I will incorporate income into the ways to calculate the propensity scores for each county.
+From @fig-dpc, we have determined that a cutoff at the 0.4 quantile is the most effective threshold for distinguishing between high and low Death Per Case (DPC) rate groups.
+Therefore, in our study, the treatment group is defined as those counties where the DPC rate exceeds the 0.4 quantile.
+Moreover, while income is not considered for determining the treatment and control groups due to its variability, its influence is still significant and shows an imbalance from @fig-dpc.
+To account for this, income will be integrated into the calculation of propensity scores for each county.
 
@@
 
-In my previous paper, I already show that proportion of residence with at least a bachelor degree, income, proportion of people with private insurance and withou health insurance, ratio of males, ratio of black and black residence are statistically in predicting the probability of a county has a high mortality rate or not.
+In my previous paper, I already showed that the proportion of residents with at least a bachelor's degree, income, the proportion of people with private insurance and without health insurance, the ratio of males, ratio of black and black residents are statistically in predicting the probability of a county has a high mortality rate or not.
 In this paper, I will directly use the results and only consider these variables when predicting the propensity scores.
@@
 The above equation shows the logistic regression model to predict the propensity score for each county.
-It seems that males is the most critical factor when predict the propensity score.
+Males seem to be the most critical factor when predicting the propensity score.
 Perhaps the majority of COVID deaths are males.
-Interestingly, the counties with a higher income levels will have a higher death per case rate.
+Interestingly, the counties with a higher income level will have a higher death per case rate.
 This may explain why richer counties are less likely to vote for Trump.
@@
 @tbl-psm summarize the results of Propensity Score Matching.
-The treatment effect is about -0.18 means that the counties with DPC rate higher than 0.4 will in average vote 0.018 less for Trump, compared to the counties with lower DPC rate.
-This value is statistically significant as its p value is about 0.02 which is lower than 0.05.
-In addition, there are 1869 treatment observations in both original and matched data, meaning that there some counties in control group that are matched more than once.
+The treatment effect of about -0.018 means that the counties with a DPC rate higher than 0.4 will, on average, vote 0.018 less for Trump compared to the counties with a lower DPC rate.
+This value is statistically significant as its p-value is about 0.02, lower than 0.05.
+In addition, there are 1869 treatment observations in both original and matched data, meaning that some counties in the control group are matched more than once.
 
@@
 
-@tbl-balance summarize each variable's mean values in the control group and standard mean difference before and after the matching.
-We can see that the mean values in the control group approximately remain same after the matching.
-Besides, there is a significant decrease in the standard mean difference for each variable.
-This indicates that the Propensity Score Matching reduced the imbalance between the treatment and control group considerablely.
+@tbl-balance summarizes each variable's mean values in the control group and standard mean difference before and after the matching.
+Notably, there is a significant decrease in the standard mean difference for each variable.
+This indicates that the Propensity Score Matching effectively reduced the imbalance between the treatment and control groups.
+Besides, we can see that the mean values in the control group approximately remain the same after the matching, enhancing the reliability of the matching process and suggesting that PSM has successfully aligned the groups based on the observed covariates.
 
@@
 
-@tbl-counter shows the summary of vote for the eleven swing states in 2020 if the difference in voting for Trump between the high DPC and low DPC eliminates.
-This equivalent to on average In 2020, the truth is that the Trump only won three of eleven swing states and he only won 232 electoral votes.
-However, with counterfactual votes, Trump will win five more states and hence bring him 52 more electoral votes.
-Therefore the total electoral votes for Trump will be 284.
-Since the cutoff to win is 270 votes, this indicate that Trump could re-elect if he can eliminate the disparities in DPC rate.
+@tbl-counter provides a comparative summary of the actual and counterfactual voting results in eleven swing states during the 2020 U.S.
+Presidential election.
+In reality, Trump only won three swing states with 232 electoral votes.
+However, the counterfactual analysis suggests that Trump would have won five additional states if the DPC rate disparities were eliminated, bringing him an extra 52 electoral votes.
+This would bring his total to 284 electoral votes, surpassing the required 270-vote threshold for reelection.
+Therefore, if Trump can eliminate the extra deaths, he could be re-elected.
 
-# Discussion
+# Discussion{#sec-discussion}
 
 ## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
 
@@
 
-Weaknesses and next steps should also be included.
+# Conclusion{#sec-conclusion}
 

### 2023-12-19 scripts/02-data_cleaning.R
Update the cleaned data sets

@@ -171,8 +171,8 @@ election_data_2016_clean <- election_data_2016 %>%
   group_by(state, county, fips, party, total_votes) %>%
   summarise(votes = sum(votes), .groups = 'drop') %>%
   pivot_wider(names_from = party, values_from = votes) %>% 
-  mutate(vote_demo = Democrat,
-         vote_rep = Republican,
+  mutate(vote_demo16 = Democrat,
+         vote_rep16 = Republican,
          vote_other = Other,
          pct_vote_demo16 = Democrat/total_votes,
          pct_vote_rep16 = Republican/total_votes,
@@ -180,8 +180,8 @@ election_data_2016_clean <- election_data_2016 %>%
                 ~replace(., is.na(.), 0)),
          rep_won16 = ifelse(pct_vote_rep16 > pct_vote_demo16, 1, 0),
          demo_won16 = ifelse(pct_vote_rep16 < pct_vote_demo16, 1, 0)) %>% 
-  dplyr::select(state, county, fips, pct_vote_demo16, pct_vote_rep16, rep_won16,
-                demo_won16)
+  dplyr::select(state, county, fips, pct_vote_demo16, pct_vote_rep16, vote_demo16,
+                vote_rep16, rep_won16, demo_won16)
 
 ## elction_data_clean combines the election data for 2016 and 2020
 election_data_clean <- election_data_2020_clean %>% 

### 2023-12-19 outputs/paper/paper.qmd
Update the final paper

@@
 ---
-title: "An Observational Study to COVID-19 and 2020 US Presidential Election: Counties with High Deaths per Case Rate Showed Diminished Support For Trump"
+title: "COVID-19 and the 2020 US Presidential Election: Counties with Extra COVID-19 Deaths Showed Less Support For Trump"
 title-block-banner: true
 author: "Yiliu Cao"
-thanks: "Code and data from this analysis are available at: https://github.com/yiliuc/covid_and_trump_loss.git"
+thanks: "Code and data from this analysis are available at: https://github.com/yiliuc/covid_and_trump_loss"
 date: "today"
 date-format: "long"
-abstract: "This study investigates the causal inference between COVID-19 and Donald Trump's loss during the 2020 US Federal Election using the data from MIT Election Data Science Lab and Johns Hopkins University CSSE.
+abstract: "This study investigates the relationship between COVID-19 and Donald Trump's loss during the 2020 US Federal Election using the data from MIT Election Data Science Lab and Johns Hopkins University CSSE.
 The main methodology used in this paper is propensity score matching with an exploratory analysis of the optimal treatment.
-The key finding suggests that the counties with a death per case rate exceeding the 0.4 quantile threshold show a reduced voting preference for Trump.
-This paper also conducts a counterfactual analysis based on the treatment effect, indicating that Trump might have been secured to re-elect if the disparities in death per case rates were addressed.
-Future research should refine the propensity score calculations by incorporating additional variables, such as the winning party in each county during the 2016 election."
+The key finding is that counties with a death per case rate exceeding the 0.4 quantile threshold show a reduced voting preference for Trump.
+This paper also conducts a counterfactual analysis based on the treatment effect, indicating that Trump might have been re-elected if the disparities of extra deaths were addressed.
+Future research should incorporate the neighborhood effects on voting between different counties, and validate the causal inference used."
 format: pdf
@@
 library(ggplot2)
+library(Matching)
+library(kableExtra)
 ```
@@
 Presidential Election was underway, ending with Trump's defeat to Biden.
-By Election Day on November 3, there were more than 9 million reported cases and about 200,000 deaths in the U.S.
+By Election Day on November 3, there were more than nine million reported cases and about 200,000 deaths in the U.S.
 When talking about Trump's loss, the public refers to his mishandling of "his greatest test" [@Greenblatt2021] and believes that his win in the 2016 Election was a historical accident [@Bryant2020].
@@
 Besides, their research emphasized the significant impact of COVID-19 on those urban areas where there were no stay-at-home orders, particularly for the "swing" states.
-In addition, Noland & Zhang [@NolandZhang2021] suggested that the deaths per case is a more crucial metric than the infection rate when analyzing the impact of COVID-19 on voting for Trump.
-However, they highlight the challenge in assessing COVID-19's impact on voting, as we may not know when the voters made their decisions.
-In contrast, Clarke et al.
-[@Clarke2021CovidTrump] conducted actual surveys before and after the election, suggesting that COVID-19 impacts voting but not the dominant one.
-The U.S.'s highly polarized political landscape might be more significant.
-Similarly, Hart [@Hart2021CovidTrump] also conducted a survey to ask about people's attitudes toward Trump.
+In addition, Noland and Zhang (2021) suggested that the deaths per case is a more crucial metric than the infection rate when analyzing the impact of COVID-19 on voting for Trump.
+However, they highlight the challenge in assessing COVID-19's impact on voting, as we may not know when the voters made their decision.
+In contrast, Clarke, Stewart, and Ho (2021) conducted surveys before and after the election, suggesting that COVID-19 impacts voting but not the dominant one.
+The U.S.'s polarized political landscape might be more significant.
+Similarly, Hart (2021) also conducted a survey to ask about people's attitudes toward Trump.
 All the participants show natural or negative possession.
@@
 However, given all counties have experienced COVID-19, this paper will first find the optimal treatment group(s).
-Finally, I will conduct a counterfactual analysis to see if Trump could be re-elected with the treatment effect eliminated.
-
-This paper will have five parts.
-I will introduce the data used in this paper and present data summaries and visualization in the @sec-data part.
-After that, I will explain the methods in this paper and the corresponding results in the @sec-methods and @sec-results parts, respectively.
-All the results will be interpreted and discussed in the @sec-discussion part, and I will conclude with limitations and drawbacks in the @sec-conclusion part.
+Finally,wewill conduct a counterfactual analysis to see if Trump could be re-elected with the treatment effect eliminated.
+
+This paper will have five parts.wewill introduce the data used in this paper and present data summaries and visualization in the @sec-data.
+After that,wewill explain the methods in this paper and the corresponding results in the @sec-methods and @sec-results, respectively.
+All the results will be interpreted and discussed in the @sec-discussion, andwewill conclude with limitations and drawbacks in the @sec-conclusion.
 
@@
 `R` [@R] will be the computer language used in the paper.
-Moreover, this paper will also use different R packages.
-The data is cleaned using the package `dplyr` [@dplyr], `stringr` [@stringr], `tidyverse` [@tidyverse], `janiotr` [@janitor], `tools` [@tools] and `tidyr` [@tidyr].
+The data is cleaned using the package `dplyr` [@dplyr], `stringr` [@stringr], `tidyverse` [@tidyverse], `janiotr` [@janitor], `tools` [@R] and `tidyr` [@tidyr].
 In addition to that, `ggplot2` [@ggplot2], `RColorBrewer` [@RColorBrewer], `maps` [@maps], `mapdata` [@mapdata], `gridExtra` [@gridExtra] and `cowplot` [@cowplot] will be used later to make tables and plot graphs.
-Besides, to perform Propensity Score Matching, this paper will also employ `Matching` [@Matching].
+To perform propensity score matching, this paper will also employ `Matching` [@Matching].
 
@@
 The data extracted is called "County Presidential Election Returns 2000-2020," [@MITElectionData2022] with about 70,000 rows containing the voting patterns for each candidate and party by county since 2000.
-Besides that, the data also indicate the types of voting, such as "EARLY VOTE" and "ELECTION DAY." To analyze the voting patterns for Donald J.
-Trump, I only filter the 2020 U.S.
+Besides that, the data also indicate the types of voting, such as "EARLY VOTE" and "ELECTION DAY." To analyze the voting patterns for Trump,weonly filter the 2020 U.S.
 Federal Election data for all counties and parties.
@@
 The data sets extracted from ACS are 2020 five-year estimates of DP02, DP03 and DP05, covering the social, economic and demographic characteristics at the county level.
-Since there are thousands of variables, to ensure simplicity, I will only use the variables that are found to be significant in predicting the COVID-19 mortality rate from my previous paper [@CaoCOVID19Repo].
+Since there are thousands of variables, to ensure simplicity,wewill only use the variables that are found to be significant in predicting the COVID-19 mortality rate from my previous paper [@CaoCOVID19Repo].
 The descriptions of all variables can be found on @tbl-descriptions.
 
-After having five data sets, they are merged into one big data by county.
-Using the merged data, I calculate the percentage of votes for each party (candidate) in the 2020 election.
-In addition, I also create new dummy variables indicating the winning party in both elections for each county.
-Moreover, to accurately compare the COVID cases and deaths across all counties, I transform the number of cases and deaths to infection and mortality rate by 10,000 citizens in each county.
-Using the conclusion from Noland & Zhang's work [@NolandZhang2021], I also calculate each county's deaths per case rate.
+The five data sets are merged into one big data by county.
+Using the merged data,wecalculate the percentage of votes for each party (candidate) in the 2020 election.
+In addition,wealso create new dummy variables indicating the winning party in both elections for each county.
+Moreover, to accurately compare the COVID cases and deaths across all counties,wetransform the number of cases and deaths to infection and mortality rate by 10,000 citizens in each county.
+Using the conclusion from Noland & Zhang's work [@NolandZhang2021],wealso calculate each county's deaths per case rate.
 The final data consists of 3107 rows with 36 columns.
@@
   summarise(n = n(),
-            case = mean(cases),
-            inf = mean(infrate),
-            deaths = mean(deaths),
-            mort = mean(mortrate),
-            dpc = mean(dpc),
-            income = mean(mean_household_income)) %>% 
+            case = round(mean(cases), 0),
+            inf = round(mean(infrate), 0),
+            deaths = round(mean(deaths), 0),
+            mort = round(mean(mortrate), 0),
+            dpc = round(mean(dpc), 0),
+            income = round(mean(mean_household_income), 0)) %>% 
   rename(`Winning Party` = winning_party,
@@
          `Mortality Rate` = mort,
-         `DPC` = dpc,
+         `Death per Case (DPC)` = dpc,
          `Income` = income) %>% 
@@
 #| label: fig-map
-#| fig-cap: The ratio of votes for the Republican and the infection rate per 100k in each county
+#| fig-cap: The ratio of Republican votes and the death per case (DPC) rate in each county
 
@@
 
-# Methods{#sec-meethods}
+# Methods{#sec-methods}
 
@@
 counties were affected by COVID-19.
-Therefore, this study begins with exploratory analysis to identify the treatment group exhibiting the most significant differences in voting patterns for Trump.
-I will establish specific cutoffs for classifying counties as having 'high' or 'low' death per case rates (DPC).
+Therefore, this study begins with exploratory analysis to identify the treatment group exhibiting the most significant differences in voting patterns for Trump.wewill establish specific cutoffs for classifying counties as having 'high' or 'low' death per case rates (DPC).
 In addition, from @fig-dpc-income, low and high-income counties also seem to have different voting behaviours.
@@
 
-Randomized controlled trials (RCT) are regarded as the most ideal and the golden way to estimate the treatment effect [@Kuss2016PropensityScore].
+Randomized controlled trials (RCT) are regarded as the most ideal way to estimate the treatment effect [@Kuss2016PropensityScore].
 The reason is that RCTs are the only way that guarantee the equal distribution of known and unknown parameters.
 Therefore, the outcomes between the treatment and control groups will not be confounded.
-However, conducting an RCT is nearly impractical and impossible in many scenarios.
+However, conducting an RCT is difficult in many scenarios.
 Back to the coronary artery bypass example [@Austin2011PropensityScore], it's not ethically feasible to randomly assign patients to new or traditional treatments.
@@
 In such situations, implementing Propensity Score Matching (PSM) can be a more practical way to find the treatment effects.
-The propensity score is the probability of an observation that receives the treatment based on the existing covariates using logistic regression.
-Using propensity scores can reduce the dimension [@Zhao2021PropensityR], meaning that we can describe an observation to a single score instead of multiple covariates.
-In addition, another significant advantage of PSM is design separation [@Zhao2021PropensityR], meaning that PSM can separate the covariates balancing and effects estimation.
+The propensity score is the probability an observation would have been treated based on the existing covariates using logistic regression.
+Using propensity scores reduces the dimension [@Zhao2021PropensityR], meaning that we can describe an observation to a single score instead of multiple covariates.
+Another significant advantage of PSM is design separation [@Zhao2021PropensityR], meaning that PSM can separate the covariates balancing and effects estimation.
 This is especially beneficial as we can observe from @fig-insurance that the distribution of counties with relatively high DPC rates is imbalanced.
-Finally, we can match two observations having similar propensity scores but from treatment and control groups separately.
-Then, we will have many matched pairs and estimate the treatment effects.
+Finally, we can match two observations with similar propensity scores, but from treatment and control groups, separately.
+Then, we will have many matched pairs and estimate the treatment effect.
 The matching process is called propensity score matching.
 
-In this paper, the propensity scores are used to estimate how likely a county is to receive treatment using multiple linear logistic regression. The equation will look like:
-
-```{=tex}
-\begin{align*}
-\log\left(\frac{P(T=1|X)}{1 - P(T=1|X)}\right) = \beta_0 + \beta_1 X_1 +\cdots +\beta_p X_p
-\end{align*}
-```
-where:\
-* $P(T=1|X)$ is the probability of a county receiving treatment\
-* $X_j\:(1\leq j\leq p)$ are the covariates\
+In this paper, propensity scores are used to estimate how likely a county is to receive treatment using multiple linear logistic regression. The equation is:
+
+$$
+\begin{aligned}
+\mbox{Pr}(y_i = 1) = & \text{logit}^{-1}(\beta_0 + \beta_1 \times \text{prop\_higher\_education}\\
+& + \beta_2 \times \text{income\_pctile} + \beta_3 \times \text{no\_insurance} \\
+& + \beta_4 \times \text{private\_insurance} + \beta_5 \times \text{males} \\
+& + \beta_6 \times \text{white\_pct} + \beta_7 \times \text{black\_pct})
+\end{aligned}
+$$
+where:
+
+* $\mbox{Pr}(y_i = 1)$ is the probability that a county has a "High" DPC rate
 * $\beta_0$ is the intercept
-* $\beta_j\:(1\leq j\leq p)$ is the corresponding coefficients of each covariate\
+* $\beta_j\:(1\leq j\leq 7)$ is the corresponding coefficients for seven predictors
 
@@
 The counterfactual analysis is particularly necessary for those "swing" states in 2020 [@WikiSwingState2023b].
-In the 2016 Federal Election, one key factor in Trump's defeat of Hillary is that he won seven out of eleven swing states.
+In the 2016 Election, Trump won seven out of eleven swing states.
 However, he only won three in 2020.
-Besides, if Donald Trump were able to hold onto three swing states which are Georgia, Arizona and Wisconsin, where the average margin of Democrats is only about 0.3%, the result would have been a 269-269 electoral tie decided in the House of Representatives.
+If Trump had been able to hold onto three swing states Georgia, Arizona and Wisconsin where the average margin of Democrats is only about 0.3%, the result would have been a 269-269 electoral tie.
 The presidential election is left up to members of the House of Representatives, and Trump could win.
 
-To find the voting patterns for Trump, especially for the "swing" states, I will follow the official election procedure and use the winners-takes-all rule [@WikiWinnerTakeAll2023c].
-That said, I will re-calculate the votes for Trump at the county level and summarize by state level.
+To find the voting patterns for Trump, especially for the "swing" states, we will follow the official election procedure and use the winners-takes-all rule [@WikiWinnerTakeAll2023c].
+That said, we will re-calculate the votes for Trump at the county level and summarize by state level.
 Then, the party with the highest total votes will take all the electoral votes in this state.
-Eventually, I will calculate the total electoral votes for each party to see whether Trump could re-elect if the treatment effects diminish.
+Eventually, we will calculate the total electoral votes for each party to see whether Trump could re-elect if the treatment effects diminish.
 
@@
 
-So, a question may arise regarding whether we should include all the income and DPC in the settings of treatment.
-The answer is NO, and we should only include the DPC.
-From @fig-dpc, a consistent quadratic relationship between income and voting patterns for Trump is evident regardless of how the cutoff for defining high DPC is adjusted.
+A question may arise regarding whether we should include all the income and DPC in the settings of treatment.
+We include only the DPC.
+Because, from @fig-dpc, a consistent quadratic relationship between income and voting patterns for Trump is evident regardless of how the cutoff for defining high DPC is adjusted.
 This pattern demonstrates that poorer counties (in the first half of the income spectrum) may show a positive or a negative correlation with voting for Trump depending on the cutoff we choose.
@@
 Hence, including income would introduce extra variability in the treatment effects and increase the probability of p-hacking.
-The @fig-treatment shows a simple visualization of why this is the case.
 
@@
 
-In my previous paper, I already showed that the proportion of residents with at least a bachelor's degree, income, the proportion of people with private insurance and without health insurance, the ratio of males, ratio of black and black residents are statistically in predicting the probability of a county has a high mortality rate or not.
-In this paper, I will directly use the results and only consider these variables when predicting the propensity scores.
+Cao (2023) already showed that the proportion of residents with at least a bachelor's degree, income, the proportion of people with private insurance and without health insurance, the ratio of males, ratio of black and black residents are statistically in predicting the probability of a county has a high mortality rate or not.
+Here,we will directly use the results and only consider these variables when predicting the propensity scores.
 
@@
 library(dplyr)
-data <- read.csv(here("outputs/data/merged_data.csv"))
-
 data_te <- data %>% 
@@
 data_te$prop_score <- predict(highinf_model, type = "response")
-
-library(Matching)
-library(kableExtra)
-# Propensity score matching
 rr <- Match(Y = data_te$pct_vote_rep, Tr = data_te$treatment,
             X = data_te$prop_score, M = 1)
-
-data.frame(estimate = -0.018342,
-           p = 0.021404,
-           on = 3115,
-           ot = 1869,
-           mo=1869,
-           mou = 5415) %>% 
+# summary(rr)
+extract_match_summary <- function(rr) {
+    summary_output <- capture.output(summary(rr))
+    estimate <- as.numeric(sub("Estimate...\\s*(-?\\d+\\.?\\d*e?-?\\d*)", "\\1", summary_output[2]))
+    p_value <- round(as.numeric(sub("p.val......\\s*(\\d+\\.?\\d*e?-?\\d*)", "\\1", summary_output[5])), 4)
+    original_num_obs <- as.numeric(sub("Original number of observations..............\\s*(\\d+)", "\\1", summary_output[7]))
+    original_num_treated <- as.numeric(sub("Original number of treated obs...............\\s*(\\d+)", "\\1", summary_output[8]))
+    matched_num_obs <- as.numeric(sub("Matched number of observations...............\\s*(\\d+)", "\\1", summary_output[9]))
+    matched_num_obs_unweighted <- as.numeric(sub("Matched number of observations  \\(unweighted\\).\\s*(\\d+)", "\\1", summary_output[10]))
+    return(list(
+        estimate = estimate,
+        p = p_value,
+        on = original_num_obs,
+        ot = original_num_treated,
+        mo = matched_num_obs,
+        mou = matched_num_obs_unweighted
+    ))
+}
+
+result <- extract_match_summary(rr)
+
+data.frame(estimate = result$estimate,
+           p = result$p,
+           on = result$on,
+           ot = result$ot,
+           mo = result$mo,
+           mou = result$mou) %>% 
   rename(`Treatment Effect` = estimate,
@@
 library(kableExtra)
+
+# mb  <- MatchBalance(treatment ~ prop_higher_education  + pctile +
+#                        no_insurance + private_insurance + males +
+#                        white_pct + black_pct, data=data_te, match.out=rr, nboots=10)
 
@@
 ```{r}
-#| tbl-cap: The summary of counterfactual votes for the eleven swing staes in 2020
+#| tbl-cap: The summary of counterfactual votes for the eleven swing states in 2020
 #| label: tbl-counter
@@
 
+```{r}
+#| include: false
+library(dplyr)
+data <- read.csv(here("outputs/data/merged_data.csv"))
+
+data_te <- data %>% 
+  filter(dpc > quantile(dpc, 0.4)) %>% 
+  mutate(rep_won20 = ifelse(pct_vote_rep > pct_vote_demo, 1, 0),
+         demo_won20 = ifelse(pct_vote_rep < pct_vote_demo, 1, 0)) %>% 
+  mutate(treatment = rep_won20)
+
+highinf_model <- glm(treatment ~ prop_higher_education  + pctile +
+                       no_insurance + private_insurance + males +
+                       white_pct + black_pct, family = binomial(),
+                      data = data_te)
+data_te$prop_score <- predict(highinf_model, type = "response")
+
+library(Matching)
+library(kableExtra)
+# Propensity score matching
+rr <- Match(Y = data_te$pct_vote_rep, Tr = data_te$treatment,
+            X = data_te$prop_score, M = 1)
+summary(rr)
+```
+
 # Discussion{#sec-discussion}
 
-## First discussion point {#sec-first-point}
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
+## Defining the treatment is the most crucial part in this study
+In an observational study, especially for conducting the causal inference, the treatment should always be settled before the analysis as it is effective to reduce the risk of p-hacking.
+If we want to focus whether the COVID-19 has generally influenced the voting for Trump compared to 2016, then we can take the voting for each county on both 2016 and 2020 and the treatment would simply be whether this county had COVID-19.
+However, this paper aims to investigate whether the extra COVID-19 death during the pandemic has influenced the voting for Trump.
+As all the U.S.
+counties had experienced COVID-19, it is necessary to manually define the cutoff of "high" extra COVID-19 deaths.
+Therefore, this study starts with an exploratory analysis in finding the optimal treatment group.
+
+From @tbl-covid and @fig-dpc-income, it seems that a county's COVID death per case rate (DPC) and income levels are the two factors that are most related to the voting for Trump.
+Therefore the treatment will only incorporate these two factors.
+However, even though the main objective in defining the cutoff is to maximize the difference between the treatment and control group, we can observe from @fig-dpc and @fig-income that the effect of income on the voting for Trump is highly dependent on the choice of cutoff as it appears a quadratic relationship.
+A quadratic pattern is not desired as different cutoffs will bring large differences on the treatment effects.
+The two groups may even show opposite relation (see @fig-treatment) or the same sign but with a high cutoff.
+However, unlike the income, DPC shows a general cubic relationship to vote of Trump and hence the choice of cutoff will not generate huge different treatment effects.
+We can verify this from @fig-income where no matter how the cutoff of "high" income changes, there will always have a linear or at least a closely linear relationship between them.
+The change of cutoff will not bring so dramatic change of treatment effects (see @fig-treatment).
+Furthermore, we can observe that the "high" and "low" DPC rate counties behave most differently with a cutoff at 0.4 quantile in @fig-income, therefore the optimal treatment group is counties with a DPC rate higher than the 0.4 quantile.
+
+```{r}
+#| label: tbl-income
+# show TE with different income levels
+generate_rr_summaries <- function(data, cutoffs, treatment_var) {
+  rr_list <- list()
+  for (cutoff in cutoffs) {
+    data_te <- data %>%
+      mutate(high_dpc = ifelse(dpc > quantile(dpc, cutoff), 1, 0),
+             high_income = ifelse(mean_household_income > quantile(mean_household_income, cutoff), 1, 0))
+    data_te$treatment <- data_te[[treatment_var]]
+    if (treatment_var == "high_income"){
+      highinf_model <- glm(treatment ~ prop_higher_education +
+                           no_insurance + private_insurance + males +
+                           white_pct + black_pct, family = binomial(), data = data_te)
+    }
+    else{
+      highinf_model <- glm(treatment ~ prop_higher_education + pctile +
+                           no_insurance + private_insurance + males +
+                           white_pct + black_pct, family = binomial(), data = data_te)
+    }
+    data_te$prop_score <- predict(highinf_model, type = "response")
+    rr <- Match(Y = data_te$pct_vote_rep, Tr = data_te$treatment, X = data_te$prop_score, M = 1)
+    rr_list[[as.character(cutoff)]] <- rr
+  }
+  summaries <- list()
+
+  for (cutoff in names(rr_list)) {
+    summaries[[cutoff]] <- extract_match_summary(rr_list[[cutoff]])
+  }
+
+  return(summaries)
+}
+cutoffs <- seq(0.1, 0.9, by = 0.1)
+
+summaries_income <- generate_rr_summaries(data, cutoffs, "high_income")
+summaries_dpc <- generate_rr_summaries(data, cutoffs, "high_dpc")
+
+extract_data_from_summaries <- function(summaries) {
+  estimates_list <- list()
+  p_values_list <- list()
+  original_num_obs_list <- list()
+  original_num_treated_list <- list()
+  matched_num_obs_list <- list()
+  matched_num_obs_unweighted_list <- list()
+
+  # Extract variables for each summary into the respective lists
+  for (cutoff in names(summaries)) {
+      estimates_list[cutoff] <- summaries[[cutoff]]$estimate
+      p_values_list[cutoff] <- summaries[[cutoff]]$p
+      original_num_obs_list[cutoff] <- summaries[[cutoff]]$on
+      original_num_treated_list[cutoff] <- summaries[[cutoff]]$ot
+      matched_num_obs_list[cutoff] <- summaries[[cutoff]]$mo
+      matched_num_obs_unweighted_list[cutoff] <- summaries[[cutoff]]$mou
+  }
+
+  # Create a data frame from the lists
+  return(data.frame(
+    cutoffs = c("0.1", "0.2", "0.3", "0.4", "0.5", "0.6", "0.7", "0.8", "0.9"),
+    estimate = unname(unlist(estimates_list)),
+    p = unname(unlist(p_values_list)),
+    on = unname(unlist(original_num_obs_list)),
+    ot = unname(unlist(original_num_treated_list)),
+    mo = unname(unlist(matched_num_obs_list)),
+    mou = unname(unlist(matched_num_obs_unweighted_list))
+  ))
+}
+
+df_income <- extract_data_from_summaries(summaries_income)
+df_dpc <- extract_data_from_summaries(summaries_dpc)
+df_income %>% 
+  rename(`Cutoffs` = cutoffs,
+       `Treatment Effect` = estimate,
+       `P value` = p,
+       `Obs.` = on,
+       `Treat. Obs.` = ot,
+       `Treat. Obs` = mo,
+       `Treat. Obs. Unw.` = mou) %>% 
+kable(booktabs = TRUE) %>% 
+add_header_above(c(" " = 3,
+                   "Original" = 2, "Matched" = 2))
+```
+
+```{r}
+#| label: tbl-dpc
+df_dpc %>% 
+  rename(`Cutoffs` = cutoffs,
+       `Treatment Effect` = estimate,
+       `P value` = p,
+       `Obs.` = on,
+       `Treat. Obs.` = ot,
+       `Treat. Obs` = mo,
+       `Treat. Obs. Unw.` = mou) %>% 
+kable(booktabs = TRUE) %>% 
+add_header_above(c(" " = 3,
+                   "Original" = 2, "Matched" = 2))
+```
+To verify our conclusions, @tbl-income and @tbl-dpc compare the treatment effects for income and DPC at different cutoffs.
+The treatment effects for income is not stable at different cutoffs as it increases as the cutoff increases gradually.
+The cutoff 0.9 may even result in a positive effect.
+Conversely, the treatment effect for DPC at different cutoffs behaves more stable, where all values are negative and around -0.01.
+In addition, even though the optimal cutoff we found, 0.4, is not the maximum one, it is only one that is statistically significant given the 0.05 significance level.
+By the way, there should have no issues of p-hacking as the intial way to choose the cutoff is to via the interaction between the income and dpc.
+
+## The COVID-19 and Trump's voting behaviours {#sec-first-point} Trump's lost can not simply blame on COVID but COVID indeed increase the uncertainty
+
+Throughout American history, Donald Trump is a very personalized president and make him unique compared other presidents in US.
+He was the first person to be elected President of the United States without any previous experience in either political office or military service, but also the first president that had been impeached twice.
+Through his presidency, Trump made many incredible accomplishments.
+US had the "Unprecedented Economic Boom", which seven million new jobs were created and the unemployment rate was the lowest in a half century.
+However, the have been constant accusations against him.
+Trump is the only US president that had been impeached twice during the presidency, given the total four times of impeachment in U.S.
+history.
+The first impeachment was in 2019 which he had solicited foreign countries to interference the 2020 Election and another one is during his last time of presidency, which he had incited the attack to the U.S.
+Capitol.
+However, the most criticized thing against to him is his mis-management of COVID-19 during 2020 and significantly affect his voting for 2020 Election.
+
+The candidates, especially the incumbent president, are always be blamed or credited on the national economy and direction.
+But 2020 Election was a quite different one because the voters need to carefully consider an extra variable "COVID-19" [citation] when they make their decisions.
+The majority of existing researches indicate that Trump was criticized more than praised due to the COVID-19, meaning that people seems to only focus on the COVID-19 and ignore the positive things that Trump made.
+In this paper, @tbl-psm and @tbl-counter shows that the counties with deaths higher than 0.4 quantile have significant less vote for Trump, and Trump can win extra five swing states if he can diminish the inequality.
+However, even though we did not incorporate income into treatment group, we can not ignore its impacts on the voting for Trump as show on @fig-dpc-income.
+The riches counties indeed vote less for Trump.
+
+In fact, if we compare the 2016 and 2020 election, we can observe that income is a crucial factor explaining why Trump lost.
+@fig-compare compares the voting pattern for the counties with extra deaths with cutoff 0.4 quantile in both 2016 and 2020 across different income levels.
+It is clear that the impact of extra death on voting for Trump is enornoumous on those richer counties.
+Specifically, in 2016, the cutoff of counties that start to increase their support to Democrat is about 75th percentile.
+However, this value shrinks to about 60th percentile, meaning that share of votes for richer counties increases compared to 2016.
+Conversely, even though the Republican took a higher share of votes in 2020, it decreases dramatically as income level increases.
+We can verify this from the number of votes for the two parties.
+The mean votes for Trump in 2020 stays very close to 2016 but the gap of votes for Democrat between 2020 and 2016 gets higher and higher as income levels increases.
+Furthermore, this pattern also prove that there were more voters in 2020 compared to 2016, but unfortunately, they all vote for Biden.
+
+```{r}
+#| fig-cap: The comparision of Trump's voting behavior of counties with extra COVID-19 death in 2016 and 2020 Election
+#| label: fig-compare
+#| message: false
+
+library(dplyr)
+library(tidyr)
+
+data1620 <- data %>% 
+  filter(dpc > quantile(dpc, 0.4)) %>% 
+  dplyr::select(fips, pct_vote_demo, pct_vote_demo16,
+                pct_vote_rep, pct_vote_rep16, vote_demo, vote_demo16,
+                vote_rep, vote_rep16)
+
+long_data_pct <- data1620 %>% 
+  pivot_longer(
+    cols = matches("pct_"),
+    names_to = c("party", "year"),
+    names_pattern = "pct_(.+)(..)$",
+    values_to = "vote_pct"
+  ) %>%
+  mutate(year = ifelse(year == "16", "2016", "2020"),
+         party = ifelse(grepl("de", party), "Democrat", "Republican")) %>% 
+  dplyr::select(fips, party, year, vote_pct)
+
+long_data_votes <- data1620 %>%
+  pivot_longer(
+    cols = matches("^vote_"),
+    names_to = c("party", "year"),
+    names_pattern = "vote_(.+)(..)$",
+    values_to = "actual_votes"
+  ) %>%
+  mutate(year = ifelse(year == "16", "2016", "2020"),
+         party = ifelse(grepl("de", party), "Democrat", "Republican")) %>% 
+  dplyr::select(fips, party, year, actual_votes)
+
+# Combine both datasets
+combined_data <- left_join(long_data_pct, long_data_votes, by = c("fips", "year", "party"))
+
+# Determine the Winning Party with both vote percentage and actual votes
+final_data <- combined_data %>%
+  group_by(fips, year) %>%
+  slice_max(vote_pct, n = 1, with_ties = FALSE) %>%
+  ungroup() %>% 
+  left_join(data, by = "fips") %>% 
+  dplyr::select(fips, mean_household_income, vote_pct, actual_votes, party, year)
+
+p1 <- final_data %>% 
+  mutate(income_pctile = ntile(mean_household_income, 100)) %>% 
+  ggplot(aes(x = income_pctile, y = vote_pct, colour = party, linetype = as.factor(year))) +
+  geom_smooth(se = FALSE) +
+  scale_linetype_manual(values = c("2016" = "twodash", "2020" = "solid")) +
+  scale_color_manual(values = c("Democrat" = "#9999FF", "Republican" = "#FF9999"),
+                     name = "Winning Party") +
+  labs(title = "Vote Percentage vs Income Percentile",
+       x = "Income Percentile",
+       y = "Votes Share(%)",
+       linetype = "Year") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+p2 <- final_data %>% 
+  mutate(income_pctile = ntile(mean_household_income, 100)) %>% 
+  ggplot(aes(x = income_pctile, y = actual_votes, colour = party, linetype = as.factor(year))) +
+  geom_smooth(se = FALSE) +
+  scale_linetype_manual(values = c("2016" = "twodash", "2020" = "solid")) +
+  scale_color_manual(values = c("Democrat" = "#9999FF", "Republican" = "#FF9999"),
+                     name = "Party") +
+  labs(title = "Vote Percentage vs Income Percentile",
+       x = "Income Percentile",
+       y = "Number of Votes",
+       linetype = "Year") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+library(cowplot)
+legend <- get_legend(p1 + theme(legend.position = "bottom"))
+combined_plot <- plot_grid(p1, p2, nrow = 1)
+combined_plot_legend <- plot_grid(combined_plot, legend, ncol = 1, rel_heights = c(1, 0.1))
+
+print(combined_plot_legend)
+```
+
+Even though we did not incorporate income into treatment, the losing confidence from richer counties are still unignorable. 书是必然的不管有没有covid
+
+need to focus on state level
+
+## Anticipations of 2024 Election
+One year ago on November 8, 2022, the 2022 United States Election which is also the Midterm Election was held.
+All seats of U.S.
+House of Representatives and 35 seats of 100 was contested to determine the new Congress [citation].
+Historically, people will have a relatively low support for the incumbent president and hence the incumbent president's party will lose a notable number of seats.
+However, even though the Republican had overperformance at traditonal Democrat state such as California and New York, the expected Red Wave was not appeared, with that the Republican only took nine more seats in the Lower House.
+This may not a good news for Trump and his party as in 2018 Midterm Election, the expected "Blue Wave" appears with Democrat taking 45 seats from the Republican.
+Even though COVID was not ended in US, it seems that the decreasing impact of COVID-19 did not bring an expected increase in support for Trump, especially the swing states as we predicted in @tbl-counter.
+
+However, one year later on November 5, 2024, the 2024 Presidential Election will be held and the new President will be elected.
+Joe Biden was already nominated by the Democrat and announced his reelection bid [citation].
+For the Republican party, even though no one has been nominated yet, Donald Trump is far ahead of his challengers within the party [citation], meaning that that the same scene in 2020 will be highly likely to happen again in 2024.
+However, the latest approval rating on Dec 16, 2023 indicates that up to 56% American disapprove the president Biden, which is the near lowest level of his presidency [citation].
+However, the support for Trump is also not optimistic, where only 42% American have favorable opinion on him.
+
+Taking the insights from this paper, it seems that Trump is at least more likely to win the 2024 Federal Election as COVID-19 will not longer be a variable that the voters need to consider.
+However, this paper only illustrate in the perspective of causal inference instead of forecasting.
+To have the accurate predictions of popular votes for Biden and Trump, other techniques like Multilevel with Post-stratification is more suitable.
+
+This paper use the methods of propensity score matching to conduct causal inference detect whether there is a relationship between high death per cases rate and voting for Trump.
+
+if there is a Yes or No question asking whether COVID-19 has negative impact on Trump's voting patterns in 2020, then the answer is Yes.
+However, regarding whether Trump lost because of COVID, the answer is Uncertain.
+
+## Limitations and Future Work
+This paper exploratorily analyze the optimal treatment group and then find the treatment effects using the Propensity Score Matching, as well as a counterfactual analysis.
+However, this paper still have several limitations.
+The data used in this paper are collected from different sources with different geographic levels.
+For Alaska, MIT Election Data Science Club used the census region whereas both ACS and JHU CSSE used different districts.
+Therefore, the data of Alaska is excluded in this research.
+However, since there is only about 10 census regions in Alaska, the exclusion of these data will not significantly influence the accuracy of the analysis.
+
+Another limitation in this paper is that the number of covariates considered to find the propensity scores may be over-simplified.
+The covariates considered in the paper are the significant factors in predicting the COVID-19 mortality rate from taken the previous research [@CaoCOVID19Repo].
+The way to calculate the propensity scores should be more complicate such as adding more predictors or interaction terms.
+A more complicated model can make the predictions of propensity score more accurate and hence the matching process will be more comprehensive.
+The resulting treatment effect is more close to the real one.
+Additionally, the way to conduct the counterfactual analysis is to add the percent of votes from treatment effects directly from Democrat to Republican.
+This might be too punished as there are some other parties which also gain lots of votes, such as Libertarian.
+Decrease the votes proportionally might be more appropriate.
+
+The future works should focus more on the neighborhood effect in voting behavior.
+Harrop et al.
+[citation] argues that neighborhood impacts is a good indicator and they are distributed by the political discussion and local party activity.
+In addition, Ron et al.
+[citation] suggest that the economic factors of communities or counties is most significant related to the voting preferences.
+They also concluded that the similar voters living in different areas will vote for different parties.
+The future works should incorporate neighborhood effect into the research.
+The potential procedure could be to find the treatment effect of high DPC for low and high income counties.
+The future research should also consider the historical voting preference for each county, like those traditionally Republican-counties will not have a hugh increase or decrease in there votes for Trump no matter how the DPC changes.
+Furthermore, the future research should also focus on how to validate the causal inference.
+Parikh et al.
+[citation] introduced a new deep generative model called "Credence" to evaluate the performance of different methods.
+The future research should take a similar way to find the best methods.
 

### 2023-12-20 outputs/paper/paper.qmd
Update the final paper

--- 2023-12-15 outputs/paper/paper.qmd
@@
 ---
-title: "An Observational Study to COVID-19 and 2020 US Presidential Election: Counties with High Deaths per Case Rate Showed Diminished Support For Trump"
+title: "COVID-19 and the 2020 U.S. Presidential Election: Counties with Extra COVID-19 Deaths Showed Less Support For Trump"
 title-block-banner: true
 author: "Yiliu Cao"
-thanks: "Code and data from this analysis are available at: https://github.com/yiliuc/covid_and_trump_loss.git"
+thanks: "Code and data from this analysis are available at: https://github.com/yiliuc/covid_and_trump_loss"
 date: "today"
 date-format: "long"
-abstract: "This study investigates the causal inference between COVID-19 and Donald Trump's loss during the 2020 US Federal Election using the data from MIT Election Data Science Lab and Johns Hopkins University CSSE.
-The main methodology used in this paper is propensity score matching with an exploratory analysis of the optimal treatment.
-The key finding suggests that the counties with a death per case rate exceeding the 0.4 quantile threshold show a reduced voting preference for Trump.
-This paper also conducts a counterfactual analysis based on the treatment effect, indicating that Trump might have been secured to re-elect if the disparities in death per case rates were addressed.
-Future research should refine the propensity score calculations by incorporating additional variables, such as the winning party in each county during the 2016 election."
+abstract: "This study investigates the relationship between extra COVID-19 deaths and Donald Trump's loss during the 2020 U.S.
+Federal Election using the data from MIT Election Data Science Lab and Johns Hopkins University CSSE.
+The main methodology used in this paper is propensity score matching with an exploratory analysis finding the optimal treatment.
+The key finding is that counties with a death per case rate exceeding the 0.4 quantile threshold show a reduced voting preference for Trump.
+Based on the treatment effect, this paper also conducts a counterfactual analysis, indicating that Trump might have been re-elected if the disparities of extra deaths were addressed.
+Future research should incorporate the neighborhood effects on voting between different counties, and validate the causal inference methods used."
 format: pdf
@@
 
-In early 2020, the United States encountered its first cases of COVID-19.
+In early 2020, the United States encountered its first case of COVID-19.
 The situation escalated rapidly, with reported cases reaching 10,000 by March 19 and soaring to 100,000 just eight days later [@WikiCOVIDUSA2023].
-Amidst this unfolding crisis, the U.S.
+Amidst this unfolding crisis, the 2020 U.S.
 Presidential Election was underway, ending with Trump's defeat to Biden.
-By Election Day on November 3, there were more than 9 million reported cases and about 200,000 deaths in the U.S.
+By Election Day on November 3, there were more than nine million reported cases and about 200,000 deaths in the U.S.
 When talking about Trump's loss, the public refers to his mishandling of "his greatest test" [@Greenblatt2021] and believes that his win in the 2016 Election was a historical accident [@Bryant2020].
-Conversely, some arguments suggest that COVID-19 was not the sole or decisive factor in Trump's defeat; his limits as a political strategy and wasting his advantage on the economy also contributed to his loss.
-This paper will investigate the causal effect between COVID and Trump's loss, and if so, to what extent it influenced the 2020 election outcome.
-
-Existing research on this subject provides varied perspectives.
-Baccini et al.
-[@Baccini2021CovidElection] suggested that Trump could win without COVID, but simply using the COVID-19 infection rate as the only factor contributing to his loss would be too naive.
-Socio-economic factors like high education attainment and race diversity are also unignorable.
+Conversely, some arguments suggest that COVID-19 was not the sole or decisive factor in Trump's defeat; his lost is the culmination of self-destructive decisions [@Liptak2020].
+This paper aims to explore the causal relationship between the COVID-19 pandemic and Trump's electoral loss in 2020, examining the extent to how the pandemic influenced the election's outcome.
+
+There are plenty of researches investigating the impacts of COVID-19 on 2020 Election with various perspectives.
+Baccini et al.(2021) suggested that Trump could win without COVID-19, but simply using the COVID-19 infection rate as the only factor contributing to his loss would be too naive.
+Socio-economic factors like high education attainment and race diversity are also essential in explaining the popular voting patterns.
 Besides, their research emphasized the significant impact of COVID-19 on those urban areas where there were no stay-at-home orders, particularly for the "swing" states.
-In addition, Noland & Zhang [@NolandZhang2021] suggested that the deaths per case is a more crucial metric than the infection rate when analyzing the impact of COVID-19 on voting for Trump.
-However, they highlight the challenge in assessing COVID-19's impact on voting, as we may not know when the voters made their decisions.
-In contrast, Clarke et al.
-[@Clarke2021CovidTrump] conducted actual surveys before and after the election, suggesting that COVID-19 impacts voting but not the dominant one.
-The U.S.'s highly polarized political landscape might be more significant.
-Similarly, Hart [@Hart2021CovidTrump] also conducted a survey to ask about people's attitudes toward Trump.
+In addition, Noland and Zhang (2021) suggested that the deaths per case (DPC) rate is a more appropriate metric than the infection rate or mortality rate when analyzing the impact of COVID-19 on voting for Trump.
+However, they highlight the challenge in assessing COVID-19's impact on voting, as we may not know when the voters made their decision.
+In contrast, Clarke, Stewart, and Ho (2021) conducted surveys before and after the election, suggesting that COVID-19 is but not the dominant factor affecting the voting.
+The U.S.'s polarized political landscape might be more significant.
+Similarly, Hart (2021) also conducted a survey to ask about people's attitudes toward Trump.
 All the participants show natural or negative possession.
-They also find that some social movements like "Murder of George Floyd" may impact voting.
-
-While all the above research investigated the correlation between COVID-19 and voting for Trump, previous studies predominantly employed models to identify variables influencing Trump's vote share and perform counterfactual analyses to estimate his voting performance in the absence of COVID-19 or reduction in deaths.
-Instead of using that method, this study employs propensity score matching to find the treatment effect and analyze whether counties more severely affected by COVID-19 exhibit distinct voting patterns.
+They also find that some social movements like "Black Lives Matter" may have potential influence on the voting.
+
+While all the above research investigated the correlation between COVID-19 and voting for Trump, previous studies predominantly employed models to identify variables significantly influencing Trump's vote share and perform counterfactual analyses to estimate his voting performance in the absence of COVID-19 or reduction in deaths.
+Instead, this study will employ the propensity score matching to find the causal effects between COVID-19 and Trump's lost.
+Furthermore, Wallace et al.
+(2023) highlights the the extra deaths during the pandemic are closely associated to a county's political preference.
+Building on the findings of Noland and Zhang (2021), the primary metric of COVID-19 in this analysis will be the excess death rate, specifically a high Death per Case (DPC) rate.
 However, given all counties have experienced COVID-19, this paper will first find the optimal treatment group(s).
-Finally, I will conduct a counterfactual analysis to see if Trump could be re-elected with the treatment effect eliminated.
-
-This paper will have five parts.
-I will introduce the data used in this paper and present data summaries and visualization in the @sec-data part.
-After that, I will explain the methods in this paper and the corresponding results in the @sec-methods and @sec-results parts, respectively.
-All the results will be interpreted and discussed in the @sec-discussion part, and I will conclude with limitations and drawbacks in the @sec-conclusion part.
+Then we will also conduct a counterfactual analysis to see if Trump could be re-elected with the treatment effect eliminated.
+
+The remainder of this paper is structured to five parts.
+@sec-data introduce the data used in the paper and summarize important variables.
+@sec-methods explains the methods used and @sec-results present all the results including the optimal treatment group, the treatment effect and the counterfactual analysis.
+All the results will be discussed in the @sec-discussion, as well as limitations and future works.
+The main findings are summarized in @sec-conclusion.
 
@@
 `R` [@R] will be the computer language used in the paper.
-Moreover, this paper will also use different R packages.
-The data is cleaned using the package `dplyr` [@dplyr], `stringr` [@stringr], `tidyverse` [@tidyverse], `janiotr` [@janitor], `tools` [@tools] and `tidyr` [@tidyr].
-In addition to that, `ggplot2` [@ggplot2], `RColorBrewer` [@RColorBrewer], `maps` [@maps], `mapdata` [@mapdata], `gridExtra` [@gridExtra] and `cowplot` [@cowplot] will be used later to make tables and plot graphs.
-Besides, to perform Propensity Score Matching, this paper will also employ `Matching` [@Matching].
+The data is cleaned using the package `dplyr` [@dplyr], `stringr` [@stringr], `tidyverse` [@tidyverse], `janiotr` [@janitor], `tools` [@R] and `tidyr` [@tidyr].
+In addition to that, `ggplot2` [@ggplot2], `RColorBrewer` [@RColorBrewer], `maps` [@maps], `mapdata` [@mapdata], `gridExtra` [@gridExtra] and `cowplot` [@cowplot] will be used to visualize the data and `kableExtra` [@kableExtra] is used to make tables.
+To perform propensity score matching, this paper will also employ `Matching` [@Matching].
 
@@
 
-This paper comprised five data sets from three different sources, each corresponding to different topics.
+This paper comprised five data sets from three sources with each corresponding to different topics.
 The primary data is from the MIT Election Data Science Club [@MITElectionLab], which builds open online data collections of the U.S.
@@
 The data extracted is called "County Presidential Election Returns 2000-2020," [@MITElectionData2022] with about 70,000 rows containing the voting patterns for each candidate and party by county since 2000.
-Besides that, the data also indicate the types of voting, such as "EARLY VOTE" and "ELECTION DAY." To analyze the voting patterns for Donald J.
-Trump, I only filter the 2020 U.S.
+Besides that, the data also indicate the types of voting, such as "EARLY VOTE" and "ELECTION DAY." To analyze the voting patterns for Trump, we only filter the 2020 U.S.
 Federal Election data for all counties and parties.
@@
 During the pandemic, they collected the U.S.
-and international COVID cases and deaths, reporting them by daily summary on their GitHub [@CSSECOVID19Data].
+and international COVID cases and deaths, reporting them by daily report on their GitHub [@CSSECOVID19Data].
 To examine the impact of COVID-19 on the 2020 Election as accurately as possible, this paper used the daily report on November 3, 2020, which is the election day of the 2020 U.S.
 Federal Election.
-The resulting data include the aggregate number of cases, deaths and recovers before Election Day and the incidence rate and case-fatality ratio for each county.
-
-This paper also incorporates the socioeconomic data from the American Community Survey (ACS) [@acs], which is an online open-source database conducted by the U.S.
+The resulting data include the aggregate number of cases, deaths and recovers before Election Day and the incidence rate for each county.
+
+This paper also uses the socioeconomic data from the American Community Survey (ACS) [@acs], which is an online open-source database conducted by the U.S.
 Census Bureau, containing the various socioeconomic factors at different geographic levels.
 The data sets extracted from ACS are 2020 five-year estimates of DP02, DP03 and DP05, covering the social, economic and demographic characteristics at the county level.
-Since there are thousands of variables, to ensure simplicity, I will only use the variables that are found to be significant in predicting the COVID-19 mortality rate from my previous paper [@CaoCOVID19Repo].
+Since there are thousands of variables, to ensure simplicity, we will only use the variables that are found to be significant in predicting the COVID-19 mortality rate from the previous research [@CaoCOVID19Repo].
 The descriptions of all variables can be found on @tbl-descriptions.
 
-After having five data sets, they are merged into one big data by county.
-Using the merged data, I calculate the percentage of votes for each party (candidate) in the 2020 election.
-In addition, I also create new dummy variables indicating the winning party in both elections for each county.
-Moreover, to accurately compare the COVID cases and deaths across all counties, I transform the number of cases and deaths to infection and mortality rate by 10,000 citizens in each county.
-Using the conclusion from Noland & Zhang's work [@NolandZhang2021], I also calculate each county's deaths per case rate.
+The five data sets are merged into one big data by county.
+Using the merged data, we calculate the percentage of votes for each party (candidate) in the 2020 election.
+In addition, we also create new dummy variables indicating the winning party for each county.
+Moreover, to accurately compare the COVID cases and deaths across all counties, we transform the number of cases and deaths to infection and mortality rate by 10,000 citizens in each county.
+Using the conclusion from Noland & Zhang's work [@NolandZhang2021], we also calculate each county's deaths per 100K case (DPC) rate.
 The final data consists of 3107 rows with 36 columns.
@@
                             "The COVID mortality rate, calculated by deaths per 10,000 residences",
-                            "The COVID death per 10,000 confirmed cases",
+                            "The COVID deaths per 10,000 confirmed cases",
                             "The percentage of votes for the Democrat in 2020",
@@
          `Descriptions` = descriptions) %>% 
-  kable(booktabs = TRUE)
+  kable(booktabs = TRUE) %>% 
+  column_spec(column = 3, width = "10cm")
 ```
@@
   summarise(n = n(),
-            case = mean(cases),
-            inf = mean(infrate),
-            deaths = mean(deaths),
-            mort = mean(mortrate),
-            dpc = mean(dpc),
-            income = mean(mean_household_income)) %>% 
-  rename(`Winning Party` = winning_party,
-         `Count` = n,
+            case = round(mean(cases), 0),
+            inf = round(mean(infrate), 0),
+            deaths = round(mean(deaths), 0),
+            mort = round(mean(mortrate), 0),
+            dpc = round(mean(dpc), 0),
+            income = round(mean(mean_household_income), 0)) %>% 
+  rename(`Party` = winning_party,
+         `County Win` = n,
          `Average Cases` = case,
@@
          `Income` = income) %>% 
-  kable(booktabs = TRUE)
-```
-
-@tbl-covid compares the COVID-19 impacts and income levels across the counties won by each party.
-Even though Republicans won in approximately five-sixths of the counties, counties that supported Biden reported nearly ten times the average number of COVID-19 cases and deaths compared to those that supported Trump.
+  kable(booktabs = TRUE) %>% 
+  column_spec(2:6, width = "1.7cm") %>% 
+  footnote(general = "DPC: Deaths per 100K Cases")
+```
+
+@tbl-covid compares the COVID-19 impacts and income levels across the counties that each party won at 2020.
+Even though Republicans won in approximately five-sixths of the counties, counties supporting Biden reported nearly as ten times high as the average number of COVID-19 cases and deaths compared to those counties supporting Trump.
 Despite similar infection and mortality rates between the two groups, there is still a notable higher death-to-case rate in counties where Democrats won.
-These patterns suggest that the counties voting for Democrats are more severely impacted by COVID-19 and are more population-intensive than those voting for the Republicans.
-Even though only around 500 counties support Democrats, they are densely populated urban areas and wealthier, so Biden can have more electoral votes.
-We can verify this from income levels, where the counties voting Democrat has a higher mean income.
+These patterns suggest that the counties voting for Democrats are more severely impacted by COVID-19.
+Besides, Biden won the election with about 500 counties, this also indicate that these counties are densely populated urban areas and hence have more electoral votes.
+We can verify this from @fig-map, which compares the relative ratio of votes for the two parties and the death to cases rate in maps.
+It is clear that more counties had a preference for the Republicans, but the counties with higher DPC rates show a preference for Democrats.
+These counties, such as California and New York, are more affluent.
+In contrast, the states that are less impacted by COVID-19, such as Utah, vote more for Republicans.
 
@@
                       limits = c(0, 5000), oob = scales::squish) +
-  labs(title = "The distribution of death per case rate by 100K",
-       fill = 'Death per cases by 100k') +
+  labs(title = "The deaths per 100K cases (DPC)",
+       fill = 'Deaths per Case') +
   coord_quickmap() +
@@
 
-We can validate our guess from @fig-map, which compares the relative ratio of votes for the two parties and the death to cases rate in maps.
-From the maps, it is clear more counties had a preference for the Republicans, but the counties with higher death-per-case rates show a preference for Democrats.
-These counties, such as California and New York, are mostly more affluent.
-In contrast, the states that are less impacted by COVID-19, such as Utah, vote more for Republicans.
-The maps match our findings from @tbl-covid.
-
-```{r}
+Given the insights from @tbl-covid and @fig-map, @fig-dpc-income delves deeper into the impact of COVID and income levels on the voting behaviors of the two parties.
+The left side of @fig-dpc-income displays the share of votes, while the right side focuses on the total number of votes.
+In terms of the share of votes, in general, there is a positive relationship between the DPC and votes for the Democrats but a negative one for the Republicans.
+However, the counties with a medium value of DPC seem less sensitive to the voting for the two parties, whereas the support of counties at the 'tails' shifts more significantly with changes in the DPC.
+Meanwhile, there are clear quadratic patterns between the vote for the two parties and income.
+The lower-income counties tend to favour Republicans, but as income increases, a greater number of counties lean towards voting for Biden rather than Trump.
+
+```{r}
+#| fig-width: 8
+#| fig-height: 5
 #| fig-cap: The correlation between voting behaviors to income and DPC rate for the Republican and Democrat
@@
 
-Given the insights from @tbl-covid and @fig-map, @fig-dpc-income delves deeper into the impact of COVID and income levels on the voting behaviours of the two parties.
-The left side of @fig-dpc-income displays the share of votes, while the right side focuses on the total number of votes, with each party represented by distinct colours.
-In terms of the percentage of votes, in general, there is a positive relationship between the DPC and votes for the Democrats but a negative one for the Republicans.
-However, the counties with a medium value of DPC seem less sensitive to the voting for the two parties, whereas the support of counties at the 'tails' shifts more significantly with changes in the DPC.
-Meanwhile, there are clear quadratic patterns between the vote for the two parties and income.
-The lower-income counties tend to favour Republicans, but as income increases, a greater number of counties lean towards voting for Biden rather than Trump.
-
 In addition, despite the average share of votes for Republicans being consistently above 0.5, it does not imply a universal loss for Democrats across all counties.
 As detailed in @tbl-covid, we know that the number of counties won by Republicans is approximately five times greater than those won by Democrats.
-With Republicans dominating more counties, they naturally have a higher average share of votes.
-This pattern also indicates that Democrats usually won the densely populated, urban areas, but Republicans won in more sparsely populated and rural regions [@Spencer2023ElectionFactCheck].
-The high population density areas have more votes, which explains why Biden beat Trump.
-We can verify this from the right side where the number of votes for Biden exceeds Trump as DPC and income level increases.
+In addition, Democrat had less "overwhelming" victories compared to the Republican.
+Democrat won the densely populated, urban areas, but Republicans won in more sparsely populated and rural regions [@Spencer2023ElectionFactCheck].
+The high population density areas have more electoral votes, which explains why Biden beat Trump despite a relatively low share of votes.
+We can verify this from the right two graphs where the number of votes for Biden exceeds Trump as DPC and income level increases.
 This pattern is particularly pronounced in the wealthiest top 50% of counties and those with DPCs exceeding the 75th percentile.
+
+@fig-insurance shows the correlation between the DPC rate and the proportion of people with private and without health insurance for the two parties.
+Unsurprisingly, there is an inverse correlation between the DPC rate and private health insurance – as private insurance coverage increases, the DPC rate tends to decrease.
+Notably, counties that favoured the Democratic Party exhibit a higher proportion of individuals with private insurance compared to Republican-leaning counties.
+In addition, in Republican-dominated counties, there is a pronounced positive correlation between the DPC rate and the lack of health insurance but a relatively flat for counties where Democrat won.
+These findings suggest that counties voting Democratic generally have more comprehensive insurance coverage, which appears to mitigate the impact of DPC rates to a greater extent than in Republican counties.
 
@@
 #| label: fig-insurance
+#| fig-width: 8
+#| fig-height: 4
+
+binned_data3 <- data %>%
+  mutate(private_insurance = round(private_insurance, 0)) %>%  
+  group_by(private_insurance, winning_party) %>%
+  summarize(dpc = mean(dpc))
+
+binned_data4 <- data %>%
+  mutate(no_insurance = round(no_insurance, 0)) %>%  
+  group_by(no_insurance, winning_party) %>%
+  summarize(dpc = mean(dpc))
 
 plot1 <- ggplot(data, aes(x = private_insurance)) +
-  #geom_point(aes(y = dpc, colour = winning_party), alpha = 0.25) +  
+  geom_point(aes(y = dpc, colour = winning_party), alpha = 0.1) +  
   geom_smooth(aes(y = dpc, colour = winning_party), 
@@
 plot2 <- ggplot(data, aes(x = no_insurance)) +
-  #geom_point(aes(y = dpc, colour = winning_party), alpha = 0.25) +  
+  geom_point(aes(y = dpc, colour = winning_party), alpha = 0.1) +  
   geom_smooth(aes(y = dpc, colour = winning_party), 
@@
                       labels = c("Democrat", "Republican")) +
-  theme(legend.position = "none") + 
-  labs(x = "No Health Insurance(%)", 
-       y = "DPC rate")
+  theme(legend.position = "none",
+        axis.title.y = element_blank()) + 
+  labs(x = "No Health Insurance(%)")
 
@@
 
-@fig-insurance shows the correlation between the DPC rate and the proportion of people with private and without health insurance for the two parties.
-Unsurprisingly, the graph reveals an inverse correlation between the DPC rate and private health insurance – as private insurance coverage increases, the DPC rate tends to decrease.
-Notably, counties that favoured the Democratic Party exhibit a higher proportion of individuals with private insurance compared to Republican-leaning counties.
-In contrast, in Republican-dominated counties, there is a pronounced positive correlation between the DPC rate and the lack of health insurance.
-However, it is relatively flat for Democratic counties.
-These findings suggest that counties voting Democratic generally have more comprehensive insurance coverage, which appears to mitigate the impact of DPC rates to a greater extent than in Republican counties.
+@fig-bar compares the ratio of the White population (race diversity) and education levels for the two parties.
+Interestingly, it suggests that counties with higher racial diversity are less likely to vote for Biden, as opposed to the trend observed for Republican-leaning counties, where increased diversity aligns with increased support.
+The average share of votes for Trump even reach 0.9 in the counties with the highest proportion of white residents.
+In terms of the ratio of people with high education attainment, since the average vote for Republicans is always higher than for Democrats, this indicate that educated people are more likely to vote for Trump.
+These observations are in line with the conclusions drawn by Baccini et al.
+(2021).
 
@@
 #| label: fig-bar
+#| fig-width: 8
+#| fig-height: 4
+
 data_we <- data %>%
@@
   scale_fill_manual(values = c("Republican" = "#FF9999", "Democrat" = "#9999FF")) +
-  labs(x = "Ratio of High-Edu. (Deciles)", y = "Mean Votes(%)", fill = "Party") +
-  theme_minimal() +
-  theme(legend.position = "none")
+  labs(x = "Ratio of High-Edu. (Deciles)", fill = "Party") +
+  theme_minimal() +
+  theme(legend.position = "none",
+        axis.title.y = element_blank())
 
@@
 
-@fig-bar compares the ratio of the White population (race diversity) and education levels for the two parties.
-Interestingly, it suggests that counties with higher racial diversity are less likely to vote for Biden, as opposed to the trend observed for Republican-leaning counties, where increased diversity aligns with increased support.
-This pattern implies that counties that predominantly vote for Democrats have a higher proportion of White residents.
-This may explain why they have relatively higher income levels.
-In terms of the ratio of people with high education attainment, since the average vote for Republicans is always higher than for Democrats, this may indicate that more educated people are likely to vote for Trump.
-
-# Methods{#sec-meethods}
+# Methods{#sec-methods}
 
@@
 The primary method implemented in this study will be the Propensity Score Matching (PSM).
-After finding the causal inference, this paper will also conduct a counterfactual analysis to see whether Trump can re-elect without COVID-19.
+After finding the treatment effects, this paper will also conduct a counterfactual analysis to see whether Trump can re-elect without COVID-19.
 
@@
 The group of people receiving the treatment is called the treatment group, and the control group for the rest.
-For instance, Austin [@Austin2011PropensityScore] examined the effectiveness of a new medical treatment called clampless off-pump coronary artery bypass (clampless OPCAB) to see if it can reduce the in-hospital mortality rate compared to the traditional method.
-Here, 'treatment' refers to the new surgical procedure, and patients who took this new surgery are the treatment group.
+For instance, Austin [@Austin2011PropensityScore] examined the effectiveness of a new medical treatment called clampless off-pump coronary artery bypass (clampless OPCAB) to see if the patients taking this treatment can have a lower in-hospital mortality rate compared to the patients taking the traditional surgery.
+Here, "treatment" refers to the new surgical procedure, and patients who took this new surgery are the treatment group.
 They compared the mortality rates between the two groups and concluded that the new treatment could statistically lower the stroke and mortality rate.
@@
 Typically, the treatment should be defined before conducting the analysis, and this aims to reduce the risk of p-hacking [@Frost2023PHacking].
-However, given the objective of this paper, which is to find whether COVID-19 influenced Trump's loss in the last Election, pre-defining the treatment is challenging as all U.S.
-counties were affected by COVID-19.
+However, given the objective of this paper, which is to find whether extra COVID-19 deaths leads to Trump's loss in the last Election, pre-defining the treatment is challenging as all U.S.
+counties experienced COVID-19.
 Therefore, this study begins with exploratory analysis to identify the treatment group exhibiting the most significant differences in voting patterns for Trump.
-I will establish specific cutoffs for classifying counties as having 'high' or 'low' death per case rates (DPC).
+We will establish specific cutoffs for classifying counties as having "high" or "low" death per case rates (DPC).
 In addition, from @fig-dpc-income, low and high-income counties also seem to have different voting behaviours.
@@
 P-hacking means the manipulation of data analysis until we can find statistically significant results [@Frost2023PHacking].
-If we simply choose one cutoff and see the significant results that match common sense, it may raise the p-hacking risk, and the analyses will be less convincing.
-Therefore, we will employ a grid of cutoffs for each variable, examining how treatment effects vary.
+If we simply choose one cutoff and see the significant results that match common sense, it may raise the p-hacking risk, and the analyses will be less convincing and comprehensive.
+Therefore, we will employ a grid of cutoffs for each variable, examining how treatment effect varies.
 For example, one treatment group can be counties with death per case and income levels higher than 0.3 quantiles.
-In this way, we should be able to find the optimal treatment group and avoid the risk of p-hacking.
+Using this way, we are not only able to find the optimal treatment group eventually, but avoid the risk of p-hacking.
 
@@
 
-Randomized controlled trials (RCT) are regarded as the most ideal and the golden way to estimate the treatment effect [@Kuss2016PropensityScore].
+Randomized controlled trials (RCT) are regarded as the most ideal way to estimate the treatment effect [@Kuss2016PropensityScore].
 The reason is that RCTs are the only way that guarantee the equal distribution of known and unknown parameters.
 Therefore, the outcomes between the treatment and control groups will not be confounded.
-However, conducting an RCT is nearly impractical and impossible in many scenarios.
+However, conducting an RCT is difficult in many scenarios.
 Back to the coronary artery bypass example [@Austin2011PropensityScore], it's not ethically feasible to randomly assign patients to new or traditional treatments.
-Hence, conducting an RCT is nearly impossible.
-
 In such situations, implementing Propensity Score Matching (PSM) can be a more practical way to find the treatment effects.
-The propensity score is the probability of an observation that receives the treatment based on the existing covariates using logistic regression.
-Using propensity scores can reduce the dimension [@Zhao2021PropensityR], meaning that we can describe an observation to a single score instead of multiple covariates.
-In addition, another significant advantage of PSM is design separation [@Zhao2021PropensityR], meaning that PSM can separate the covariates balancing and effects estimation.
-This is especially beneficial as we can observe from @fig-insurance that the distribution of counties with relatively high DPC rates is imbalanced.
-Finally, we can match two observations having similar propensity scores but from treatment and control groups separately.
-Then, we will have many matched pairs and estimate the treatment effects.
-The matching process is called propensity score matching.
-
-In this paper, the propensity scores are used to estimate how likely a county is to receive treatment using multiple linear logistic regression. The equation will look like:
-
-```{=tex}
-\begin{align*}
-\log\left(\frac{P(T=1|X)}{1 - P(T=1|X)}\right) = \beta_0 + \beta_1 X_1 +\cdots +\beta_p X_p
-\end{align*}
-```
-where:\
-* $P(T=1|X)$ is the probability of a county receiving treatment\
-* $X_j\:(1\leq j\leq p)$ are the covariates\
+
+The propensity score is the probability an observation would have been treated based on the existing covariates using logistic regression.
+With propensity scores, it can help us to reduce the dimension [@Zhao2021PropensityR], meaning that we can describe an observation by a single value instead of multiple covariates.
+Another significant advantage of PSM is design separation [@Zhao2021PropensityR], meaning that PSM can separate the covariates balancing and effects estimation.
+This is especially beneficial as we can observe from @fig-insurance that the distribution of people with and without (private) insurance is unbalanced for the counties with a high and low DPC rate.
+Using propensity scores, we can match two observations with similar propensity scores, but from treatment and control groups, separately.
+We can then estimate the treatment effect based on many matched pairs with different propensity score.
+Therefore, propensity score mathcing can solve the limitation that we can not estimate causal effects from observational studies.
+
+In this paper, propensity scores are used to estimate how likely a county is to receive treatment using multiple linear logistic regression. The equation is:
+
+$$
+\begin{aligned}
+\mbox{Pr}(y_i = 1) = & \text{logit}^{-1}(\beta_0 + \beta_1 \times \text{prop\_higher\_education}\\
+& + \beta_2 \times \text{income\_pctile} + \beta_3 \times \text{no\_insurance} \\
+& + \beta_4 \times \text{private\_insurance} + \beta_5 \times \text{males} \\
+& + \beta_6 \times \text{white\_pct} + \beta_7 \times \text{black\_pct})
+\end{aligned}
+$$
+where:
+
+* $\mbox{Pr}(y_i = 1)$ is the probability that a county has a "High" DPC rate
 * $\beta_0$ is the intercept
-* $\beta_j\:(1\leq j\leq p)$ is the corresponding coefficients of each covariate\
+* $\beta_j\:(1\leq j\leq 7)$ is the corresponding coefficients for seven predictors
 
@@
 
-## Choise of treatment groups
+## Choice of treatment groups
+
+@fig-dpc shows the correlation between votes for Trump and income levels, segmented into "high DPC" and "low DPC" counties at varying cutoffs.
+For instance, a cutoff of 0.3 classifies counties with a DPC rate below 0.3 quantile as "low DPC" and those above as "high DPC." By comparing these groups across different cutoffs, we aim to identify the cutoff where the difference in voting for Trump across different income levels is most pronounced.
+From the graph, we can find that, except for the cutoff 0.4, the two lines converge as income increases (right tail).
+Only for the 0.4 cutoff, the lines of the two groups converge at the middle income but do not overlap and diverge at the tails.
+Additionally, regardless of the cutoff applied, there is a common pattern that the support for Trump increases in poorer counties up to around the eighth bin, which is approximately the 0.4 income quantile, and then decreases in wealthier counties.
+This indicates a significant voting disparity between counties below and above the 0.4 income quantile.
 
@@
 #| fig-height: 5
-#| fig-cap: 'The Correlation Between Voting for Trump and Income Levels, Categorized by DPC Rate at Varied Quantile Cutoffs (e.g., Top 40% as High-DPC and bottom 60% as Low-DPC for "Cutoff: 0.6")'
+#| fig-cap: The Correlation Between Voting for Trump and Income Levels, Categorized by DPC Rate at Varied Quantile Cutoffs
 #| label: fig-dpc
@@
                                                         c(0.4, 0.8)))
-
-library(cowplot)
-legend <- get_legend(plots[[1]] + theme(legend.position = "bottom",  # Place legend at the bottom
-        legend.justification = "center",  # Justify legend in the center
+legend <- get_legend(plots[[1]] + theme(legend.position = "bottom",
+        legend.justification = "center",
         legend.box.just = "bottom"))
@@
 
-@fig-dpc shows the correlation between votes for Trump and income levels, segmented into "High-DPC" and "Low-DPC" county groups at varying DPC cutoffs.
-For instance, a cutoff of 0.3 classifies counties with a DPC rate below this threshold as "Low DPC" and those above as "High DPC." By comparing these groups across different cutoffs, we aim to identify the cutoff where the difference in voting for Trump across different income levels is most pronounced.
-That said, we can find that, except for the cutoff 0.4, the two lines converge as income increases (right tail).
-Only for the 0.4 cutoff, the lines of the two groups converge at the middle income but do not overlap and diverge at the tails.
-Additionally, regardless of the cutoff applied, there is a common pattern that the support for Trump increases in poorer counties up to around the eighth bin, which is approximately the 0.4 income quantile, and then decreases in wealthier counties.
-This indicates a significant voting disparity between counties below and above the 0.4 income quantile.
+@fig-income shows the correlation between Trump's vote share and the DPC rate but this time dividing counties based on income levels into "high Income" and "low Income" groups.
+Each subplot in this analysis corresponds to a different income cutoff.
+Compared to @fig-dpc, all curves, regardless of the cutoff, display a linear or near-linear relationship with voting patterns.
+As the income cutoff increases, the difference in voting behaviour between the high and low-income counties becomes more distinct, especially at the 0.9 cutoff.
+This suggests that in counties with higher incomes, voting patterns vary significantly from those in lower-income counties as the DPC rate changes.
+Therefore, the most significant pattern is observed at the 0.9 income cutoff, contradicting the patterns identified in @fig-dpc.
 
@@
 #| fig-height: 5
-#| fig-cap: 'The Correlation Between Voting for Trump and DPC Rate, Categorized by Income Levels at Varied Quantile Cutoffs (e.g., Top 40% as High Income and bottom 60% as Low Income for "Cutoff: 0.6")'
+#| fig-cap: The Correlation Between Voting for Trump and DPC Rate, Categorized by Income Levels at Varied Quantile Cutoffs
 #| label: fig-income
@@
 
-@fig-income shows the correlation between Trump's vote share and the DPC rate but this time dividing counties based on income levels into "High Income" and "Low Income" groups.
-Each subplot in this analysis corresponds to a different income cutoff.
-Compared to @fig-dpc, all curves, regardless of the cutoff, display a linear or near-linear relationship with voting patterns.
-As the income cutoff increases, the difference in voting behaviour between the high and low-income counties becomes more distinct, especially at the 0.9 cutoff.
-This suggests that in counties with higher incomes, voting patterns vary significantly from those in lower-income counties as the DPC rate changes.
-Therefore, the most significant pattern is observed at the 0.9 income cutoff, contradicting the patterns identified in @fig-dpc.
-
-So, a question may arise regarding whether we should include all the income and DPC in the settings of treatment.
-The answer is NO, and we should only include the DPC.
-From @fig-dpc, a consistent quadratic relationship between income and voting patterns for Trump is evident regardless of how the cutoff for defining high DPC is adjusted.
+A question may arise regarding whether we should incorporate both the income and DPC in the settings of treatment.
+We include only the DPC because, from @fig-dpc, a consistent quadratic relationship between income and voting patterns for Trump is evident regardless of how the cutoff for defining high DPC is adjusted.
 This pattern demonstrates that poorer counties (in the first half of the income spectrum) may show a positive or a negative correlation with voting for Trump depending on the cutoff we choose.
@@
 We should exclude income because the choice of cutoff will significantly change the treatment effects; DPC is more "stable" than income.
-We can verify this from @fig-income, where each line shows a negative slope, meaning that the choice of cutoff for the "High DPC" group will not change the general voting patterns for Trump.
+We can verify this from @fig-income, where each line shows a negative slope, meaning that the choice of cutoff for the "high DPC" group will not change the general voting patterns for Trump.
 Hence, including income would introduce extra variability in the treatment effects and increase the probability of p-hacking.
-The @fig-treatment shows a simple visualization of why this is the case.
+
+@fig-treatment shows a simple example when we test different cutoffs for DPC rate and income.
+From the graph, we can see that no matter which cutoff we set, the low DPC and high DPC counties always have a negative value of slope.
+The only difference is the absolute values of the slope.
+Conversely, the impact of income levels shows notable variability.
+For instance, at a cutoff of the 0.4 quantiles, low and high-income groups display opposite voting patterns for Trump.
+However, when we increase the cutoff to the 0.9 quantile, both income groups negatively correlate with Trump's vote share.
+This underscores the instability of income.
 
@@
 #| fig-width: 8
-#| fig-height: 3
+#| fig-height: 4
 #| label: fig-treatment
@@
 
-@fig-treatment shows a simple example when we test different cutoffs for DPC rate and income.
-From the graph, we can see that no matter which cutoff we set, the low DPC and high DPC counties always have a negative value of slope.
-The only difference is the absolute values of the slope.
-Conversely, the impact of income levels shows notable variability.
-For instance, at a cutoff of the 0.4 quantiles, low and high-income groups display opposite voting patterns for Trump.
-However, when we increase the cutoff to the 0.9 quantile, both income groups negatively correlate with Trump's vote share.
-This underscores the instability of income.
-
 From @fig-dpc, we have determined that a cutoff at the 0.4 quantile is the most effective threshold for distinguishing between high and low Death Per Case (DPC) rate groups.
@@
 
-In my previous paper, I already showed that the proportion of residents with at least a bachelor's degree, income, the proportion of people with private insurance and without health insurance, the ratio of males, ratio of black and black residents are statistically in predicting the probability of a county has a high mortality rate or not.
-In this paper, I will directly use the results and only consider these variables when predicting the propensity scores.
-
-```{=tex}
-\begin{align*}
-\log\left(\frac{P(T=1|X)}{1 - P(T=1|X)}\right) &= 8.06 \\
-& -0.02 \times \text{prop\_higher\_education} \\
-& +0.01 \times \text{pctile} \\
-& +0.01 \times \text{no\_insurance} \\
-& -0.03 \times \text{private\_insurance} \\
-& -0.13 \times \text{males} \\
-& +0.00 \times \text{white\_pct} \\
-& +0.06 \times \text{black\_pct}
-\end{align*}
-```
+Cao (2023) already showed that the proportion of residents with at least a bachelor's degree, income, the proportion of people with private insurance and without health insurance, the ratio of males, ratio of black and black residents are statistically in predicting the probability of a county has a high mortality rate or not.
+Here, we will directly use the results and only consider these variables when predicting the propensity scores.
+
+$$
+\begin{aligned}
+\mbox{Pr}(y_i = 1) = & \text{logit}^{-1}(8.06 - 0.016 \times \text{prop\_higher\_education}\\
+& + 0.012 \times \text{income\_pctile} + 0.013 \times \text{no\_insurance} \\
+& - 0.034 \times \text{private\_insurance} - 0.128 \times \text{males} \\
+& + 0.003 \times \text{white\_pct} + 0.06 \times \text{black\_pct})
+\end{aligned}
+$$
 
 The above equation shows the logistic regression model to predict the propensity score for each county.
-Males seem to be the most critical factor when predicting the propensity score.
+The proportion of males seems to be mostly correlated to the high DPC rate.
 Perhaps the majority of COVID deaths are males.
@@
 This may explain why richer counties are less likely to vote for Trump.
+In addition, the role of health insurance is unignorable.
+The counties with high private health insurance coverage seems to significantly less likely to have extra COVID-19 deaths.
+
+Using the treatment we found previously, @tbl-psm summarize the results of Propensity Score Matching.
+The treatment effect of about -0.018 means that the counties with a DPC rate higher than 0.4 quantile will, on average, have a 0.018 less vote for Trump compared to the counties with a low DPC rate.
+This value is statistically significant as its p-value is about 0.02.
+In addition, there are 1869 treatment observations in both original and matched data, meaning that some counties in the control group are matched more than once.
 
@@
 
-@tbl-psm summarize the results of Propensity Score Matching.
-The treatment effect of about -0.018 means that the counties with a DPC rate higher than 0.4 will, on average, vote 0.018 less for Trump compared to the counties with a lower DPC rate.
-This value is statistically significant as its p-value is about 0.02, lower than 0.05.
-In addition, there are 1869 treatment observations in both original and matched data, meaning that some counties in the control group are matched more than once.
+To test the performance of PSM in balancing the covariates, @tbl-balance summarizes each variable's mean values in the control group and standard mean difference before and after the matching.
+Notably, there is a significant decrease in the standard mean difference for each variable, indicating that the PSM effectively reduced the imbalance between the treatment and control groups.
+Besides, we can see that the mean values in the control group approximately remain the same after the matching.
+This denote the reliability of the matching and suggest that PSM has successfully aligned the groups based on the observed covariates.
 
@@
 
-@tbl-balance summarizes each variable's mean values in the control group and standard mean difference before and after the matching.
-Notably, there is a significant decrease in the standard mean difference for each variable.
-This indicates that the Propensity Score Matching effectively reduced the imbalance between the treatment and control groups.
-Besides, we can see that the mean values in the control group approximately remain the same after the matching, enhancing the reliability of the matching process and suggesting that PSM has successfully aligned the groups based on the observed covariates.
-
 ## Counterfactual Analysis
 
-```{r}
-#| tbl-cap: The summary of counterfactual votes for the eleven swing staes in 2020
+@tbl-counter provides a comparative summary of the actual and counterfactual voting results in eleven swing states during the 2020 U.S.
+Presidential election.
+The counterfactual votes is calculated based on the treatment effects.
+That said, given other variables constant, Trump will have about 1.8% more votes in the counties with a DPC rate higher than 0.4 quantile, whereas Democrat will lose the same amount of votes in these counties.
+In reality, Trump only won three swing states and the total electoral votes was 232.
+However, the counterfactual analysis suggests that Trump would have won five additional states if the DPC rate disparities were eliminated, bringing him an extra 52 electoral votes.
+Then his total electoral votes will be 284, surpassing the required 270-vote threshold for reelection.
+Therefore, if Trump can eliminate the extra deaths, he could be re-elected.
+
+```{r}
+#| tbl-cap: The summary of counterfactual votes for the eleven swing states in 2020
 #| label: tbl-counter
@@
 data_new <- data_te %>%
-  mutate(vote_rep = ifelse(treatment == 1, total_votes*(pct_vote_rep + 0.018342), vote_rep),
-         vote_demo = ifelse(treatment == 1, total_votes*(pct_vote_demo - 0.018342), vote_demo))
+  mutate(vote_rep2 = ifelse(treatment == 1, total_votes*(pct_vote_rep + 0.018342), vote_rep),
+         vote_demo = ifelse(treatment == 1, vote_demo - (vote_rep2 - vote_rep), vote_demo),
+         vote_rep = vote_rep2)
 
@@
   group_by(state) %>% 
-  summarise(total_votes_rep_counter = sum(vote_rep),
-            total_votes_demo_counter = sum(vote_demo))
+  summarise(total_votes_rep_counter = round(sum(vote_rep), 0),
+            total_votes_demo_counter = round(sum(vote_demo), 0))
 
@@
   kable(booktabs = TRUE) %>% 
-  add_header_above(c(" " = 2, "Actual Results" = 3, "Simulated Results*" = 3))
-```
-
-@tbl-counter provides a comparative summary of the actual and counterfactual voting results in eleven swing states during the 2020 U.S.
-Presidential election.
-In reality, Trump only won three swing states with 232 electoral votes.
-However, the counterfactual analysis suggests that Trump would have won five additional states if the DPC rate disparities were eliminated, bringing him an extra 52 electoral votes.
-This would bring his total to 284 electoral votes, surpassing the required 270-vote threshold for reelection.
-Therefore, if Trump can eliminate the extra deaths, he could be re-elected.
+  add_header_above(c(" " = 2, "Actual Results" = 3, "Counterfactual Results*" = 3))
+```
 
@@
 
-## First discussion point {#sec-first-point}
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
+## Defining the treatment is the most crucial part in this study
+
+In an observational study, especially when conducting the causal inference, the treatment should always be settled before the analysis as it can effectively reduce the risk of p-hacking.
+If we want to focus on whether COVID-19 has generally influenced the voting for Trump compared to 2016, then we can take the voting for each county in both 2016 and 2020, and the treatment group would simply be the counties in 2020.
+However, this paper aims to delve deeper to see whether the extra COVID-19 deaths during the pandemic has influenced the voting for Trump.
+As all the U.S.
+counties have experienced COVID-19, it is necessary to manually define the cutoff of "high" extra COVID-19 deaths.
+Therefore, this study starts with an exploratory analysis to find the optimal treatment group.
+
+From @sec-data, it seems that a county's COVID death per case rate (DPC) and income levels are the two factors most related to voting for Trump.
+Therefore, the treatment will only incorporate these two factors and our objective is to define the cutoffs that maximize the difference between the treatment and control group.
+However, from @fig-dpc and @fig-income, it seems that the effect of income on the voting for Trump is highly dependent on the choice of cutoff as it appears a quadratic relationship.
+A quadratic pattern is not desired as different cutoffs will bring significant differences in the treatment effects.
+Furthermore, from @fig-treatment, we can observe that the two groups show an opposite relation to voting when the cutoff is low but the same relation for a higher cutoff.
+In contrast, DPC shows a general cubic relationship to Trump's vote; hence, the choice of cutoff will not generate huge different treatment effects.
+We can verify this from @fig-income, where no matter how the cutoff for defining "high" income changes, there will always be a linear or at least a closely linear relationship between them and the shift in cutoff will not bring a dramatic change in treatment effects (see @fig-treatment).
+Besides, from @fig-income, we can observe that the "high" and "low" DPC rate counties behave most differently with a cutoff at 0.4 quantiles.
+Therefore, the optimal treatment group is counties with a DPC rate higher than the 0.4 quantile.
+
+To verify our conclusions, @tbl-income and @tbl-dpc compare the treatment effects for income and DPC at different cutoffs.
+From @tbl-income, we can observe that the treatment effects for income are unstable at different cutoffs.
+It increases gradually as we increase the cutoff and even shows a positive value if we set it to 0.9.
+This indicates that the voting difference for Trump between the high and low-income counties increases as cutoff increases.
+However, the reason why a cutoff 0.9 will generate a postive treatment effect is the both high and low income groups shows a negative correlation to the vote for Trump (see @fig-treatment).
+Hence we should not incorporate income into the treatment.
+
+```{r}
+#| label: tbl-income
+#| tbl-cap: The treatment effect of income under different cutoffs (exlude income from calculation of propensity scores)
+# show TE with different income levels
+generate_rr_summaries <- function(data, cutoffs, treatment_var) {
+  rr_list <- list()
+  for (cutoff in cutoffs) {
+    data_te <- data %>%
+      mutate(high_dpc = ifelse(dpc > quantile(dpc, cutoff), 1, 0),
+             high_income = ifelse(mean_household_income > quantile(mean_household_income, cutoff), 1, 0))
+    data_te$treatment <- data_te[[treatment_var]]
+    if (treatment_var == "high_income"){
+      highinf_model <- glm(treatment ~ prop_higher_education +
+                           no_insurance + private_insurance + males +
+                           white_pct + black_pct, family = binomial(), data = data_te)
+    }
+    else{
+      highinf_model <- glm(treatment ~ prop_higher_education + pctile +
+                           no_insurance + private_insurance + males +
+                           white_pct + black_pct, family = binomial(), data = data_te)
+    }
+    data_te$prop_score <- predict(highinf_model, type = "response")
+    rr <- Match(Y = data_te$pct_vote_rep, Tr = data_te$treatment, X = data_te$prop_score, M = 1)
+    rr_list[[as.character(cutoff)]] <- rr
+  }
+  summaries <- list()
+
+  for (cutoff in names(rr_list)) {
+    summaries[[cutoff]] <- extract_match_summary(rr_list[[cutoff]])
+  }
+
+  return(summaries)
+}
+cutoffs <- seq(0.1, 0.9, by = 0.1)
+
+summaries_income <- generate_rr_summaries(data, cutoffs, "high_income")
+summaries_dpc <- generate_rr_summaries(data, cutoffs, "high_dpc")
+
+extract_data_from_summaries <- function(summaries) {
+  estimates_list <- list()
+  p_values_list <- list()
+  original_num_obs_list <- list()
+  original_num_treated_list <- list()
+  matched_num_obs_list <- list()
+  matched_num_obs_unweighted_list <- list()
+
+  # Extract variables for each summary into the respective lists
+  for (cutoff in names(summaries)) {
+      estimates_list[cutoff] <- summaries[[cutoff]]$estimate
+      p_values_list[cutoff] <- summaries[[cutoff]]$p
+      original_num_obs_list[cutoff] <- summaries[[cutoff]]$on
+      original_num_treated_list[cutoff] <- summaries[[cutoff]]$ot
+      matched_num_obs_list[cutoff] <- summaries[[cutoff]]$mo
+      matched_num_obs_unweighted_list[cutoff] <- summaries[[cutoff]]$mou
+  }
+
+  # Create a data frame from the lists
+  return(data.frame(
+    cutoffs = c("0.1", "0.2", "0.3", "0.4", "0.5", "0.6", "0.7", "0.8", "0.9"),
+    estimate = unname(unlist(estimates_list)),
+    p = unname(unlist(p_values_list)),
+    on = unname(unlist(original_num_obs_list)),
+    ot = unname(unlist(original_num_treated_list)),
+    mo = unname(unlist(matched_num_obs_list)),
+    mou = unname(unlist(matched_num_obs_unweighted_list))
+  ))
+}
+
+df_income <- extract_data_from_summaries(summaries_income)
+df_dpc <- extract_data_from_summaries(summaries_dpc)
+df_income %>% 
+  rename(`Cutoffs` = cutoffs,
+       `Treatment Effect` = estimate,
+       `P value` = p,
+       `Obs.` = on,
+       `Treat. Obs.` = ot,
+       `Treat. Obs` = mo,
+       `Treat. Obs. Unw.` = mou) %>% 
+kable(booktabs = TRUE) %>% 
+add_header_above(c(" " = 3,
+                   "Original" = 2, "Matched" = 2))
+```
+
+Compared to the income, the treatment effect for DPC at different cutoffs behaves more stable, where all values are negative and around -0.01.
+Besides, even though the number of treatment observations is the same as income, the number of unweighted treatment observations when setting treatment using DPC is much larger than income.
+This represents a higher number of samples, and therefore more accurate estimates.
+In addition, even though the optimal cutoff we found, 0.4 quantiles, does not have the maximum treatment effect, it is the only statistically significant one given 0.05 significance level.
+
+```{r}
+#| label: tbl-dpc
+#| tbl-cap: The treatment effect of income under different cutoffs
+
+df_dpc %>% 
+  rename(`Cutoffs` = cutoffs,
+       `Treatment Effect` = estimate,
+       `P value` = p,
+       `Obs.` = on,
+       `Treat. Obs.` = ot,
+       `Treat. Obs` = mo,
+       `Treat. Obs. Unw.` = mou) %>% 
+kable(booktabs = TRUE) %>% 
+add_header_above(c(" " = 3,
+                   "Original" = 2, "Matched" = 2))
+```
+
+## Beyond COVID-19, chain reaction of losing confidence from richer votes may be more decisive on Trump's loss.
+
+Throughout American history, Donald Trump has been a very personalized president, which makes him unique compared to other presidents in the U.S.
+He was the first president of the United States without any previous experience in either political office or military service, but also the first president that had been impeached twice.
+Throughout his presidency, Trump made many incredible accomplishments like the "Unprecedented Economic Boom" [@TrumpWH2021].
+However, there have been constant accusations against him.
+Trump is the only U.S.
+president who has been impeached twice during the presidency [@ImpeachmentTrump2023], given the total four times of impeachment in U.S.
+history.
+The first impeachment was in 2019 when he had solicited foreign countries to interfere in the 2020 Election, and another one was which he incited an attack on the U.S.
+Capitol.
+However, when talk about Trump, the most criticized thing against him is his mismanagement of COVID-19.
+
+The candidates, especially the incumbent president, are always blamed or credited for the national economy and direction.
+But the 2020 Election was quite different because the voters needed to consider an extra variable "COVID-19" when making decisions.
+Most existing research indicates that Trump was criticized more than praised due to COVID-19, meaning that people only focus on COVID-19 and ignore the positive things that Trump made.
+In this paper, we see from @tbl-psm and @tbl-counter that the counties with deaths higher than 0.4 quantiles have significantly fewer votes for Trump, and Trump can win an extra five swing states if he can diminish the inequality.
+However, even though we did not incorporate income into the treatment group, we can not ignore its impacts on voting for Trump, as shown on @fig-dpc-income.
+The wealthy counties indeed vote less for Trump.
+
+Income seems to have a higher impact on voting then we expected.
+@fig-compare compares the voting pattern for the same treatment group in both 2016 and 2020 U.S.
+Election across different income levels.
+The impact of extra deaths on voting for Trump is enormous in those more affluent counties.
+Specifically, in 2016, the cutoff of counties that started to increase their support to Democrats is about 75th percentile.
+However, this value shrinks to about the 60th percentile, meaning that the share of votes for Democrats from richer counties increases compared to 2016.
+Conversely, although Trump generally took a higher share of votes in 2020, it decreased dramatically as income increased.
+We can verify this from the number of votes for the two parties.
+The mean votes for Trump in 2020 stay very close to 2016, but the gap of votes for Democrats between the two election gets higher and higher as income levels increase.
+Furthermore, this pattern also proves that there were more voters in 2020 than in 2016, but unfortunately, they all voted for Biden.
+
+```{r}
+#| fig-cap: The comparision of Trump's voting behavior of counties with extra COVID-19 death in 2016 and 2020 Election
+#| label: fig-compare
+#| message: false
+#| fig-width: 8
+#| fig-height: 5
+
+library(dplyr)
+library(tidyr)
+
+data1620 <- data %>% 
+  filter(dpc > quantile(dpc, 0.4)) %>% 
+  dplyr::select(fips, pct_vote_demo, pct_vote_demo16,
+                pct_vote_rep, pct_vote_rep16, vote_demo, vote_demo16,
+                vote_rep, vote_rep16)
+
+long_data_pct <- data1620 %>% 
+  pivot_longer(
+    cols = matches("pct_"),
+    names_to = c("party", "year"),
+    names_pattern = "pct_(.+)(..)$",
+    values_to = "vote_pct"
+  ) %>%
+  mutate(year = ifelse(year == "16", "2016", "2020"),
+         party = ifelse(grepl("de", party), "Democrat", "Republican")) %>% 
+  dplyr::select(fips, party, year, vote_pct)
+
+long_data_votes <- data1620 %>%
+  pivot_longer(
+    cols = matches("^vote_"),
+    names_to = c("party", "year"),
+    names_pattern = "vote_(.+)(..)$",
+    values_to = "actual_votes"
+  ) %>%
+  mutate(year = ifelse(year == "16", "2016", "2020"),
+         party = ifelse(grepl("de", party), "Democrat", "Republican")) %>% 
+  dplyr::select(fips, party, year, actual_votes)
+
+# Combine both datasets
+combined_data <- left_join(long_data_pct, long_data_votes, by = c("fips", "year", "party"))
+
+# Determine the Winning Party with both vote percentage and actual votes
+final_data <- combined_data %>%
+  group_by(fips, year) %>%
+  slice_max(vote_pct, n = 1, with_ties = FALSE) %>%
+  ungroup() %>% 
+  left_join(data, by = "fips") %>% 
+  dplyr::select(fips, mean_household_income, vote_pct, actual_votes, party, year)
+
+p1 <- final_data %>% 
+  mutate(income_pctile = ntile(mean_household_income, 100)) %>% 
+  ggplot(aes(x = income_pctile, y = vote_pct, colour = party, linetype = as.factor(year))) +
+  geom_smooth(se = FALSE) +
+  scale_linetype_manual(values = c("2016" = "twodash", "2020" = "solid")) +
+  scale_color_manual(values = c("Democrat" = "#9999FF", "Republican" = "#FF9999"),
+                     name = "Winning Party") +
+  labs(x = "Income Percentile",
+       y = "Votes Share(%)",
+       linetype = "Year") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+p2 <- final_data %>% 
+  mutate(income_pctile = ntile(mean_household_income, 100)) %>% 
+  ggplot(aes(x = income_pctile, y = actual_votes, colour = party, linetype = as.factor(year))) +
+  geom_smooth(se = FALSE) +
+  scale_linetype_manual(values = c("2016" = "twodash", "2020" = "solid")) +
+  scale_color_manual(values = c("Democrat" = "#9999FF", "Republican" = "#FF9999"),
+                     name = "Party") +
+  labs(x = "Income Percentile",
+       y = "Number of Votes",
+       linetype = "Year") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+library(cowplot)
+legend <- get_legend(p1 + theme(legend.position = "bottom"))
+combined_plot <- plot_grid(p1, p2, nrow = 1)
+combined_plot_legend <- plot_grid(combined_plot, legend, ncol = 1, rel_heights = c(1, 0.1))
+
+print(combined_plot_legend)
+```
+
+## Anticipations of 2024 Election
+
+On November 8, 2022, the 2022 United States Election, which is also the Midterm Election, was held.
+All seats of the U.S.
+House of Representatives and 35 seats of 100 were contested to determine the new Congress [@2022USElection].
+Historically, people will have a relatively low support for the incumbent president, and hence the incumbent president's party will lose a notable number of seats.
+However, even though the Republican had overperformance in traditional Democrat states such as California and New York, the expected "Red Wave" was not appeared, with the Republican only taking nine more seats from Democrat in the Lower House [@2022USElection].
+This is not be good news for Trump and his party, as in the 2018 Midterm Election, the expected "Blue Wave" appeared with Democrats taking 45 seats from the Republicans [@2018USElection].
+Even though COVID was not officially ended and still had impacts on the U.S.
+in last year miterm election, it seems that the decreasing effect of COVID-19 did not bring an expected increase in support for Trump, especially the swing states, as we predicted in @tbl-counter.
+
+However, on November 5, 2024, the 2024 Presidential Election will be held, and the new President will be elected.
+Both parties have not nominated any candidates so far.
+However, Joe Biden was seeking another term at the White House, and Trump is far ahead of his challengers within the party [@2024election], meaning that the same scene in 2020 will likely happen again in 2024.
+However, the latest approval rating indicates that up to 53% of Americans disapprove of President Biden, which is the near lowest level of his presidency [@BidenOpinion].
+Meanwhile, the support for Trump is also not optimistic, where 54.9% of Americans have a favourable opinion of him [@TrumpPolls].
+
+Taking the insights from this paper, it seems that Trump is at least more likely to win the next election than 2020, as COVID-19 will no longer be a variable that the voters need to consider.
+However, this paper only illustrates the perspective of causal inference in 2020 instead of forecasting the next election.
+Other techniques, like Multilevel with Post-stratification, are more suitable for accurate predictions of popular votes for Biden and Trump in 2024.
+
+## Limitations and Future Work
+
+This paper exploratorily analyzes the optimal treatment group and then finds the treatment effects using the propensity score matching, as well as a counterfactual analysis.
+However, this paper still has several limitations.
+The data used in this paper are collected from different sources at different geographic levels.
+MIT Election Data Science Club used the census region for Alaska, whereas both ACS and JHU CSSE used "districts." Therefore, the data of Alaska is excluded from this research.
+However, since there are only about ten census regions in Alaska, the exclusion of these data will not significantly influence the accuracy of the analysis.
+
+Another limitation of this paper is that the number of covariates considered to find the propensity scores may be over-simplified.
+The covariates considered in the paper are the significant factors in predicting the COVID-19 mortality rate from the previous research [@CaoCOVID19Repo].
+The way to calculate the propensity scores should be more complicated, such as adding more predictors or interaction terms.
+A more complex model can make the predictions of propensity scores more comprehensive, and hence, the resulting treatment effect will also be closer to the real one.
+Additionally, the way to conduct the counterfactual analysis is to add the percent of votes from treatment effects directly from Democrat to Republican.
+This might be too punished for Democrats as some other parties also won lots of votes, such as Libertarian.
+Re-assigning the votes proportionally might be more appropriate.
+
+Future works should focus more on the neighbourhood effect on voting behaviour.
+Harrop et al.
+[@Harrop2007] argue that neighbourhood impacts are a good indicator, and they are distributed by political discussion and local party activity.
+In addition, Ron et al.
+[@SpatialPolarization] suggest that the economic factors of communities or counties are significantly related to voting preferences.
+Similar voters living in different areas will vote for different parties.
+The potential procedure to incorporate the neighbourhood effect in this study could be to find the treatment effect for low and high-income counties, respectively.
+Additionally, the historical voting preference also matters for each county.
+For example, traditionally Republican counties may not have a considerable increase or decrease in their votes for Trump no matter how the DPC changes.
+Furthermore, future research should also focus on how to validate the methods used to conduct causal inference.
+Parikh et al.
+[@validatingci] introduced a new deep generative model called "Credence" to evaluate the performance of different methods.
+Future research should take a similar approach to evaluate the reliability of the methods used and find the best one.
 
 # Conclusion{#sec-conclusion}
+In conclusion, this paper uses the methods of propensity score matching to conduct causal inference to detect whether there is a relationship between extra COVID-19 deaths and voting for Trump in 2020.
+This paper shows that the extra COVID-19 deaths had a negative impact on Trump's voting patterns in 2020, and Trump could re-elect if he can address this disparity.
+However, this paper also shows that the effects of COVID-19 on popular votes for Trump are not the sole or decisive factor causing him to lose in 2020.
+Losing confidence from richer voters may be another reason.
+To analyze the impact of COVID-19 on voting more comprehensively, the neighbourhood impact should also be considered.
 

### 2023-12-20 scripts/00-simulate_data.R
Update the scripts

@@ -1,11 +1,10 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Simulate the data
+# Author: Yiliu Cao
+# Date: Dec 20 2023
+# Contact: yiliu.cao@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: None
 
 
 #### Workspace setup ####

### 2023-12-20 scripts/01-download_data.R

@@ -1,13 +1,10 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from ACS API
+# Purpose: Downloads and saves the data from ACS API and JHU CSSE
 # Author: Yiliu Cao
-# Date: 1 October 2023
+# Date: Dec 20 2023
 # Contact: yiliu.cao@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: None
-
-
-#### Workspace setup ####
 
 
 #### Download data ####
@@ -30,6 +27,9 @@ url_covid2 <- paste0(
   "01-20-2021.csv")
 
 covid_data <- read.csv(url_covid, header = TRUE, sep = ",")
+
+# The election data is downloaded directly from their website at 
+# https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/VOQCHQ
 
 #### Save data ####
 write.csv(DP02, "inputs/data/DP_02.csv", row.names = FALSE)

### 2023-12-20 scripts/02-data_cleaning.R

@@ -1,11 +1,10 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Data cleaning and export the cleaned data
+# Author: Yiliu Cao
+# Date: Dec 20 2023
+# Contact: yiliu.cao@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: None
 
 #### Workspace setup ####
 library(dplyr)

### 2023-12-21 outputs/paper/paper.qmd

@@
 
-There are plenty of researches investigating the impacts of COVID-19 on 2020 Election with various perspectives.
+There are plenty of researchers investigating the impacts of COVID-19 on 2020 Election with various perspectives.
 Baccini et al.(2021) suggested that Trump could win without COVID-19, but simply using the COVID-19 infection rate as the only factor contributing to his loss would be too naive.
@@
 Furthermore, Wallace et al.
-(2023) highlights the the extra deaths during the pandemic are closely associated to a county's political preference.
+(2023) highlights the extra deaths during the pandemic are closely associated to a county's political preference.
 Building on the findings of Noland and Zhang (2021), the primary metric of COVID-19 in this analysis will be the excess death rate, specifically a high Death per Case (DPC) rate.
@@
 
-@tbl-covid compares the COVID-19 impacts and income levels across the counties that each party won at 2020.
+@tbl-covid compares the COVID-19 impacts and income levels across the counties that each party won in 2020.
 Even though Republicans won in approximately five-sixths of the counties, counties supporting Biden reported nearly as ten times high as the average number of COVID-19 cases and deaths compared to those counties supporting Trump.
@@
 ```{r}
-#| fig-cap: The correlation between DPC rate with proportons of people with private insurance and without health insurance for the two parties
+#| fig-cap: The correlation between DPC rate with proportions of people with private insurance and without health insurance for the two parties
 #| label: fig-insurance
@@
 Interestingly, it suggests that counties with higher racial diversity are less likely to vote for Biden, as opposed to the trend observed for Republican-leaning counties, where increased diversity aligns with increased support.
-The average share of votes for Trump even reach 0.9 in the counties with the highest proportion of white residents.
+The average share of votes for Trump even reached 0.9 in the counties with the highest proportion of white residents.
 In terms of the ratio of people with high education attainment, since the average vote for Republicans is always higher than for Democrats, this indicate that educated people are more likely to vote for Trump.
@@
 Therefore, this study begins with exploratory analysis to identify the treatment group exhibiting the most significant differences in voting patterns for Trump.
-We will establish specific cutoffs for classifying counties as having "high" or "low" death per case rates (DPC).
+We will establish specific cut-offs for classifying counties as having "high" or "low" death per case rates (DPC).
 In addition, from @fig-dpc-income, low and high-income counties also seem to have different voting behaviours.
@@
 
-Nevertheless, choosing cutoffs is critical, and we need to avoid p-hacking or data dredging risks.
+Nevertheless, choosing cut-offs is critical, and we need to avoid p-hacking or data dredging risks.
 P-hacking means the manipulation of data analysis until we can find statistically significant results [@Frost2023PHacking].
 If we simply choose one cutoff and see the significant results that match common sense, it may raise the p-hacking risk, and the analyses will be less convincing and comprehensive.
-Therefore, we will employ a grid of cutoffs for each variable, examining how treatment effect varies.
+Therefore, we will employ a grid of cut-offs for each variable, examining how treatment effect varies.
 For example, one treatment group can be counties with death per case and income levels higher than 0.3 quantiles.
@@
 We can then estimate the treatment effect based on many matched pairs with different propensity score.
-Therefore, propensity score mathcing can solve the limitation that we can not estimate causal effects from observational studies.
+Therefore, propensity score matching can solve the limitation that we cannot estimate causal effects from observational studies.
 
@@
 
-@fig-dpc shows the correlation between votes for Trump and income levels, segmented into "high DPC" and "low DPC" counties at varying cutoffs.
-For instance, a cutoff of 0.3 classifies counties with a DPC rate below 0.3 quantile as "low DPC" and those above as "high DPC." By comparing these groups across different cutoffs, we aim to identify the cutoff where the difference in voting for Trump across different income levels is most pronounced.
+@fig-dpc shows the correlation between votes for Trump and income levels, segmented into "high DPC" and "low DPC" counties at varying cut-offs.
+For instance, a cutoff of 0.3 classifies counties with a DPC rate below 0.3 quantile as "low DPC" and those above as "high DPC." By comparing these groups across different cut-offs, we aim to identify the cutoff where the difference in voting for Trump across different income levels is most pronounced.
 From the graph, we can find that, except for the cutoff 0.4, the two lines converge as income increases (right tail).
@@
 #| fig-height: 5
-#| fig-cap: The Correlation Between Voting for Trump and Income Levels, Categorized by DPC Rate at Varied Quantile Cutoffs
+#| fig-cap: The Correlation Between Voting for Trump and Income Levels, Categorized by DPC Rate at Varied Quantile Cut-offs
 #| label: fig-dpc
@@
 #| fig-height: 5
-#| fig-cap: The Correlation Between Voting for Trump and DPC Rate, Categorized by Income Levels at Varied Quantile Cutoffs
+#| fig-cap: The Correlation Between Voting for Trump and DPC Rate, Categorized by Income Levels at Varied Quantile Cut-offs
 #| label: fig-income
@@
 
-@fig-treatment shows a simple example when we test different cutoffs for DPC rate and income.
+@fig-treatment shows a simple example when we test different cut-offs for DPC rate and income.
 From the graph, we can see that no matter which cutoff we set, the low DPC and high DPC counties always have a negative value of slope.
@@
   scale_color_manual(values = c("65" = "lightgrey", "40" = "#E69F00"),
-                       name = "Cutoffs",
+                       name = "Cut-offs",
                        labels = c("65" = "0.65 Quantile", "40" = "0.4 Quantile")) +
@@
   scale_color_manual(values = c("90" = "lightgrey", "40" = "#E69F00"),
-                       name = "Cutoffs",
+                       name = "Cut-offs",
                        labels = c("90" = "0.90 Quantile", "40" = "0.4 Quantile")) +
@@
 Besides, we can see that the mean values in the control group approximately remain the same after the matching.
-This denote the reliability of the matching and suggest that PSM has successfully aligned the groups based on the observed covariates.
+This denotes the reliability of the matching and suggest that PSM has successfully aligned the groups based on the observed covariates.
 
@@
 #| label: tbl-balance
-#| tbl-cap: The balance of each covariates before and after the mathcing
+#| tbl-cap: The balance of each covariate before and after the matching
 # Load necessary libraries
@@
 The counterfactual votes is calculated based on the treatment effects.
-That said, given other variables constant, Trump will have about 1.8% more votes in the counties with a DPC rate higher than 0.4 quantile, whereas Democrat will lose the same amount of votes in these counties.
+That said, given other variables constant, Trump will have about 1.8% more votes in the counties with a DPC rate higher than 0.4 quantile, whereas Democrats will lose the same number of votes in these counties.
 In reality, Trump only won three swing states and the total electoral votes was 232.
 However, the counterfactual analysis suggests that Trump would have won five additional states if the DPC rate disparities were eliminated, bringing him an extra 52 electoral votes.
-Then his total electoral votes will be 284, surpassing the required 270-vote threshold for reelection.
+Then his total electoral votes will be 284, surpassing the required 270-vote threshold for re-election.
 Therefore, if Trump can eliminate the extra deaths, he could be re-elected.
@@
 From @sec-data, it seems that a county's COVID death per case rate (DPC) and income levels are the two factors most related to voting for Trump.
-Therefore, the treatment will only incorporate these two factors and our objective is to define the cutoffs that maximize the difference between the treatment and control group.
+Therefore, the treatment will only incorporate these two factors and our objective is to define the cut-offs that maximize the difference between the treatment and control group.
 However, from @fig-dpc and @fig-income, it seems that the effect of income on the voting for Trump is highly dependent on the choice of cutoff as it appears a quadratic relationship.
-A quadratic pattern is not desired as different cutoffs will bring significant differences in the treatment effects.
+A quadratic pattern is not desired as different cut-offs will bring significant differences in the treatment effects.
 Furthermore, from @fig-treatment, we can observe that the two groups show an opposite relation to voting when the cutoff is low but the same relation for a higher cutoff.
@@
 
-To verify our conclusions, @tbl-income and @tbl-dpc compare the treatment effects for income and DPC at different cutoffs.
-From @tbl-income, we can observe that the treatment effects for income are unstable at different cutoffs.
+To verify our conclusions, @tbl-income and @tbl-dpc compare the treatment effects for income and DPC at different cut-offs.
+From @tbl-income, we can observe that the treatment effects for income are unstable at different cut-offs.
 It increases gradually as we increase the cutoff and even shows a positive value if we set it to 0.9.
 This indicates that the voting difference for Trump between the high and low-income counties increases as cutoff increases.
-However, the reason why a cutoff 0.9 will generate a postive treatment effect is the both high and low income groups shows a negative correlation to the vote for Trump (see @fig-treatment).
+However, the reason why a cutoff 0.9 will generate a postive treatment effect is the both high- and low-income groups shows a negative correlation to the vote for Trump (see @fig-treatment).
 Hence we should not incorporate income into the treatment.
@@
 #| label: tbl-income
-#| tbl-cap: The treatment effect of income under different cutoffs (exlude income from calculation of propensity scores)
+#| tbl-cap: The treatment effect of income under different cut-offs (exlude income from calculation of propensity scores)
 # show TE with different income levels
@@
 df_income %>% 
-  rename(`Cutoffs` = cutoffs,
+  rename(`Cut-offs` = cutoffs,
        `Treatment Effect` = estimate,
@@
 
-Compared to the income, the treatment effect for DPC at different cutoffs behaves more stable, where all values are negative and around -0.01.
+Compared to the income, the treatment effect for DPC at different cut-offs behaves more stable, where all values are negative and around -0.01.
 Besides, even though the number of treatment observations is the same as income, the number of unweighted treatment observations when setting treatment using DPC is much larger than income.
@@
 #| label: tbl-dpc
-#| tbl-cap: The treatment effect of income under different cutoffs
+#| tbl-cap: The treatment effect of income under different cut-offs
 
 df_dpc %>% 
-  rename(`Cutoffs` = cutoffs,
+  rename(`Cut-offs` = cutoffs,
        `Treatment Effect` = estimate,
@@
 In this paper, we see from @tbl-psm and @tbl-counter that the counties with deaths higher than 0.4 quantiles have significantly fewer votes for Trump, and Trump can win an extra five swing states if he can diminish the inequality.
-However, even though we did not incorporate income into the treatment group, we can not ignore its impacts on voting for Trump, as shown on @fig-dpc-income.
+However, even though we did not incorporate income into the treatment group, we cannot ignore its impacts on voting for Trump, as shown on @fig-dpc-income.
 The wealthy counties indeed vote less for Trump.
@@
 ```{r}
-#| fig-cap: The comparision of Trump's voting behavior of counties with extra COVID-19 death in 2016 and 2020 Election
+#| fig-cap: The comparison of Trump's voting behavior of counties with extra COVID-19 death in 2016 and 2020 Election
 #| label: fig-compare
@@
 However, even though the Republican had overperformance in traditional Democrat states such as California and New York, the expected "Red Wave" was not appeared, with the Republican only taking nine more seats from Democrat in the Lower House [@2022USElection].
-This is not be good news for Trump and his party, as in the 2018 Midterm Election, the expected "Blue Wave" appeared with Democrats taking 45 seats from the Republicans [@2018USElection].
+This is not good news for Trump and his party, as in the 2018 Midterm Election, the expected "Blue Wave" appeared with Democrats taking 45 seats from the Republicans [@2018USElection].
 Even though COVID was not officially ended and still had impacts on the U.S.
-in last year miterm election, it seems that the decreasing effect of COVID-19 did not bring an expected increase in support for Trump, especially the swing states, as we predicted in @tbl-counter.
+in last year's mid-term election, it seems that the decreasing effect of COVID-19 did not bring an expected increase in support for Trump, especially the swing states, as we predicted in @tbl-counter.
 
